### 2015-02-05 doc/uo2_paper/uo2-II.tex
AM: generated tex for NR

%% LyX 2.0.6 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,twocolumn,showkeys,superscriptaddress]{revtex4}
\usepackage[T1]{fontenc}
\setcounter{secnumdepth}{3}
\usepackage{color}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{esint}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\@ifundefined{textcolor}{}
{%
 \definecolor{BLACK}{gray}{0}
 \definecolor{WHITE}{gray}{1}
 \definecolor{RED}{rgb}{1,0,0}
 \definecolor{GREEN}{rgb}{0,1,0}
 \definecolor{BLUE}{rgb}{0,0,1}
 \definecolor{CYAN}{cmyk}{1,0,0,0}
 \definecolor{MAGENTA}{cmyk}{0,1,0,0}
 \definecolor{YELLOW}{cmyk}{0,0,1,0}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{mathptmx}
\usepackage{times}
\usepackage{microtype}
\def\thetable{\arabic{table}}

\makeatother

\usepackage{babel}
\begin{document}

\title{Uranyl solvation by the three dimensional reference interaction site
model}

\author{Alexei Matveev}

\email{alexei.matveev@gmail.com}

\selectlanguage{english}%

\author{Bo Li}

\affiliation{Department Chemie, Technische Universität München, 85747 Garching,
Germany}

\author{Notker Rösch}

\affiliation{Department Chemie and Catalysis Research Center, Technische Universität
München, 85747 Garching, Germany}

\date{Jan 2015}
\begin{abstract}
We report an implementation of the three-dimensional reference interaction
site model (3D RISM) specifically describing the treatment of the
long-range Coulomb field of charged species represented by point charges
and/or distributed charge density. A comparison of 1D and 3D results
for atomic ions demonstrates the reasonable accuracy even for moderate
box size and grid resolution. In an application to uranyl aqua complexes
with 4\textendash{}6 explicit aqua ligands and an implicit bulk solvent
modeled by RISM, we demonstrate that the 3D technique is not susceptible
to the deficiencies of 1D technique exposed in our previous work {[}Li,
Matveev, Krüger, Rösch, \emph{Comp. Theor. Chem.} \textbf{1051}, p.~151,
(2015){]}. The 3D method eliminates the artificial superposition of
explicit aqua ligands and the RISM medium. It predicts essentially
the same values for uranyl and uranyl-water bond lengths as a state
of the art polarizable continuum model. We find that with the first
solvation shell treated explicitly, the observables are nearly independent
of the closure order. The water exchange activation barrier obtained
in a hybrid approach combining the quantum mechanical description
of the uranyl interaction within the first solvation shell and the
RISM model for the bulk solvent is in close agreement with the experimental
and the best theoretical estimates. 
\end{abstract}

\keywords{Reference interaction site model, aqueous solution, uranyl, solvation}

\maketitle

\section{Introduction}

Thermodynamic and structural properties of the uranyl(VI) cation,
UO$_{2}^{2+}$, the most common uranium species in aqueous solution
are crucial for understanding its environmental chemistry as well
as chemical processes in radioactive waste disposal and recycling
of used nuclear fuel \cite{book:Morss06,book:Grenthe92,Silva95}.
Quantum mechanics (QM) and classical molecular mechanics (MM) simulations
provide the two complimentary ways to address microscopic structure
and energetics of solvated uranyl species. Till now, the most popular
strategy to combine the accuracy of QM methods and the complexity
of the statistical solvent description is to apply a polarizable continuum
model (PCM) to represent the effective electrostatic field of the
solvent, usually employed in combination with phenomenological corrections
for dispersion-repulsion interactions and cavity formation energy
using empiric parameters \cite{Tomasi05,Fuchs02}. The state of the
art PCM is owing its success in part to the conductor-like screening
model (COSMO) of electrostatic interactions \cite{Klamt95} and has
successfully been applied in various studies of uranyl complexes \cite{Fuchs02,Moskaleva04,Cao05,Shamov05,Gutowski06}.
On the other hand, there is a significant effort to improve the empirical
model potentials for MM simulations, such as molecular dynamics (MD)
or Monte-Carlo (MC) methods, that yield the most explicit statistical
solvent description \cite{Guilbaud93,Guilbaud96,Rai12,Pomogaev13,Kerisit13,Tiwari14}.
Though several studies \cite{Gordon00,hagberg05,Frick09,Tirler13}
attempt to combine QM and stochastic MM models such combinations will
inevitably lack the precise reproducibility  and efficiency of analytic
continuum models due to statistical sampling techniques. Yet more
demanding QM dynamic simulations \cite{Buehl05,Buehl06,Buehl11,Nichols08}
are restricted to comparably small numbers of solvent molecules.

Meanwhile, the integral equation theories of molecular liquids \cite{book:Hirata03,book:Hansen06},
may provide sufficient accuracy at an affordable cost for simulating
chemical processes in solution by employing standard numerical methods
on contemporary hardware. Two approaches of this type are the reference
interaction site model (RISM) \cite{Chandler72,Hirata81,Hirata82,Kinoshita97}
and its three-dimensional generalization (3D RISM) \cite{Beglov95,Kovalenko98}
which provide the solvent structure in form of site-site radial distribution
functions (RDFs) or three-dimensional distribution functions of solvent
sites in an external (solute) field, respectively. Both 1D and 3D
theories provide variational expressions for the solvation energy
in the form of the excess chemical potential for several closure types
\cite{Singer85,Kast08} and, hence, intuitive expressions for the
first derivatives. The dielectrically consistent RISM (DRISM) \cite{Perkyns92}
was designed to compensate the error of the gas-like dielectric response
of RISM for electrolytes and ionic solutes \cite{Perkyns94,Chiodo07,Chuev08,Joung13}.

A correct handling of the electrostatic field of a charge distribution
requires additional care when periodic boundary conditions are employed
to simulate finite systems \cite{Makov95}. It was understood early
that the long-range asymptotes of the radial correlation functions
induced by Coulomb interactions in RISM need a separate treatment
as well \cite{Ng74}. For 3D RISM the few practical approaches include
the background charge correction technique \cite{Kovalenko00b}, monopole
correction \cite{Kloss08b}, and an exact re-summation for a superposition
of gaussian charges \cite{Perkyns10}. In this work we present a generalization
of the exact re-summation technique for arbitrary solute charge densities.

The integral equation theory has been combined with Hartree--Fock
electronic structure theory for self-consistent field calculations
(RISM SCF) in the pioneering work of Ten-no, Hirata, and Kato and
has been first applied to formaldehyde and carbonyl compounds in water
\cite{Tenno93,Tenno94}. An extension to 3D RISM SCF was developed
for density functional theory (DFT) \cite{Kovalenko99} and \emph{ab
initio} calculations \cite{Kloss08}. The analytical gradients for
the self-consistent combination of 3D RISM and DFT \cite{Gusarov06}
simplify explorations of free energy landscapes. Some of the recent
developments of hybrid approaches combining integral equation theories
and quantum chemistry are discussed in the recent review by Sato \cite{Sato13}.

In this work, we first describe the implementation of the 3D RISM
approach and a numerical treatment of the long-range Coulomb field
of the charged solute, Section~\ref{sec:methodology}, and then validate
the implementation by comparing the results of 1D and 3D techniques
for alkali and halide ions, Section~\ref{sec:simple-ions}. We further
evaluate the performance of the method for the uranyl(VI) dication
in aqueous solution represented by aqua complexes $[\text{UO}_{2}(\text{H}_{2}\text{O})_{n}]^{2+}$,
($n=4-6$), using a force field, Section~\ref{sub:MM+RISM}, and
a quantum mechanical representation of interactions within the first
solvation shell, Section~\ref{sub:QM+RISM}. We also compare the
results to those of PCM calculations. The activation barrier of water
exchange in the first solvation shell is analyzed in Section~\ref{sub:water-exchange}
using QM and MM models of interactions in the first solvation shell.

\section{Methodology \label{sec:methodology}}

\subsection{3D RISM equations}

In the 3D RISM approach, the solvent sites are exposed to an external
field of a solute of general shape. The solute-solvent interactions
are thus assumed to be pairwise, which in this formulation excludes
\emph{polarizable} solutes. Further, the interactions are usually
approximated by superposition of pairwise site-site contributions.
The 3D RISM equations operating with site-specific distributions can
be written in the following form \cite{book:Hirata03}:

\begin{equation}
\begin{cases}
h=\chi\ast c\\
h=c+t\\
h=f(-v+t)
\end{cases}\label{eq:rsim-h-c-t}
\end{equation}
Here $h$, $c$, and $t$ are the vectors of total-, direct-, and
indirect correlation functions $h_{\alpha}(\vec{r})$, $c_{\alpha}(\vec{r})$,
and $t_{\alpha}(\vec{r})$. All are dimensionless unknowns of the
equations that will be represented on a uniform grid of points. The
greek subscripts, such as $\alpha$, enumerate solvent interaction
sites.

The solvent susceptibility $\chi$ in the first of the three Eqs.~\ref{eq:rsim-h-c-t}
is represented by its (square symmetric) matrix $\chi_{\alpha\gamma}(r)$
derived from the radial site-site correlation functions of the pure
solvent in 1D RISM equations for a specific temperature and solvent
number density $\rho$. \cite{book:Hirata03}. The star denotes the
convolution integral and the sum over corresponding matrix indices.
For that reason it is convenient to represent the solvent susceptibility
by its (dimensionless) Fourier image $\tilde{\chi}_{\alpha\gamma}(k)$
\cite{Chandler72,book:Hirata03}. The solvent susceptibility as a
function of either $r$ or $k$ is spherically symmetric by construction.

The second of the Eqs.~\ref{eq:rsim-h-c-t} introduces the indirect
correlation $t$ that is conveniently used in the closure relation
--- the third and last of Eqs.~\ref{eq:rsim-h-c-t}. Closure is a
local relation between the values of site potential $v_{\alpha}(\vec{r})$
and, in this case, the indirect correlation $t_{\alpha}(\vec{r})$
at each point $\vec{r}$ and the value of the direct correlation $h_{\alpha}(\vec{r})$
at this point. It is the only non-linear relation between the unknowns
$h$, $c$ and $t$ in Eqs.~\ref{eq:rsim-h-c-t}. The site-specific
potentials $v$ are measured in units of temperature and thus are
also dimensionless. For the purposes of this work we have chosen a
very specific, though popular, form of the closure which happens to
be a relation between $h$ and the sum $x=-v+t$ sometimes referred
to as a renormalized indirect correlation \cite{Kast08}:
\begin{equation}
f(x)=\begin{cases}
\mathrm{e}^{x}-1 & x\le0\\
\sum_{k=1}^{n}x^{k}/k! & x>0
\end{cases}\label{eq:closure-pse}
\end{equation}
In the limit of $n\to\infty$ one recovers the hypernetted chain (HNC)
closure. Another popular closure is obtained for $n=1$. It was proposed
by Kovalenko and Hirata (KH) for improving the numerical stability
of the equations \cite{Kovalenko99}. Higher order closures with $n>1$
are referred to as partial series expansion closures of order $n$
(PSE$n$). For consistency, we will also refer to the KH closure as
PSE1 in this work.

We used pair-wise interactions between solute and solvent sites $i$
and $\alpha$ to define the solute field: $v_{\alpha}(\vec{r})=\beta\sum_{i}u_{i\alpha}(|\vec{r}-\vec{r}_{i}|)$
where $\beta$ is the inverse temperature. Each potential $u_{i\alpha}$
is defined as the sum of the Coulomb interaction $u_{i\alpha}^{\text{C}}(r)=q_{i}q_{\alpha}a[f_{\text{CS}}(ar)+f_{\text{CL}}(ar)]$,
where we separated the singular short-range $f_{\text{CS}}(r)=\mathrm{erfc}(r)/r$
from the regular long-range $f_{\text{CL}}(r)=\mathrm{erf}(r)/r$
part, and a Lennard--Jones (LJ) type potential, $u_{i\alpha}^{\text{LJ}}(r)=\epsilon_{i\alpha}f_{\text{LJ}}(r/\sigma_{i\alpha})$
with a LJ form $f_{\text{LJ}}(r)=4(r^{-12}-r^{-6})$. Parameters of
the pair interactions were derived from site parameters using the
Lorentz--Berthelot rules \cite{book:Hirata03}. The Ewald range parameter
$a=1.2$~Å$^{-1}$ is common for all site pairs. Further, to circumvent
numerical problems with singularities the primitive forms $f_{\text{CS}}(r)$
and $f_{\text{LJ}}(r)$ were regularized by substituting them with
quadratic functions $\bar{f}(r)=f(b)+f^{\prime}(b)(r^{2}-b^{2})/2b$
for $r<b$. For our systems at room temperature the results were almost
indistinguishable for finite $b\le0.5$ so that we settled with $b=0.2$
in this work.

\subsection{Free-energy surface}

The analytical expression for the excess chemical potential $\mu$
of an unstructured solute and the HNC closure has been derived by
Singer and Chandler (here in temperature units) \cite{Singer85}:
\begin{equation}
\mu_{\text{HNC}}=\rho\int d^{3}r\sum_{\alpha}\left[h_{\alpha}^{2}/2-c_{\alpha}-c_{\alpha}h_{\alpha}/2\right].\label{eq:chem-pot-hnc}
\end{equation}
A modification for the KH closure \cite{Kovalenko99} and a generalization
for the arbitrary closure of the form $h=f(x)$ with $x=-v+t$ are
available: 
\begin{equation}
\mu=\mu_{\text{HNC}}+\rho\int\sum_{\alpha}\phi(x_{\alpha})d^{3}r
\end{equation}
with $\phi(x)=f(x)-x-\int_{0}^{x}f(y)dy$ having particularly a simple
form for the PSE$n$ closures: $\phi_{n}(x)=-x^{n+1}/(n+1)!$ for
$x>0$ and zero otherwise \cite{Kast08}. In the case of the KH (PSE1)
closure the additional term $-x^{2}/2$ cancels the $h^{2}/2$ term
of the HNC integrand, Eq.~\ref{eq:chem-pot-hnc}, in the regions
where $x=h>0$.

The analytical expression for the excess chemical potential is a path-independent
integral of the work required to ``switch on'' the external (solute)
field $v$ \cite{Singer85,Kast08}. This property leads to straightforward
expressions for derivatives of the chemical potential with respect
to parameters affecting the external field, such as the location of
solute sites. Such a derivative is an expectation value of the corresponding
force, $d\mu/ds=\int(dv/ds)gd^{3}r$, with $g=h+1$.

The free energy surface as a function of solute parameters $s$ is
constructed as a composite functional
\begin{equation}
G[n_{\text{e}},s]=E[n_{\text{e}},s]+\mu(s),
\end{equation}
where $E[n_{\text{e}},s]$ is the solute internal energy depending
on the solutes' geometric degrees of freedom $s$ and, in the case
of a QM solute, also on the electronic density $n_{\text{e}}$, but
not on the solvent site distribution functions $h$ \cite{Bo15}.

The uranyl solvation energy is then estimated as\textcolor{green}{{}
}
\begin{eqnarray}
\Delta G_{\text{sol}} & = & G[\text{UO}_{2}(\text{H}_{2}\text{O})_{n}^{2+}]-nG[\text{H}_{2}\text{O}]-E[\text{UO}_{2}^{2+}]\label{eq:dG-sol-definition}
\end{eqnarray}
for $n\ge0$ explicit aqua ligands. Note that without proper statistical
sampling over the solute degrees of freedom $s$ this approximation
of the solvation energy, justified to some extent for sufficiently
rigid solutes, breaks for increasingly many explicit water molecules
outside of the tightly bound first solvation shell as $n\to\infty$.
Indeed, in a gedankenexperiment the difference $G[(\text{H}_{2}\text{O})_{n}]-nG[\text{H}_{2}\text{O}]$
for the above definition of $G$ applied to water droplets asymptotically
approaches $n(\bar{E}[\text{H}_{2}\text{O}]-G[\text{H}_{2}\text{O}])$
where $\bar{E}$ is the average configurational energy of a water
molecule which is about $-$9.9~kcal/mol \cite{Berendsen87,Mahoney00}
to be compared to the solvation energy of $-$6.3~kcal/mol derived
from the ratio of the equilibrium vapor/liquid concentrations \cite{Bridgeman64}.

\subsection{Numerical solution of 3D RISM equations}

Because of the highly non-linear closure relation, a single variable
fix point iteration scheme, $t^{(n+1)}=T[t^{(n)}]$, with $T[t]=(\chi-1)\ast c(t)$
and $c(t)=f(-v+t)-t$ as derived from the non-linear equation system,
Eqs.~\ref{eq:rsim-h-c-t}, is bound to diverge for all but the simplest
cases. To overcome the divergence problem we solve the reduced non-linear
problem $F[t]=T[t]-t=0$ using the Newton--Krylov non-linear solver
\cite{petsc-user-ref} with the analytic matrix-free Jacobian, $J[t]=\delta F/\delta t$,
occasionally preceded by a few Picard iterations.

The long-range field of a charged solute requires a special treatment
in the 3D case. All of the practical approaches, including the Ng
scheme \cite{Ng74}, separate the smooth asymptote $v_{\text{L}}$
of the site specific potential off the indirect correlation $t$ as
well as $-v_{\text{L}}$ off the direct correlation $c$ to operate
primarily with the short range counterparts $v_{\text{S}}$ , $c_{\text{S}}$,
and $t_{\text{S}}$. An iteration scheme for the new primary variable
$t_{\text{S}}$ (and the resulting non-linear problem) will look slightly
different, though:

\begin{eqnarray}
T_{\text{S}}[t_{\text{S}}] & = & (\chi-1)\ast c_{\text{S}}(t_{\text{S}})-\tau
\end{eqnarray}
with $c_{\text{S}}(t_{\text{S}})=f(-v_{\text{S}}+t_{\text{S}})-t_{S}$
and the only term that involves a long-range operand being the constant
``renormalization'' term $\tau=\chi\ast v_{\text{L}}$. This term
is the convolution of solvent susceptibility with the asymptotic potential.

The low $k$ behavior of the asymptotic potential $v_{\text{L}}$
originates from the Coulomb potential and is, in general, singular.
With our choice of separation into short- and long-range parts it
reads 
\begin{equation}
\tilde{v}_{\text{L}}(\vec{k})=q\tilde{\varphi}(k)\tilde{n}(\vec{k})\sim4\pi qQ/k^{2}
\end{equation}
with $\tilde{n}(\vec{k})$ being the solute electric form factor,
$Q=\tilde{n}(0)$ being the net solute charge, $q$ being the vector
of the solvent site charges $q_{\alpha}$ and $\tilde{\varphi}(k)=4\pi k^{-2}\exp(-k^{2}/4a^{2})$
the Fourier transform of the long range potential function $\varphi(r)=\mathrm{erf}(ar)/r$.
A straightforward representation of this convolution operand on a
3D grid is thus problematic and should be avoided. The result of the
convolution, $\tau$, is of short range, though.

Indeed, the low $k$ behavior of another convolution operand, the
solvent susceptibility is regular 
\begin{equation}
\tilde{\chi}_{\alpha\gamma}(k)\sim\rho\kappa+\frac{1}{2}d_{\alpha\gamma}k^{2}
\end{equation}
with the limit for $k\to0$ being proportional to the solvent compressibility
$\kappa$ (here in temperature units) \cite{book:Hansen06}. More
importantly, this limit is the same for all site pairs $\alpha$ and
$\gamma$. Though not enforced, this property is obeyed in practical
calculations with reasonable accuracy. The second, $o(k^{2})$, term
is site pair specific and is in particular responsible for the dielectric
response of the solvent --- it is this term that is addressed by the
dielectric correction of the DRISM approach \cite{Perkyns92,Cummings81,Kovalenko00b}.
The result of the convolution, 
\begin{equation}
\tilde{\tau}_{\alpha}(\vec{k})=\tilde{n}(\vec{k})\left[\tilde{\varphi}(k)\sum_{\gamma}\tilde{\chi}_{\alpha\gamma}(k)q_{\gamma}\right]\sim2\pi Q\sum_{\gamma}d_{\alpha\gamma}q_{\gamma}\label{eq:tau}
\end{equation}
is regular assuming net neutrality of the solvent, $\sum_{\gamma}q_{\gamma}=0$,
which makes the compressibility term vanish. Note that the terms in
square brackets, $\tilde{\varphi}(k)\sum_{\gamma}\tilde{\chi}_{\alpha\gamma}(k)q_{\gamma}=:\tilde{\theta}_{\alpha}(k)$
(one per solvent site $\alpha$), are convolutions of spherically
symmetric functions neither of which is solute specific. The functions
$\tilde{\theta}_{\alpha}(k)$ are site-specific solvent properties
representing response to a unit (here gaussian-broadened) charge that
should be computed using 1D techniques, just as the solvent susceptibility
$\tilde{\chi}_{\alpha\gamma}(k)$ itself. In the 1D case our deliberate
choice of the self-inverse discrete sine transform of type IV (DST-IV)
as a basic operation eliminates the need to treat the singularities
at $r=0$ and $k=0$ as the discrete representation do not include
these points by construction \cite{FFTW05}. The solute part is the
electric form factor $\tilde{n}(\vec{k})$ of an arbitrary solute
charge distribution $n(\vec{r})$ with a low $k$ expansion including
all multipoles of the charge distribution; it is treated by regular
3D techniques.

A similar approach was applied before to a monopole field of the solute
\cite{Kloss08b} and to a superposition of gaussian charge distributions
\cite{Perkyns10}.  The argument line used above assumes the net
neutrality of the solvent species and is not directly applicable to
electrolytes, though even then the low $k$ behavior of a convolution
like in Eq.~\ref{eq:tau} can be assumed to be regular, cp.\ the
discussion of the background charge correction in Ref.~\cite{Kovalenko00b}.
An interpretation of Eq.~\ref{eq:tau} as a linear response of the
pure solvent distribution to a (weak) electrostatic field of the charge
distribution $n(\vec{r})$ offers an intuitive way to judge the range
assumptions for most solvent-solute combinations, cf.~Eqs.~\ref{eq:rsim-h-c-t}
for sound closures that fulfill $f(x)\sim x$ upon linearization.

\section{Computational details \label{sec:computational-details}}

To model uranyl solvation we employ the SPC/E water model \cite{Berendsen87},
modified as suggested by Pettitt and Rossky \cite{Pettitt82} by addition
of a Lennard--Jones (LJ) term for hydrogen atoms with $\sigma_{\text{H}}=0.4$~Å
and $\epsilon_{\text{H}}=0.046$~kcal/mol (PR-SPC/E). For atomic
ions we adopted the coincident SPC/E water model (cSPC/E) with $\sigma_{\text{H}}=1.1658$~Å
and $\epsilon_{\text{H}}=0.01553$~kcal/mol as proposed by Luchko
and Kovalenko \cite{Luchko10} in order to compare with published
results. The uranyl-water interaction parameters were derived from
the second model proposed by Kerisit and Liu to strengthen the uranyl\textendash{}water
interaction with the uranium charge of 3.5~\emph{e} (KL2) \cite{Kerisit13}.
For convenience the reader may refer to our previous work \cite{Bo15}
where we summarized the literal values for the charge and LJ parameters
of several force-fields.

All free energy calculations were carried out for the temperature
of 298~K, with the dielectric constant set to 78.4 and the solvent
particle density $\rho=0.0333295$~Å$^{-3}$. These parameters correspond
to a pressure of 1~atm. The 3D domain was represented by a cubic
cell of $20^{3}$~Å$^{3}$ and a Cartesian grid with 96$^{3}$ points
and a resolution of about 0.208~Å. The radial grid used to precompute
and interpolate the solvent susceptibility and its contractions with
the Coulomb kernel for use in Eq.~\ref{eq:tau} was chosen with a
radial range of 40~Å and 1536 points corresponding to a resolution
of about 0.026~Å.

All-electron QM calculations were carried out with the linear combination
of Gaussian-type orbitals fitting-functions density functional method
\cite{Dunlap90} as implemented in the parallel code ParaGauss \cite{ParaGaussV3.1}.
The atomic approximation \cite{Matveev08} to the second-order Douglas\textendash{}Kroll\textendash{}Hess
scalar-relativistic approach \cite{rev:roesch96} was used to account
for relativistic effects; this approximation was shown to be very
accurate for uranyl aqua complexes. We employed the gradient-corrected
exchange-correlation functional (generalized gradient approximation,
GGA) \cite{Goerling99,kochDFT}, as suggested by Becke and Perdew
(BP) \cite{Becke88,Perdew86c}. The Kohn\textendash{}Sham orbitals
were represented by flexible Gaussian-type basis sets, contracted
in a generalized fashion using atomic eigenvectors. For U, we used
a basis set of the size (24s, 19p, 16d, 11f), contracted to {[}10s,
7p, 7d, 4f{]}; O and H atoms were described by standard basis sets,
(9s, 5p, 1d) $\to$ {[}5s, 4p, 1d{]} and (6s, 1p) $\to$ {[}4s, 1p{]},
respectively. These basis sets and further computational details are
the same as in earlier studies on uranyl species \cite{Moskaleva04,Moskaleva06b,Moskaleva06a,Kremleva12b,Bo15}.

We used a COSMO PCM solvation model for comparison \cite{Fuchs02},
with the cavity defined via an effective solvent radius of 1.4~Å
and van der Waals radii of solute atoms as tabulated by Bondi \cite{Bondi64}
and scaled by 1.125 (except for hydrogen). We carried out the cavity
tessellation with the FIXPVA approach \cite{Su09,Kremleva12b} (fixed
points with variable areas) to obtain numerically stable energies
and forces \cite{Kremleva12b}.

Geometry optimizations have been carried out using the quasi-Newton
method and the constrained minimization algorithm as implemented in
the utility package ParaTools \cite{Chaffey-Millar11,Nikodem13}.
In QM+RISM calculations of symmetric uranyl aqua complexes the U\textendash{}O
bonds were relaxed but constrained to have the same value by symmetry.
In the corresponding MM+RISM calculations uranyl was rigid with the
U\textendash{}O distances fixed to 179~pm motivated by earlier results
\cite{Moskaleva04,Cao05,Buehl05}. However, for a study of the water
exchange with MM+RISM we adopted a flexible uranyl model without symmetry
constraints and compared two sets of intra-molecular parameters \cite{Kerisit13,Tiwari14}
with unconstrained QM+RISM results. The two uranyl force field models
we employed differ only by uranyl equilibrium bond length and force
constants for stretching and bending modes. The SPC/E geometry of
explicit water ligands in the aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$,
$n=4-6$, features an O\textendash{}H distance of 100~pm and an H\textendash{}O\textendash{}H
angle of 109.47$^{\circ}$.

\section{Simple ions \label{sec:simple-ions}}

To verify our 3D RISM implementation for charged solutes, we performed
a comparison to 1D RISM results for mono-atomic alkali and halide
ions \cite{Joung13,Bo15} applying the cSPC/E water model \cite{Berendsen87,Luchko10}.
With moderate settings  for the grid resolution of $64^{3}$ points
in total and a cubic box volume of $20^{3}$~Å$^{3}$ we are able
to match the 1D results up to 0.3~kcal/mol or better. The accuracy
is slightly better for small cations than for the larger anions. The
values are comparable to MD with SPC/E water and experimental results.
The agreement may be slightly improved by a correction proportional
to the molar volume of the solute \cite{Sergiievskyi14} as was demonstrated
in our previous work \cite{Bo15}.

\begin{table}
\begin{centering}
\caption{The solvation energy (excess chemical potential) for alkali and halide
ions by 3D and 1D RISM approaches with the cSPC/E water model. MD
and experimental values are also shown for comparison.$^{a}$ \label{tab:simple-ions}}

\par\end{centering}

\begin{ruledtabular}

\begin{tabular}{lrrrrrr}
 & Li$^{+}$ & Na$^{+}$ & K$^{+}$ & Cl$^{-}$ & Br$^{-}$ & I$^{-}$\tabularnewline
\hline 
3D KH & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.4 & $-$72.5 & $-$62.9\tabularnewline
1D KH$^{b}$ & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.5 & $-$72.6 & $-$63.1\tabularnewline
3D HNC & $-$111.8 & $-$86.3 & $-$69.3 & $-$79.8 & $-$74.1 & $-$64.7\tabularnewline
1D HNC$^{b}$ & $-$111.7 & $-$86.3 & $-$69.3 & $-$79.9 & $-$74.2 & $-$64.9\tabularnewline
MD$^{c}$ & $-$113.3 & $-$88.4 & $-$71.0 & $-$89.3 & $-$82.7 & $-$74.4\tabularnewline
Exp$^{d}$ & $-$113.8 & $-$88.7 & $-$71.2 & $-$89.1 & $-$82.7 & $-$74.3\tabularnewline
\end{tabular}

\end{ruledtabular}

\footnotetext{$^{a}$~Energies in kcal/mol. KH and HNC closure relations
have been used. 3D domain is $20^{3}$ Å$^{3}$ / $64^{3}$ pts. $^{b}$~Ref.~\cite{Bo15}.
$^{c}$~MD with SPC/E water from Ref.~\cite{Joung08}. $^{d}$~Ref.~\cite{Schmid00}.} 
\end{table}

\section{uranyl aqua complexes \label{sec:uranyl-aqua-complexes}}

\subsection{MM+RISM \label{sub:MM+RISM}}

In this section we examine a combination of 3D RISM for bulk water
and a force field potential energy surface for the interactions of
uranyl and $n$ explicit water ligands for $n=4-6$. The complex with
$n=4$ features a missing equatorial ligand and demonstrates a qualitative
difference of the optimal structures obtained by minimizing the MM
and and MM+RISM expressions of the potential (free) energy. Similar
to the findings in our previous work, using the 1D RISM formalism
\cite{Bo15}, the MM+RISM combination features a vacancy in the equatorial
plane of the uranyl corresponding to the location of the missing water
ligand, Figure~\ref{fig:spatial-distribution-4w}. The MM potential
energy surface features a minimum at a $D_{4\text{h}}$-like structure
with four water ligands evenly distributed in the equatorial plane;
structures featuring a vacancy in the equatorial plane are by 13--15~kcal/mol
higher on that potential energy surface. The average U--OW bond in
the MM-optimized structure of 239~pm is slightly shorter than the
corresponding bonds of 240--242~pm of the structures optimized at
the MM+RISM level, Table~\ref{tab:mm-rism}. The shorter distances
indicate weaker steric ligand-ligand repulsions in the under-coordinated
gas-phase complex. Note, however, there is a non-obvious opposite
effect of shortening U--OW distances from 243--244~pm to 240~pm
in five-coordinated uranyl complexes after embedding a gas phase complex
into the RISM medium, Table~\ref{tab:mm-rism}. This could be an
indication of a partial screening of steric repulsions of the equatorial
ligands by the solvent of the second solvation shell in saturated
complexes, also observed in PCM calculations, see below.

The mechanism by which the MM+RISM model compensates for the missing
ligand is a local increase of the water density at the location of
the missing explicit ligand, Figure~\ref{fig:spatial-distribution-4w}.
Due to a notable width of the local distribution, both in equatorial
and apical directions and less so in radial direction the RISM representation
of the missing ligand is not an exact substitute for an explicit water
ligand. Indeed, there is a measurable difference of the average U--OW
distances between the complexes with 4 and 5 explicit waters in the
first solvation shell: 241--242~pm for $n=4$ and 240~pm for $n=5$,
Table~\ref{tab:mm-rism}.

The strong localization of the solvent medium at the equatorial vacancy
leads to divergence of the RISM equations with the PSE3 closure. This
is manifested by the missing values in Table~\ref{tab:mm-rism}.
Results for the five-coordinated uranyl models, on the other hand,
are essentially independent of the closure choice. The bare uranyl
model, where the axial symmetry makes the localization issue less
severe, does converge for all three closures, though the results are
significantly affected by the closure choice --- the solvation energy
at the PSE3 level of theory is by 20~kcal/mol lower than at the PSE1
level. A different source of errors is the approximation of the solute
by a single structure minimizing the conformation free energy instead
of an ensemble at thermal equilibrium. Whereas this could be justified
to some extent for solutes with rigid intra-molecular and strong ligand
bonds, the weakly associated water molecules of the second solvation
shell need a better statistical treatment. The model with six explicit
waters with one of them in the second shell ($n=5+1$, Table~\ref{tab:mm-rism})
illustrates the point. Whereas the structure of the first solvation
shell is very similar for $n=5$ and $n=5+1$ explicit waters, the
solvation energy, computed by Eq.~\ref{eq:dG-sol-definition} is
by about 5~kcal lower in the latter case. This is likely an artifact
of a missing statistical sum over positions and orientations of the
second shell water molecule. Hence a successful model shall treat
the strongly bound aqua ligands explicitly as a part of the solute
and weakly associated or free-floating water as a RISM medium in order
to be numerically stable and closure independent.

\begin{table}
\caption{Average uranyl--water distance of the first solvation shell, $d$(U--OW),
and an estimate for the uranyl solvation energy, $\Delta G_{\text{sol}}$,
obtained using gas phase MM (GP) and MM+RISM models with PSE closures
of order $n=1-3$ (PSE$n$), see Eq.~\ref{eq:closure-pse}.$^{a}$
\label{tab:mm-rism}}

\begin{ruledtabular}

\begin{tabular}{llrrrr}
$n$ &  & 0 & 5$-$1$^{b}$ & 5 & 5+1$^{c}$\tabularnewline
\hline 
$d$(U--OW) & GP & --- & 239 & 244 & 243\tabularnewline
 & PSE1 & --- & 241 & 240 & 240\tabularnewline
 & PSE2 & --- & 242 & 240 & 240\tabularnewline
 & PSE3 & --- & na$^{d}$ & 240 & 240\tabularnewline
$\Delta G_{\text{sol}}$ & GP$^{e}$ & --- & $-$233 & $-$272 & $-$295\tabularnewline
 & PSE1 & $-$363 & $-$382 & $-$392 & $-$397\tabularnewline
 & PSE2 & $-$379 & $-$389 & $-$392 & $-$397\tabularnewline
 & PSE3 & $-$383 & na$^{d}$ & $-$392 & $-$397\tabularnewline
\end{tabular}

\end{ruledtabular}

\footnotetext{ $^{a}$~Bond lengths in pm, energies in kcal/mol.
$^{b}$~GP structure is $D_{4\text{h}}$ but the RISM structures
feature a vacancy in the equatorial plane corresponding to the missing
ligand, see Figure~\ref{fig:spatial-distribution-4w}. $^{c}$~Five-coordinated
structure with the sixth water in the second shell. $^{d}$~Value
not available due to divergence. $^{e}$~The GP row shows binding
energies of $n$ water molecules to a rigid uranyl with $d$(U--O)
of 179~pm. }
\end{table}
\begin{figure}
\caption{Water oxygen distribution around the uranyl aqua complex with four
explicit water ligands. The isosurface corresponds to $h=2$. The
area around the missing equatorial ligand is characterized by values
of $h$ more than ten times as high. Other lopes of the isosurface
correspond to second solvation shell and feature $h<6$. \label{fig:spatial-distribution-4w}}

\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/4w}
\end{figure}

\subsection{QM+RISM \label{sub:QM+RISM}}

We use symmetric models of uranyl aqua complexes to compare PCM and
RISM representations of the solvent in QM calculations. The under-coordinated
species does not feature a vacancy in the first solvation shell. Instead
all water ligands are equivalent due to symmetry, also for the five-
and six-coordinated uranyl. These models feature an additional intra-molecular
degree of freedom, the uranyl bond length, in contrast to the MM models
discussed in the previous section where the uranyl bond was fixed
to a reasonable value.

In our previous work exploring these models with 1D RISM \cite{Bo15}
we observed consistently longer U--O and U--OW bond lengths in QM+RISM
calculations in comparison to QM+PCM and experiment. We tentatively
attributed that to the excess coordination of uranyl, manifested by
the presence of a peak in the U--OW radial distribution at the distance
corresponding to the first solvation shell for models with explicit
ligands. This artifact is specific to 1D RISM and is not reproduced
in 3D RISM calculations, cp.\ Figure~\ref{fig:rdfs-vs-closure-5w}
showing no features at that separation and Figure~\ref{fig:rdfs-vs-closure-0w}.
The number integrals of 5.3, 6.2, and 6.4 for the first peak of U--OW
RDFs with PSE closures of order 1 to 3, respectively, for uranyl without
explicit aqua ligands (Figure~\ref{fig:rdfs-vs-closure-0w}) are
overestimated though.

Apart from the bare uranyl model the QM+RISM and QM+PCM results for
uranyl bond lengths of 179--180~pm agree in Table~\ref{tab:qm+rism-vs-qm+pcm}.
For the bare uranyl model the QM+RISM result is qualitatively better.
The QM+PCM predicts 174~pm for the uranyl bond length which is shorter
than for uranyl with just four water ligands. The QM+RISM results,
on the other hand, are around 179~pm being equally long (PSE1) or
even marginally longer by 0.3~pm and 0.7~pm for (PSE2 and PSE3,
respectively) than for four-coordinated uranyl (not visible in Table~\ref{tab:qm+rism-vs-qm+pcm}
due to rounding).  The effect of the RISM medium on the bond length
of the bare uranyl is thus comparable to that of the explicit water
ligands; the elongation of the uranyl bond by 7--8~pm corresponds
to 5--6~kcal/mol change in internal energy.

Another notable discrepancy between the PCM and 1D RISM results mentioned
in our previous work is also resolved by the 3D RISM approach: the
uranyl--water distances U--OW are consistently shorter in solvated
complexes than in the gas phase by measurable 4--6~pm. The 1D RISM
approach did not show this effect convincingly \cite{Bo15}. Again,
we interpret this effect of the medium as a screening of repulsive
ligand-ligand interactions in the first coordination shell by the
solvent.

The magnitude of the uranyl solvation energy estimated with QM+RISM
models is comparable with that obtained in our previous work using
1D RISM. The estimates of $-$388 and $-$389~kcal/mol using the
standard model of hydrated uranyl with five explicit water ligands
vary only little with the closure order and agree well with $-$390~kcal/mol
from our previous work \cite{Bo15}. However, the relative order of
the estimates based on the four- and six-coordinated complexes is
not reproduced. The estimate of $\Delta G_{\text{sol}}$ based on
the four-coordinated complex is by $\sim$10~kcal/mol less negative
than the result of the standard model, Table~\ref{tab:qm+rism-vs-qm+pcm}.
This difference (but not the absolute numbers) is the same as the
PCM result, Table~\ref{tab:qm+rism-vs-qm+pcm}. However, this difference
is significantly larger than the value of 1.6~kcal/mol obtained using
1D RISM \cite{Bo15}, which is remarkably close to the free energy
difference corresponding to the relative weights of four- and five-coordinated
species determined from an average coordination number in a high-energy
x-ray scattering (HEXS) experiment \cite{Soderholm05}. Given the
limitations of 1D RISM exposed by the current 3D RISM approach we
interpret that past agreement as fortuitous.  $\Delta G_{\text{sol}}$
calculated from the six-coordinated $D_{3\text{d}}$ complex is by
4--5~kcal/mol more negative than the result of the standard model
with five explicit water ligands, Table~\ref{tab:qm+rism-vs-qm+pcm}.
The corresponding PCM difference is only 1~kcal/mol and of opposite
sign. It is likely that the limitations of the Eq.~\ref{eq:dG-sol-definition}
with the free energy of the complex approximated from a single symmetry-constrained
conformation is the source of this counter-intuitive ordering of estimates
for $\Delta G_{\text{sol}}$. For similar reasons the MM+RISM estimates
of the uranyl solvation energy using the models with different number
of explicit water molecules agree between each other no better than
to 3--5~kcal/mol even at the highest closure level, Table~\ref{tab:mm-rism}.
 Again, the solvation energy as determined by the QM+PCM model without
explicit water ligands ($n=0$) is by $\sim$100~kcal/mol off the
results obtained with $n=4-6$ explicit ligands, Table~\ref{tab:qm+rism-vs-qm+pcm}.
 The QM+PCM solvation energy of $-426$~kcal/mol given by the standard
five-coordinated uranyl model appears to be too low \cite{Moskaleva04}
in the view of most recent experimental estimate of $-369\pm15$~kcal/mol
\cite{Gibson05}. The corresponding QM+RISM solvation energy of $-389$~kcal/mol
is significantly closer to that interval. The properties in Table~\ref{tab:qm+rism-vs-qm+pcm}
obtained using the models with explicit water ligands separating the
uranyl di-cation and the RISM medium ($n=4-6$) hardly depend on the
closure order. This observation may justify the widespread use of
the lowest order PSE1 (KH) closure.

\begin{table}
\caption{Uranyl bond length, $d$(U--O), uranyl--water distance, $d$(U--OW),
and the uranyl solvation energy estimate, $\Delta G_{\text{sol}}$,
obtained using gas phase QM, QM+PCM and QM+RISM models.$^{a}$ \label{tab:qm+rism-vs-qm+pcm}}

\begin{ruledtabular}

\begin{tabular}{llrrrr}
$n$ &  & 0 & 4$^{b}$ & 5$^{c}$ & 6$^{d}$\tabularnewline
\hline 
$d$(U--O) & GP & 172 & 177 & 177 & 178\tabularnewline
 & PCM & 174 & 179 & 179 & 180\tabularnewline
 & PSE1 & 179 & 179 & 179 & 180\tabularnewline
 & PSE2 & 179 & 179 & 179 & 180\tabularnewline
 & PSE3 & 179 & 179 & 179 & 180\tabularnewline
$d$(U--OW) & GP & --- & 242 & 249 & 253\tabularnewline
 & PCM & --- & 237 & 243 & 249\tabularnewline
 & PSE1 & --- & 237 & 243 & 248\tabularnewline
 & PSE2 & --- & 237 & 243 & 248\tabularnewline
 & PSE3 & --- & 237 & 243 & 248\tabularnewline
$\Delta G_{\text{sol}}$ & GP$^{e}$ & --- & $-$244 & $-$270 & $-$289\tabularnewline
 & PCM & $-$322 & $-$416 & $-$426 & $-$425\tabularnewline
 & PSE1 & $-$357 & $-$377 & $-$388 & $-$393\tabularnewline
 & PSE2 & $-$373 & $-$379 & $-$389 & $-$394\tabularnewline
 & PSE3 & $-$378 & $-$379 & $-$389 & $-$393\tabularnewline
\end{tabular}

\end{ruledtabular}

\footnotetext{ $^{a}$~Bond lengths in pm, energies in kcal/mol.
Symmetric uranyl aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$,
$4\leq n\leq6$ and a bare uranyl ($n=0$) are treated by QM while
interactions with the PR-SPC/E water solvent are treated by KL2 force
field. $^{b}$~$D_{4\text{h}}$ structure. $^{c}$~$D_{5\text{h}}$
structure. $^{d}$~$D_{3\text{d}}$ structure. $^{e}$~The row shows
binding energies of $n$ water ligands. }
\end{table}

\begin{figure}
\caption{Radial distribution functions of the solvated uranyl without explicit
aqua ligands: U--OW (top panel, black), U--HW (top panel, gray), O--OW
(bottom panel, black) and O--HW (bottom panel, gray). Results employing
PSE1 closure are shown with the thick lines in the foreground. Results
employing PSE2 and PSE3 closures are displayed using thinner lines
in the background.  \label{fig:rdfs-vs-closure-0w}}

\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig2}
\end{figure}

\begin{figure}
\caption{Radial distribution functions of the solvated uranyl complex with
five explicit water ligands in the equatorial plane. See Fig.~\ref{fig:rdfs-vs-closure-0w}
for explanation. \label{fig:rdfs-vs-closure-5w}}

\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig3}
\end{figure}

\subsection{Water exchange \label{sub:water-exchange}}

By way of the Eyring equation the activation energy of aqua ligand
exchange process in the first solvation shell of uranyl computed from
the experimentally available water exchange rate estimate of 1.3$\times$10$^{6}$~s$^{-1}$
is about 9~kcal/mol \cite{Farkas00}. Many theoretical studies predict
much faster water exchange than the available experimental data suggests
\cite{Vallet01,hagberg05,Frick09,Nichols08,Kerisit10,Tiwari14}. A
simple way to obtain an estimate of the activation energy for the
water exchange may be the inspection of the U--OW potential of mean
force (PMF), $W(r)=-T\ln g(r)$, derived from the uranyl--water oxygen
RDF $g(r)$, Fig.\ \ref{fig:potential-of-mean-force}. In Table~\ref{tab:pmf-extrema}
we give the locations of the extrema of such PMF profiles, obtained
using the bare uranyl model and three different closure relations.
The activation energy estimated by 3D RISM as the PMF difference of
the first-shell minimum and the barrier top ranges from 3.4 to 5.2
kcal/mol, increasing with the closure order. The corresponding number
from the MD RDF is 4.4 kcal/mol. These values are comparable to the
5~kcal/mol derived from the residence time by Tiwari et al.~\cite{Tiwari14},
but significantly lower than the experimental estimate. The transition
between the first- and second solvation shells happens at U--OW separations
$r_{\text{b}}=309-324$~pm according to the radial PMF profiles.
Note that the one-particle radial PMF $W(r)$ provides no information
on the collective behavior of the water ligands during the water exchange,
thus it offers only limited insight into possible mechanisms. Also,
the RISM model for bare uranyl assumes a rigid linear geometry whereas
it was shown that uranyl bending essentially facilitates the water
exchange \cite{Tiwari14}.

Additional insight is provided by a three-dimensional PMF $W(\vec{r})=-T\ln g(\vec{r})$
given as $W(x,y)$ in Fig.\ \ref{fig:potential-of-mean-force-2d}
on a quarter of the axial plane with the uranium atom at the origin
and the uranyl oxygen on the $y$-axis. The location of the saddle
point  in Fig.\ \ref{fig:potential-of-mean-force-2d} corresponds
to a U--OW distance of 318~pm comparable with the respective PSE1
value in Table~\ref{tab:pmf-extrema}. The angle of attack for the
incoming/leaving water ligand estimated from that location is about
60$^{\circ}$ to the uranyl axis. This value is in good agreement
with the result of 61.8$^{\circ}$ from the transition state search
at Hartree--Fock level of theory with water solvent described by CPCM
model \cite{Vallet01}. A hypothetical water exchange in the equatorial
plane appears to be significantly hindered at room temperature. The
PMF values at the location of the saddle point and at the equatorial
minimum are 1.4~kcal/mol and $-1.7$~kcal/mol, respectively (PSE1).
Thus the activation barrier is about 3.1~kcal/mol which is only slightly
less than the estimate from the radial PMF, Table~\ref{tab:pmf-extrema}.
Again it is not possible to draw a conclusion on a collective behavior
of the incoming and leaving ligands from this one-particle property.

\begin{figure}
\caption{Potential of mean force, $W(r)$, for U--OW distribution functions
obtained with PSE1, PSE2 and PSE3 closures (solid lines of decreasing
weight). The dashed line represents the MD result. \label{fig:potential-of-mean-force}}

\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf}
\end{figure}

\begin{table}
\begin{centering}
\caption{PMF extrema $\{r_{i},W(r_{i})\}$,$^{a}$ and an estimate for the
water exchange activation energy, $\Delta W_{\text{ab}}=W(r_{\text{b}})-W(r_{\text{a}})$
extracted from the curves in Fig.\ \ref{fig:potential-of-mean-force}.
\label{tab:pmf-extrema}}

\par\end{centering}

\begin{ruledtabular}

\begin{tabular}{lrrrrrrr}
 & $r_{\text{a}}$ & $W(r_{\text{a}})$ & $r_{\text{b}}$ & $W(r_{\text{b}})$ & $r_{\text{c}}$ & $W(r_{\text{c}})$ & $\Delta W_{\text{ab}}$\tabularnewline
\hline 
PSE1 & 236 & $-$0.9 & 324 & 2.5 & 465 & $-$0.3 & 3.4\tabularnewline
PSE2 & 238 & $-$1.4 & 313 & 3.4 & 462 & $-$0.3 & 4.8\tabularnewline
PSE3 & 240 & $-$1.5 & 309 & 3.7 & 461 & $-$0.3 & 5.2\tabularnewline
MD & 246 & $-$1.5 & 297 & 2.9 & 460 & $-$0.3 & 4.4\tabularnewline
\end{tabular}

\end{ruledtabular}

\footnotetext{$^{a}$~Distances in pm, energies in kcal/mol. Locations
$r_{\text{a}}$ and $r_{\text{c}}$ correspond to the minima of first-
and second solvation shells, respectively; $r_{\text{b}}$ is the
highest point of the barrier connecting the two minima. } 
\end{table}

\begin{figure}
\caption{Potential of mean force, $W(x,y)$, for the U--OW distribution functions
obtained with PSE1 closure. Here $x$ and $y$ are the radial and
axial displacements off the uranium atom. \label{fig:potential-of-mean-force-2d}}

\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf-2d}
\end{figure}

\begin{figure}
\caption{Free energy profiles, $G(q)$, along the water exchange pathway using
two intra-molecular force fields with soft- and hard uranyl bending
modes (squares and diamonds, respectively). The QM+RISM profile is
represented by three stationary states (circles) connected by strait
lines to guide the eye. The reaction coordinate $q$ is the difference
of U--OW distances of two designated water molecules. \label{fig:free-energy-profiles}}

\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/profiles}
\end{figure}

To investigate the collective behavior of the ligands in the water
exchange process we need to treat the first solvation shell and the
incoming water molecule explicitly and also employ a flexible uranyl
model as it was shown that the uranyl bending may affect the exchange
rate \cite{Tiwari14}. The two intra-molecular force fields for a
flexible uranyl species we compare below differ primarily by the force
constant corresponding to the bending modes of uranyl and the equilibrium
bond length $d_{0}$(U--O). The force field here referred to as ``soft''
corresponds to the frequency of the bending mode of 159~cm$^{-1}$
and $d_{0}$(U--O) of 176~pm \cite{Tiwari14}. The alternative ``hard''
force field features a significantly higher frequency of the uranyl
bending mode of 393~cm$^{-1}$ and a somewhat too long a uranyl bond
 $d_{0}$(U--O) of 180~pm \cite{Kerisit13}. The bending vibration
in crystals and solution has been experimentally characterized by
frequencies in the range of 180--270~cm$^{-1}$ \cite{Denning76,Denning79,Toth81,Gal92}.
The DFT models with 5--6 explicit water ligands that we employ in
this and earlier works predict the bending frequency in this experimental
range \cite{Matveev08}. For the gas phase uranyl the standard DFT
predicts too soft a bending mode likely because of the self-interaction
error for $f$-electrons \cite{Ramakrishnan11} while hybrid DFT and
CCSD(T) estimate the bending frequency around 160--180~cm$^{-1}$
\cite{Jackson08}. In this light the force field with the ``soft''
bending mode appears to be a more reliable choice.

However, a comparison of the free energy profiles obtained with QM
and the two force field models of uranyl aqua complexes in RISM medium
reveals significant quantitative deficiencies for both force field
models. The Fig.~\ref{fig:free-energy-profiles} shows the three
free energy profiles along the reaction coordinate, $q=d(\text{U--OW}_{1})-d(\text{U--OW}_{2})$,
which is a difference of two uranium--water oxygen distances for incoming
and leaving water molecules. The QM+RISM profile is approximated by
just three stationary points. All of the three profiles feature an
intermediate metastable state at around $q=0$ roughly corresponding
to the six-coordinated $D_{3\text{d}}$-like uranyl aqua complex.
With the ``soft'' force field the uranyl species in the metastable
state is bent featuring an O--U--O angle of 155$^{\circ}$ and distances
to the two water ligands participating in exchange which are by about
15~pm longer than the distances to the other four water ligands;
the deviations from the $D_{3\text{d}}$ geometry for the other two
models, QM and ``hard'' force field, are less pronounced.  The
transition state approximated by the structure with the highest free
energy is characterized by $q>0.6$~Å for both the QM and the ``soft''
uranyl models with the longest U--OW distance of 303~pm and 310~pm,
respectively. These distances are comparable to or slightly below
the estimates derived from the 1D and 3D PMFs above. The ``hard''
uranyl model, on the other hand, suggests a transition state with
a much shorter distance to the outermost water ligand of about 280~pm
and a shallow metastable state. This is likely due to the significantly
longer U--O bond of the solvated uranyl species of about 186~pm which
is more than 6~pm longer than those of the other models employed
here and in other works \cite{Tiwari14,Kerisit13} and should be considered
a deficiency of the parametrization.  In turn, the ``soft'' model
does not convincingly show the preference for five-coordinated uranyl
with the sixth water in the second solvation shell over the six-coordinated
broken symmetry metastable structure. The QM and ``hard'' force
field models predict the six-coordinated $D_{3\text{d}}$-like state
by 5.6~kcal/mol and 3.6~kcal/mol less stable than the five-coordinated
structure with an extra water in the second shell, respectively. The
presence of such a six-coordinated intermediate suggests an associative
two-step water exchange mechanism as one of the possibilities. This
is consistent with the findings of the other theoretical studies \cite{Tsushima07,Vallet01,Buehl06b,Rotzinger07,Tiwari14}
suggesting that the associative mechanism is more favorable than the
dissociative mechanism.

The collective behavior of the uranyl aqua ligands is unspectacular
and can be described as a redistribution in the torus-shaped first
solvation shell roughly corresponding to the low PMF region, cp.~Figs.~\ref{fig:potential-of-mean-force}
and \ref{fig:potential-of-mean-force-2d}, while the sixth water molecule
enters or leaves the first solvation shell. The water exchange barrier
of 2.3 and 4.4~kcal/mol calculated with the ``soft'' and ``hard''
force field models, respectively, though not very far from the estimates
we obtained above from the PMF analysis, are significantly lower than
the 9.1~kcal/mol barrier of the QM+RISM model, Fig.~\ref{fig:free-energy-profiles},
which is in good agreement with the experimental estimate of 9~kcal/mol
\cite{Farkas00}.  A comparison of locations and orientations of
the second-shell explicit water which is connected by hydrogen bonds
to the two explicit aqua ligands of uranyl between the three models
suggests that the force field models may slightly underestimate the
strength and directional nature of the hydrogen bonds. This may explain
the low water exchange activation energy and the stability of the
six-coordinated metastable state relative to the structure with an
explicit water in the second shell, as predicted by the force field
models, Fig.~\ref{fig:free-energy-profiles}.

It should be noted that water activation energy barrier computed by
state of the art QM is sensitive to the choice of reaction path, level
of theory, and solvation model \cite{Tiwari14}. Tsushima \cite{Tsushima07}
reported the activation energy barriers of 10.1~kcal/mol and 4.8~kcal/mol
at the B3LYP level with the conductor-like polarizable continuum model
(CPCM) by considering the terminal states with one and two hydrogen
bonds, respectively. The latter value is close to the prediction of
4.5~kcal/mol in an earlier study \cite{Vallet01}, yet both are lower
than the experimental result. The comparison of PCM and a spherical
cavity self-consistent reaction field (SCRF) solvation models is reported
by Rotzinger \cite{Rotzinger07} using the complete active space self-consistent
field (CAS SCF) technique. The corresponding energy barriers of the
associative pathway are equal to 9.4~kcal/mol and 6.5~kcal/mol,
respectively. Water exchange energy barrier based on the thermodynamic
sampling is also estimated by Bühl et.\ al \cite{Buehl06b} using
the \textit{ab initio} molecular dynamics (AIMD). Their result of
6.7~kcal/mol is however lower than the experiment and the QM+RISM
value in this study. The good agreement of QM+RISM result with the
experiment and the best of other theoretical models indicates that
the RISM model provides a viable alternative solvation model for uranyl
aqua complexes in hybrid calculations and offers a reasonable accuracy
at an affordable cost.

\section{Conclusions}

We presented a parallelized implementation of 3D RISM for charged
species and verified its performance on examples of atomic ions and
uranyl aqua complexes. We demonstrated that the method is not susceptible
to a deficiency exposed in our previous work for the 1D RISM technique
where an artificial superposition of explicit aqua ligands and the
RISM medium leads to an overestimation of effective uranyl--water
coordination and an excessive elongation of the U--O and U--OW bonds:
for the uranyl aqua complexes with 4--6 explicit ligands RISM predict
essentially the same structure of the first solvation shell as PCM.
A net effect of a missing explicit aqua ligand is qualitatively correctly
represented by an excess density of the RISM medium, leading to comparable
results for uranyl with 5 and 4 explicit water ligands. However, because
of the strong closure order dependence of the results and occasional
divergence for higher order closures a successful model will avoid
treating the first solvation shell of highly charged ions implicitly.
With a first solvation shell treated by explicit aqua ligands, the
observables are nearly independent of the closure order.

A hybrid approach employing two very different models of ion--water
interactions, \emph{ab initio} and force field, relies heavily on
the consistence of the two interaction models. Use of the KL2 force
field, featuring high charges of interacting sites of uranyl, as selected
in our previous study as the one predicting the highest uranyl solvation
energy, leads to a substitution of explicit QM aqua ligands by the
RISM medium far too easily for routine calculations to be practical.
Such artifacts affected the 1D RISM approach as well suggesting that
this particular QM+RISM combination may need to be revisited.

Nevertheless, by employing the 3D RISM model for the second solvation
shell and the bulk water, the hybrid QM+RISM model produces meaningful
structure of the first solvation shell and solvation energies competitive
to the state of the art PCM. By examining the water exchange activation
energy we demonstrated the usefulness of the hybrid QM+RISM models
to address properties that are not easily reproduced by pure force
field models. The activation energy is predicted too low with the
force field models, likely due to the misrepresentation of hydrogen
bonding. It is in remarkably good agreement with the experiment when
the hybrid QM+RISM model is employed.

\section*{Acknowledgments}

We thank Prof.~H.-J.~Bungartz and co-workers for valuable discussions
and Dr.~L.~Jager and Prof.~M.~Griebel for providing the source
of the BGY3D program package. BL is grateful for a fellowship by the
International Graduate School of Science and Engineering at Technische
Universität München. This work was partially supported by Bundesministerium
für Bildung und Forschung (grant No.\ 01RC1106A).

\bibliography{uo2}

\end{document}

### 2015-02-06 doc/uo2_paper/uo2-II.tex
AM: include figures from cwd
AM: re-generate with elsevier style

@@
 %% Do not edit unless you really know what you are doing.
-\documentclass[english,twocolumn,showkeys,superscriptaddress]{revtex4}
+\documentclass[english,5p,sort&compress,times]{elsarticle}
 \usepackage[T1]{fontenc}
-\setcounter{secnumdepth}{3}
 \usepackage{color}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\@ifundefined{textcolor}{}
-{%
- \definecolor{BLACK}{gray}{0}
- \definecolor{WHITE}{gray}{1}
- \definecolor{RED}{rgb}{1,0,0}
- \definecolor{GREEN}{rgb}{0,1,0}
- \definecolor{BLUE}{rgb}{0,0,1}
- \definecolor{CYAN}{cmyk}{1,0,0,0}
- \definecolor{MAGENTA}{cmyk}{0,1,0,0}
- \definecolor{YELLOW}{cmyk}{0,0,1,0}
-}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-\usepackage{mathptmx}
-\usepackage{times}
+%\usepackage{mathptmx}
+%\usepackage{times}
+%\def\thetable{\arabic{table}}
 \usepackage{microtype}
-\def\thetable{\arabic{table}}
+\journal{Computational and Theoretical Chemistry}
 
@@
 
-\author{Alexei Matveev}
-
-\email{alexei.matveev@gmail.com}
-
-\selectlanguage{english}%
-
-\author{Bo Li}
-
-\affiliation{Department Chemie, Technische Universität München, 85747 Garching,
+\author[tum]{Alexei Matveev\corref{corr}}
+
+\ead{alexei.matveev@gmail.com}
+
+\author[tum]{Bo Li}
+
+\ead{fixme@abc.com}
+
+\author[tum,crc]{Notker Rösch}
+
+\ead{fixme@xyz.com}
+
+\cortext[corr]{Corresponding author}
+
+\address[tum]{Department Chemie, Technische Universität München, 85747 Garching,
 Germany}
 
-\author{Notker Rösch}
-
-\affiliation{Department Chemie and Catalysis Research Center, Technische Universität
-München, 85747 Garching, Germany}
-
-\date{Jan 2015}
+\address[crc]{Catalysis Research Center, Technische Universität München, 85747
+Garching, Germany}
 \begin{abstract}
@@
 RISM model for the bulk solvent is in close agreement with the experimental
-and the best theoretical estimates. 
-\end{abstract}
-
-\keywords{Reference interaction site model, aqueous solution, uranyl, solvation}
-
+and the best theoretical estimates. \end{abstract}
+\begin{keyword}
+Reference interaction site model, uranyl, aqueous solution
+\end{keyword}
 \maketitle
@@
 as chemical processes in radioactive waste disposal and recycling
-of used nuclear fuel \cite{book:Morss06,book:Grenthe92,Silva95}.
+of used nuclear fuel \citep{book:Morss06,book:Grenthe92,Silva95}.
 Quantum mechanics (QM) and classical molecular mechanics (MM) simulations
@@
 for dispersion-repulsion interactions and cavity formation energy
-using empiric parameters \cite{Tomasi05,Fuchs02}. The state of the
+using empiric parameters \citep{Tomasi05,Fuchs02}. The state of the
 art PCM is owing its success in part to the conductor-like screening
-model (COSMO) of electrostatic interactions \cite{Klamt95} and has
-successfully been applied in various studies of uranyl complexes \cite{Fuchs02,Moskaleva04,Cao05,Shamov05,Gutowski06}.
+model (COSMO) of electrostatic interactions \citep{Klamt95} and has
+successfully been applied in various studies of uranyl complexes \citep{Fuchs02,Moskaleva04,Cao05,Shamov05,Gutowski06}.
 On the other hand, there is a significant effort to improve the empirical
@@
 or Monte-Carlo (MC) methods, that yield the most explicit statistical
-solvent description \cite{Guilbaud93,Guilbaud96,Rai12,Pomogaev13,Kerisit13,Tiwari14}.
-Though several studies \cite{Gordon00,hagberg05,Frick09,Tirler13}
+solvent description \citep{Guilbaud93,Guilbaud96,Rai12,Pomogaev13,Kerisit13,Tiwari14}.
+Though several studies \citep{Gordon00,hagberg05,Frick09,Tirler13}
 attempt to combine QM and stochastic MM models such combinations will
@@
 continuum models due to statistical sampling techniques. Yet more
-demanding QM dynamic simulations \cite{Buehl05,Buehl06,Buehl11,Nichols08}
+demanding QM dynamic simulations \citep{Buehl05,Buehl06,Buehl11,Nichols08}
 are restricted to comparably small numbers of solvent molecules.
 
-Meanwhile, the integral equation theories of molecular liquids \cite{book:Hirata03,book:Hansen06},
+Meanwhile, the integral equation theories of molecular liquids \citep{book:Hirata03,book:Hansen06},
 may provide sufficient accuracy at an affordable cost for simulating
@@
 on contemporary hardware. Two approaches of this type are the reference
-interaction site model (RISM) \cite{Chandler72,Hirata81,Hirata82,Kinoshita97}
-and its three-dimensional generalization (3D RISM) \cite{Beglov95,Kovalenko98}
+interaction site model (RISM) \citep{Chandler72,Hirata81,Hirata82,Kinoshita97}
+and its three-dimensional generalization (3D RISM) \citep{Beglov95,Kovalenko98}
 which provide the solvent structure in form of site-site radial distribution
@@
 in the form of the excess chemical potential for several closure types
-\cite{Singer85,Kast08} and, hence, intuitive expressions for the
-first derivatives. The dielectrically consistent RISM (DRISM) \cite{Perkyns92}
+\citep{Singer85,Kast08} and, hence, intuitive expressions for the
+first derivatives. The dielectrically consistent RISM (DRISM) \citep{Perkyns92}
 was designed to compensate the error of the gas-like dielectric response
-of RISM for electrolytes and ionic solutes \cite{Perkyns94,Chiodo07,Chuev08,Joung13}.
+of RISM for electrolytes and ionic solutes \citep{Perkyns94,Chiodo07,Chuev08,Joung13}.
 
@@
 requires additional care when periodic boundary conditions are employed
-to simulate finite systems \cite{Makov95}. It was understood early
+to simulate finite systems \citep{Makov95}. It was understood early
 that the long-range asymptotes of the radial correlation functions
 induced by Coulomb interactions in RISM need a separate treatment
-as well \cite{Ng74}. For 3D RISM the few practical approaches include
-the background charge correction technique \cite{Kovalenko00b}, monopole
-correction \cite{Kloss08b}, and an exact re-summation for a superposition
-of gaussian charges \cite{Perkyns10}. In this work we present a generalization
-of the exact re-summation technique for arbitrary solute charge densities.
+as well \citep{Ng74}. For 3D RISM the few practical approaches include
+the background charge correction technique \citep{Kovalenko00b},
+monopole correction \citep{Kloss08b}, and an exact re-summation for
+a superposition of gaussian charges \citep{Perkyns10}. In this work
+we present a generalization of the exact re-summation technique for
+arbitrary solute charge densities.
 
@@
 has been first applied to formaldehyde and carbonyl compounds in water
-\cite{Tenno93,Tenno94}. An extension to 3D RISM SCF was developed
-for density functional theory (DFT) \cite{Kovalenko99} and \emph{ab
-initio} calculations \cite{Kloss08}. The analytical gradients for
-the self-consistent combination of 3D RISM and DFT \cite{Gusarov06}
+\citep{Tenno93,Tenno94}. An extension to 3D RISM SCF was developed
+for density functional theory (DFT) \citep{Kovalenko99} and \emph{ab
+initio} calculations \citep{Kloss08}. The analytical gradients for
+the self-consistent combination of 3D RISM and DFT \citep{Gusarov06}
 simplify explorations of free energy landscapes. Some of the recent
 developments of hybrid approaches combining integral equation theories
-and quantum chemistry are discussed in the recent review by Sato \cite{Sato13}.
+and quantum chemistry are discussed in the recent review by Sato \citep{Sato13}.
 
@@
 The 3D RISM equations operating with site-specific distributions can
-be written in the following form \cite{book:Hirata03}:
+be written in the following form \citep{book:Hirata03}:
 
@@
 solvent in 1D RISM equations for a specific temperature and solvent
-number density $\rho$. \cite{book:Hirata03}. The star denotes the
+number density $\rho$. \citep{book:Hirata03}. The star denotes the
 convolution integral and the sum over corresponding matrix indices.
@@
 by its (dimensionless) Fourier image $\tilde{\chi}_{\alpha\gamma}(k)$
-\cite{Chandler72,book:Hirata03}. The solvent susceptibility as a
+\citep{Chandler72,book:Hirata03}. The solvent susceptibility as a
 function of either $r$ or $k$ is spherically symmetric by construction.
@@
 be a relation between $h$ and the sum $x=-v+t$ sometimes referred
-to as a renormalized indirect correlation \cite{Kast08}:
+to as a renormalized indirect correlation \citep{Kast08}:
 \begin{equation}
@@
 by Kovalenko and Hirata (KH) for improving the numerical stability
-of the equations \cite{Kovalenko99}. Higher order closures with $n>1$
-are referred to as partial series expansion closures of order $n$
-(PSE$n$). For consistency, we will also refer to the KH closure as
-PSE1 in this work.
+of the equations \citep{Kovalenko99}. Higher order closures with
+$n>1$ are referred to as partial series expansion closures of order
+$n$ (PSE$n$). For consistency, we will also refer to the KH closure
+as PSE1 in this work.
 
@@
 the pair interactions were derived from site parameters using the
-Lorentz--Berthelot rules \cite{book:Hirata03}. The Ewald range parameter
+Lorentz--Berthelot rules \citep{book:Hirata03}. The Ewald range parameter
 $a=1.2$~Å$^{-1}$ is common for all site pairs. Further, to circumvent
@@
 of an unstructured solute and the HNC closure has been derived by
-Singer and Chandler (here in temperature units) \cite{Singer85}:
+Singer and Chandler (here in temperature units) \citep{Singer85}:
 \begin{equation}
@@
 \end{equation}
-A modification for the KH closure \cite{Kovalenko99} and a generalization
+A modification for the KH closure \citep{Kovalenko99} and a generalization
 for the arbitrary closure of the form $h=f(x)$ with $x=-v+t$ are
@@
 form for the PSE$n$ closures: $\phi_{n}(x)=-x^{n+1}/(n+1)!$ for
-$x>0$ and zero otherwise \cite{Kast08}. In the case of the KH (PSE1)
+$x>0$ and zero otherwise \citep{Kast08}. In the case of the KH (PSE1)
 closure the additional term $-x^{2}/2$ cancels the $h^{2}/2$ term
@@
 integral of the work required to ``switch on'' the external (solute)
-field $v$ \cite{Singer85,Kast08}. This property leads to straightforward
+field $v$ \citep{Singer85,Kast08}. This property leads to straightforward
 expressions for derivatives of the chemical potential with respect
@@
 of a QM solute, also on the electronic density $n_{\text{e}}$, but
-not on the solvent site distribution functions $h$ \cite{Bo15}.
+not on the solvent site distribution functions $h$ \citep{Bo15}.
 
@@
 where $\bar{E}$ is the average configurational energy of a water
-molecule which is about $-$9.9~kcal/mol \cite{Berendsen87,Mahoney00}
+molecule which is about $-$9.9~kcal/mol \citep{Berendsen87,Mahoney00}
 to be compared to the solvation energy of $-$6.3~kcal/mol derived
-from the ratio of the equilibrium vapor/liquid concentrations \cite{Bridgeman64}.
+from the ratio of the equilibrium vapor/liquid concentrations \citep{Bridgeman64}.
 
@@
 problem $F[t]=T[t]-t=0$ using the Newton--Krylov non-linear solver
-\cite{petsc-user-ref} with the analytic matrix-free Jacobian, $J[t]=\delta F/\delta t$,
+\citep{petsc-user-ref} with the analytic matrix-free Jacobian, $J[t]=\delta F/\delta t$,
 occasionally preceded by a few Picard iterations.
@@
 in the 3D case. All of the practical approaches, including the Ng
-scheme \cite{Ng74}, separate the smooth asymptote $v_{\text{L}}$
+scheme \citep{Ng74}, separate the smooth asymptote $v_{\text{L}}$
 of the site specific potential off the indirect correlation $t$ as
@@
 with the limit for $k\to0$ being proportional to the solvent compressibility
-$\kappa$ (here in temperature units) \cite{book:Hansen06}. More
+$\kappa$ (here in temperature units) \citep{book:Hansen06}. More
 importantly, this limit is the same for all site pairs $\alpha$ and
@@
 response of the solvent --- it is this term that is addressed by the
-dielectric correction of the DRISM approach \cite{Perkyns92,Cummings81,Kovalenko00b}.
+dielectric correction of the DRISM approach \citep{Perkyns92,Cummings81,Kovalenko00b}.
 The result of the convolution, 
@@
 at $r=0$ and $k=0$ as the discrete representation do not include
-these points by construction \cite{FFTW05}. The solute part is the
+these points by construction \citep{FFTW05}. The solute part is the
 electric form factor $\tilde{n}(\vec{k})$ of an arbitrary solute
@@
 A similar approach was applied before to a monopole field of the solute
-\cite{Kloss08b} and to a superposition of gaussian charge distributions
-\cite{Perkyns10}.  The argument line used above assumes the net
+\citep{Kloss08b} and to a superposition of gaussian charge distributions
+\citep{Perkyns10}.  The argument line used above assumes the net
 neutrality of the solvent species and is not directly applicable to
@@
 like in Eq.~\ref{eq:tau} can be assumed to be regular, cp.\ the
-discussion of the background charge correction in Ref.~\cite{Kovalenko00b}.
+discussion of the background charge correction in Ref.~\citep{Kovalenko00b}.
 An interpretation of Eq.~\ref{eq:tau} as a linear response of the
@@
 
-To model uranyl solvation we employ the SPC/E water model \cite{Berendsen87},
-modified as suggested by Pettitt and Rossky \cite{Pettitt82} by addition
-of a Lennard--Jones (LJ) term for hydrogen atoms with $\sigma_{\text{H}}=0.4$~Å
+To model uranyl solvation we employ the SPC/E water model \citep{Berendsen87},
+modified as suggested by Pettitt and Rossky \citep{Pettitt82} by
+addition of a Lennard--Jones (LJ) term for hydrogen atoms with $\sigma_{\text{H}}=0.4$~Å
 and $\epsilon_{\text{H}}=0.046$~kcal/mol (PR-SPC/E). For atomic
@@
 and $\epsilon_{\text{H}}=0.01553$~kcal/mol as proposed by Luchko
-and Kovalenko \cite{Luchko10} in order to compare with published
+and Kovalenko \citep{Luchko10} in order to compare with published
 results. The uranyl-water interaction parameters were derived from
 the second model proposed by Kerisit and Liu to strengthen the uranyl\textendash{}water
-interaction with the uranium charge of 3.5~\emph{e} (KL2) \cite{Kerisit13}.
-For convenience the reader may refer to our previous work \cite{Bo15}
+interaction with the uranium charge of 3.5~\emph{e} (KL2) \citep{Kerisit13}.
+For convenience the reader may refer to our previous work \citep{Bo15}
 where we summarized the literal values for the charge and LJ parameters
@@
 of Gaussian-type orbitals fitting-functions density functional method
-\cite{Dunlap90} as implemented in the parallel code ParaGauss \cite{ParaGaussV3.1}.
-The atomic approximation \cite{Matveev08} to the second-order Douglas\textendash{}Kroll\textendash{}Hess
-scalar-relativistic approach \cite{rev:roesch96} was used to account
+\citep{Dunlap90} as implemented in the parallel code ParaGauss \citep{ParaGaussV3.1}.
+The atomic approximation \citep{Matveev08} to the second-order Douglas\textendash{}Kroll\textendash{}Hess
+scalar-relativistic approach \citep{rev:roesch96} was used to account
 for relativistic effects; this approximation was shown to be very
@@
 exchange-correlation functional (generalized gradient approximation,
-GGA) \cite{Goerling99,kochDFT}, as suggested by Becke and Perdew
-(BP) \cite{Becke88,Perdew86c}. The Kohn\textendash{}Sham orbitals
+GGA) \citep{Goerling99,kochDFT}, as suggested by Becke and Perdew
+(BP) \citep{Becke88,Perdew86c}. The Kohn\textendash{}Sham orbitals
 were represented by flexible Gaussian-type basis sets, contracted
@@
 respectively. These basis sets and further computational details are
-the same as in earlier studies on uranyl species \cite{Moskaleva04,Moskaleva06b,Moskaleva06a,Kremleva12b,Bo15}.
-
-We used a COSMO PCM solvation model for comparison \cite{Fuchs02},
+the same as in earlier studies on uranyl species \citep{Moskaleva04,Moskaleva06b,Moskaleva06a,Kremleva12b,Bo15}.
+
+We used a COSMO PCM solvation model for comparison \citep{Fuchs02},
 with the cavity defined via an effective solvent radius of 1.4~Å
-and van der Waals radii of solute atoms as tabulated by Bondi \cite{Bondi64}
+and van der Waals radii of solute atoms as tabulated by Bondi \citep{Bondi64}
 and scaled by 1.125 (except for hydrogen). We carried out the cavity
-tessellation with the FIXPVA approach \cite{Su09,Kremleva12b} (fixed
+tessellation with the FIXPVA approach \citep{Su09,Kremleva12b} (fixed
 points with variable areas) to obtain numerically stable energies
-and forces \cite{Kremleva12b}.
+and forces \citep{Kremleva12b}.
 
@@
 method and the constrained minimization algorithm as implemented in
-the utility package ParaTools \cite{Chaffey-Millar11,Nikodem13}.
+the utility package ParaTools \citep{Chaffey-Millar11,Nikodem13}.
 In QM+RISM calculations of symmetric uranyl aqua complexes the U\textendash{}O
@@
 U\textendash{}O distances fixed to 179~pm motivated by earlier results
-\cite{Moskaleva04,Cao05,Buehl05}. However, for a study of the water
+\citep{Moskaleva04,Cao05,Buehl05}. However, for a study of the water
 exchange with MM+RISM we adopted a flexible uranyl model without symmetry
-constraints and compared two sets of intra-molecular parameters \cite{Kerisit13,Tiwari14}
+constraints and compared two sets of intra-molecular parameters \citep{Kerisit13,Tiwari14}
 with unconstrained QM+RISM results. The two uranyl force field models
@@
 a comparison to 1D RISM results for mono-atomic alkali and halide
-ions \cite{Joung13,Bo15} applying the cSPC/E water model \cite{Berendsen87,Luchko10}.
-With moderate settings  for the grid resolution of $64^{3}$ points
-in total and a cubic box volume of $20^{3}$~Å$^{3}$ we are able
-to match the 1D results up to 0.3~kcal/mol or better. The accuracy
-is slightly better for small cations than for the larger anions. The
-values are comparable to MD with SPC/E water and experimental results.
-The agreement may be slightly improved by a correction proportional
-to the molar volume of the solute \cite{Sergiievskyi14} as was demonstrated
-in our previous work \cite{Bo15}.
-
-\begin{table}
+ions \citep{Joung13,Bo15} applying the cSPC/E water model \citep{Berendsen87,Luchko10}.
+With moderate settings for the grid \textcolor{blue}{size} of $64^{3}$
+points in total and a cubic box volume of $20^{3}$~Å$^{3}$ we are
+able to match the 1D results up to \textcolor{blue}{0.2}~kcal/mol
+or better. \textcolor{blue}{With a 50\% larger box size at the same
+grid density the agreement improves to 0.07~kcal/mol; a further increase
+of the grid density by 50\% improves the results insignificantly.}
+The accuracy is slightly better for small cations than for the larger
+anions. The values are comparable to MD with SPC/E water and experimental
+results. The agreement may be slightly improved by a correction proportional
+to the molar volume of the solute \citep{Sergiievskyi14} as was demonstrated
+in our previous work \citep{Bo15}.
+
+\begin{table}[t]
 \begin{centering}
@@
 
-\begin{ruledtabular}
-
+\small\vspace{1ex}
+
+\begin{centering}
 \begin{tabular}{lrrrrrr}
+\hline 
  & Li$^{+}$ & Na$^{+}$ & K$^{+}$ & Cl$^{-}$ & Br$^{-}$ & I$^{-}$\tabularnewline
 \hline 
-3D KH & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.4 & $-$72.5 & $-$62.9\tabularnewline
-1D KH$^{b}$ & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.5 & $-$72.6 & $-$63.1\tabularnewline
-3D HNC & $-$111.8 & $-$86.3 & $-$69.3 & $-$79.8 & $-$74.1 & $-$64.7\tabularnewline
-1D HNC$^{b}$ & $-$111.7 & $-$86.3 & $-$69.3 & $-$79.9 & $-$74.2 & $-$64.9\tabularnewline
+KH & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.4 & $-$72.\textcolor{blue}{6} & $-$62.9\tabularnewline
+KH$^{b}$ & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.5 & $-$72.6 & $-$63.1\tabularnewline
+HNC & $-$111.\textcolor{blue}{7} & $-$86.3 & $-$69.3 & $-$79.8 & $-$74.1 & $-$64.7\tabularnewline
+HNC$^{b}$ & $-$111.7 & $-$86.3 & $-$69.3 & $-$79.9 & $-$74.2 & $-$64.9\tabularnewline
 MD$^{c}$ & $-$113.3 & $-$88.4 & $-$71.0 & $-$89.3 & $-$82.7 & $-$74.4\tabularnewline
@@
 \end{tabular}
-
-\end{ruledtabular}
-
-\footnotetext{$^{a}$~Energies in kcal/mol. KH and HNC closure relations
-have been used. 3D domain is $20^{3}$ Å$^{3}$ / $64^{3}$ pts. $^{b}$~Ref.~\cite{Bo15}.
-$^{c}$~MD with SPC/E water from Ref.~\cite{Joung08}. $^{d}$~Ref.~\cite{Schmid00}.} 
+\par\end{centering}
+
+\vspace{1ex}\footnotesize
+
+$^{a}$~Energies in kcal/mol. KH and HNC closure relations have been
+used. 3D domain is $20^{3}$ Å$^{3}$ / $64^{3}$ pts. $^{b}$~1D
+RISM from Ref.~\citep{Bo15}. $^{c}$~MD with SPC/E water from Ref.~\citep{Joung08}.
+$^{d}$~Ref.~\citep{Schmid00}.
 \end{table}
 
-\section{uranyl aqua complexes \label{sec:uranyl-aqua-complexes}}
+\section{Uranyl aqua complexes \label{sec:uranyl-aqua-complexes}}
 
@@
 to the findings in our previous work, using the 1D RISM formalism
-\cite{Bo15}, the MM+RISM combination features a vacancy in the equatorial
+\citep{Bo15}, the MM+RISM combination features a vacancy in the equatorial
 plane of the uranyl corresponding to the location of the missing water
@@
 
-\begin{table}
+\begin{table}[t]
 \caption{Average uranyl--water distance of the first solvation shell, $d$(U--OW),
@@
 \label{tab:mm-rism}}
-
-\begin{ruledtabular}
-
+\small\vspace{1ex}
+
+\begin{centering}
 \begin{tabular}{llrrrr}
+\hline 
 $n$ &  & 0 & 5$-$1$^{b}$ & 5 & 5+1$^{c}$\tabularnewline
@@
 \end{tabular}
-
-\end{ruledtabular}
-
-\footnotetext{ $^{a}$~Bond lengths in pm, energies in kcal/mol.
-$^{b}$~GP structure is $D_{4\text{h}}$ but the RISM structures
-feature a vacancy in the equatorial plane corresponding to the missing
-ligand, see Figure~\ref{fig:spatial-distribution-4w}. $^{c}$~Five-coordinated
-structure with the sixth water in the second shell. $^{d}$~Value
-not available due to divergence. $^{e}$~The GP row shows binding
-energies of $n$ water molecules to a rigid uranyl with $d$(U--O)
-of 179~pm. }
+\par\end{centering}
+
+\vspace{1ex}\footnotesize
+
+$^{a}$~Bond lengths in pm, energies in kcal/mol. $^{b}$~GP structure
+is $D_{4\text{h}}$ but the RISM structures feature a vacancy in the
+equatorial plane corresponding to the missing ligand, see Figure~\ref{fig:spatial-distribution-4w}.
+$^{c}$~Five-coordinated structure with the sixth water in the second
+shell. $^{d}$~Value not available due to divergence. $^{e}$~The
+GP row shows binding energies of $n$ water molecules to a rigid uranyl
+with $d$(U--O) of 179~pm. 
 \end{table}
-\begin{figure}
+\begin{figure}[t]
 \caption{Water oxygen distribution around the uranyl aqua complex with four
@@
 
-In our previous work exploring these models with 1D RISM \cite{Bo15}
+In our previous work exploring these models with 1D RISM \citep{Bo15}
 we observed consistently longer U--O and U--OW bond lengths in QM+RISM
@@
 complexes than in the gas phase by measurable 4--6~pm. The 1D RISM
-approach did not show this effect convincingly \cite{Bo15}. Again,
+approach did not show this effect convincingly \citep{Bo15}. Again,
 we interpret this effect of the medium as a screening of repulsive
@@
 vary only little with the closure order and agree well with $-$390~kcal/mol
-from our previous work \cite{Bo15}. However, the relative order of
-the estimates based on the four- and six-coordinated complexes is
-not reproduced. The estimate of $\Delta G_{\text{sol}}$ based on
-the four-coordinated complex is by $\sim$10~kcal/mol less negative
+from our previous work \citep{Bo15}. However, the relative order
+of the estimates based on the four- and six-coordinated complexes
+is not reproduced. The estimate of $\Delta G_{\text{sol}}$ based
+on the four-coordinated complex is by $\sim$10~kcal/mol less negative
 than the result of the standard model, Table~\ref{tab:qm+rism-vs-qm+pcm}.
@@
 is significantly larger than the value of 1.6~kcal/mol obtained using
-1D RISM \cite{Bo15}, which is remarkably close to the free energy
+1D RISM \citep{Bo15}, which is remarkably close to the free energy
 difference corresponding to the relative weights of four- and five-coordinated
 species determined from an average coordination number in a high-energy
-x-ray scattering (HEXS) experiment \cite{Soderholm05}. Given the
+x-ray scattering (HEXS) experiment \citep{Soderholm05}. Given the
 limitations of 1D RISM exposed by the current 3D RISM approach we
@@
  The QM+PCM solvation energy of $-426$~kcal/mol given by the standard
-five-coordinated uranyl model appears to be too low \cite{Moskaleva04}
+five-coordinated uranyl model appears to be too low \citep{Moskaleva04}
 in the view of most recent experimental estimate of $-369\pm15$~kcal/mol
-\cite{Gibson05}. The corresponding QM+RISM solvation energy of $-389$~kcal/mol
+\citep{Gibson05}. The corresponding QM+RISM solvation energy of $-389$~kcal/mol
 is significantly closer to that interval. The properties in Table~\ref{tab:qm+rism-vs-qm+pcm}
@@
 
-\begin{table}
+\begin{table}[t]
 \caption{Uranyl bond length, $d$(U--O), uranyl--water distance, $d$(U--OW),
@@
 
-\begin{ruledtabular}
-
+\small\vspace{1ex}
+
+\begin{centering}
 \begin{tabular}{llrrrr}
+\hline 
 $n$ &  & 0 & 4$^{b}$ & 5$^{c}$ & 6$^{d}$\tabularnewline
@@
 \end{tabular}
-
-\end{ruledtabular}
-
-\footnotetext{ $^{a}$~Bond lengths in pm, energies in kcal/mol.
-Symmetric uranyl aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$,
-$4\leq n\leq6$ and a bare uranyl ($n=0$) are treated by QM while
-interactions with the PR-SPC/E water solvent are treated by KL2 force
-field. $^{b}$~$D_{4\text{h}}$ structure. $^{c}$~$D_{5\text{h}}$
-structure. $^{d}$~$D_{3\text{d}}$ structure. $^{e}$~The row shows
-binding energies of $n$ water ligands. }
+\par\end{centering}
+
+\vspace{1ex}\footnotesize
+
+$^{a}$~Bond lengths in pm, energies in kcal/mol. Symmetric uranyl
+aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$, $4\leq n\leq6$
+and a bare uranyl ($n=0$) are treated by QM while interactions with
+the PR-SPC/E water solvent are treated by KL2 force field. $^{b}$~$D_{4\text{h}}$
+structure. $^{c}$~$D_{5\text{h}}$ structure. $^{d}$~$D_{3\text{d}}$
+structure. $^{e}$~The row shows binding energies of $n$ water ligands. 
 \end{table}
 
-\begin{figure}
+\begin{figure}[t]
 \caption{Radial distribution functions of the solvated uranyl without explicit
@@
 
-\begin{figure}
+\begin{figure}[t]
 \caption{Radial distribution functions of the solvated uranyl complex with
@@
 the experimentally available water exchange rate estimate of 1.3$\times$10$^{6}$~s$^{-1}$
-is about 9~kcal/mol \cite{Farkas00}. Many theoretical studies predict
+is about 9~kcal/mol \citep{Farkas00}. Many theoretical studies predict
 much faster water exchange than the available experimental data suggests
-\cite{Vallet01,hagberg05,Frick09,Nichols08,Kerisit10,Tiwari14}. A
-simple way to obtain an estimate of the activation energy for the
+\citep{Vallet01,hagberg05,Frick09,Nichols08,Kerisit10,Tiwari14}.
+A simple way to obtain an estimate of the activation energy for the
 water exchange may be the inspection of the U--OW potential of mean
@@
 from the MD RDF is 4.4 kcal/mol. These values are comparable to the
-5~kcal/mol derived from the residence time by Tiwari et al.~\cite{Tiwari14},
+5~kcal/mol derived from the residence time by Tiwari et al.~\citep{Tiwari14},
 but significantly lower than the experimental estimate. The transition
@@
 it was shown that uranyl bending essentially facilitates the water
-exchange \cite{Tiwari14}.
+exchange \citep{Tiwari14}.
 
@@
 at Hartree--Fock level of theory with water solvent described by CPCM
-model \cite{Vallet01}. A hypothetical water exchange in the equatorial
+model \citep{Vallet01}. A hypothetical water exchange in the equatorial
 plane appears to be significantly hindered at room temperature. The
@@
 
-\begin{figure}
+\begin{figure}[t]
 \caption{Potential of mean force, $W(r)$, for U--OW distribution functions
@@
 
-\begin{table}
+\begin{table}[t]
 \begin{centering}
@@
 
-\begin{ruledtabular}
-
+\small\vspace{1ex}
+
+\begin{centering}
 \begin{tabular}{lrrrrrrr}
+\hline 
  & $r_{\text{a}}$ & $W(r_{\text{a}})$ & $r_{\text{b}}$ & $W(r_{\text{b}})$ & $r_{\text{c}}$ & $W(r_{\text{c}})$ & $\Delta W_{\text{ab}}$\tabularnewline
@@
 \end{tabular}
-
-\end{ruledtabular}
-
-\footnotetext{$^{a}$~Distances in pm, energies in kcal/mol. Locations
-$r_{\text{a}}$ and $r_{\text{c}}$ correspond to the minima of first-
-and second solvation shells, respectively; $r_{\text{b}}$ is the
-highest point of the barrier connecting the two minima. } 
+\par\end{centering}
+
+\vspace{1ex}\footnotesize
+
+$^{a}$~Distances in pm, energies in kcal/mol. Locations $r_{\text{a}}$
+and $r_{\text{c}}$ correspond to the minima of first- and second
+solvation shells, respectively; $r_{\text{b}}$ is the highest point
+of the barrier connecting the two minima. 
 \end{table}
 
-\begin{figure}
+\begin{figure}[t]
 \caption{Potential of mean force, $W(x,y)$, for the U--OW distribution functions
@@
 
-\begin{figure}
+\begin{figure}[t]
 \caption{Free energy profiles, $G(q)$, along the water exchange pathway using
@@
 model as it was shown that the uranyl bending may affect the exchange
-rate \cite{Tiwari14}. The two intra-molecular force fields for a
+rate \citep{Tiwari14}. The two intra-molecular force fields for a
 flexible uranyl species we compare below differ primarily by the force
@@
 corresponds to the frequency of the bending mode of 159~cm$^{-1}$
-and $d_{0}$(U--O) of 176~pm \cite{Tiwari14}. The alternative ``hard''
+and $d_{0}$(U--O) of 176~pm \citep{Tiwari14}. The alternative ``hard''
 force field features a significantly higher frequency of the uranyl
 bending mode of 393~cm$^{-1}$ and a somewhat too long a uranyl bond
- $d_{0}$(U--O) of 180~pm \cite{Kerisit13}. The bending vibration
+ $d_{0}$(U--O) of 180~pm \citep{Kerisit13}. The bending vibration
 in crystals and solution has been experimentally characterized by
-frequencies in the range of 180--270~cm$^{-1}$ \cite{Denning76,Denning79,Toth81,Gal92}.
+frequencies in the range of 180--270~cm$^{-1}$ \citep{Denning76,Denning79,Toth81,Gal92}.
 The DFT models with 5--6 explicit water ligands that we employ in
 this and earlier works predict the bending frequency in this experimental
-range \cite{Matveev08}. For the gas phase uranyl the standard DFT
+range \citep{Matveev08}. For the gas phase uranyl the standard DFT
 predicts too soft a bending mode likely because of the self-interaction
-error for $f$-electrons \cite{Ramakrishnan11} while hybrid DFT and
-CCSD(T) estimate the bending frequency around 160--180~cm$^{-1}$
-\cite{Jackson08}. In this light the force field with the ``soft''
+error for $f$-electrons \citep{Ramakrishnan11} while hybrid DFT
+and CCSD(T) estimate the bending frequency around 160--180~cm$^{-1}$
+\citep{Jackson08}. In this light the force field with the ``soft''
 bending mode appears to be a more reliable choice.
@@
 is more than 6~pm longer than those of the other models employed
-here and in other works \cite{Tiwari14,Kerisit13} and should be considered
-a deficiency of the parametrization.  In turn, the ``soft'' model
-does not convincingly show the preference for five-coordinated uranyl
-with the sixth water in the second solvation shell over the six-coordinated
-broken symmetry metastable structure. The QM and ``hard'' force
-field models predict the six-coordinated $D_{3\text{d}}$-like state
-by 5.6~kcal/mol and 3.6~kcal/mol less stable than the five-coordinated
+here and in other works \citep{Tiwari14,Kerisit13} and should be
+considered a deficiency of the parametrization.  In turn, the ``soft''
+model does not convincingly show the preference for five-coordinated
+uranyl with the sixth water in the second solvation shell over the
+six-coordinated broken symmetry metastable structure. The QM and ``hard''
+force field models predict the six-coordinated $D_{3\text{d}}$-like
+state by 5.6~kcal/mol and 3.6~kcal/mol less stable than the five-coordinated
 structure with an extra water in the second shell, respectively. The
@@
 two-step water exchange mechanism as one of the possibilities. This
-is consistent with the findings of the other theoretical studies \cite{Tsushima07,Vallet01,Buehl06b,Rotzinger07,Tiwari14}
+is consistent with the findings of the other theoretical studies \citep{Tsushima07,Vallet01,Buehl06b,Rotzinger07,Tiwari14}
 suggesting that the associative mechanism is more favorable than the
@@
 which is in good agreement with the experimental estimate of 9~kcal/mol
-\cite{Farkas00}.  A comparison of locations and orientations of
+\citep{Farkas00}.  A comparison of locations and orientations of
 the second-shell explicit water which is connected by hydrogen bonds
@@
 state of the art QM is sensitive to the choice of reaction path, level
-of theory, and solvation model \cite{Tiwari14}. Tsushima \cite{Tsushima07}
+of theory, and solvation model \citep{Tiwari14}. Tsushima \citep{Tsushima07}
 reported the activation energy barriers of 10.1~kcal/mol and 4.8~kcal/mol
@@
 bonds, respectively. The latter value is close to the prediction of
-4.5~kcal/mol in an earlier study \cite{Vallet01}, yet both are lower
-than the experimental result. The comparison of PCM and a spherical
+4.5~kcal/mol in an earlier study \citep{Vallet01}, yet both are
+lower than the experimental result. The comparison of PCM and a spherical
 cavity self-consistent reaction field (SCRF) solvation models is reported
-by Rotzinger \cite{Rotzinger07} using the complete active space self-consistent
-field (CAS SCF) technique. The corresponding energy barriers of the
-associative pathway are equal to 9.4~kcal/mol and 6.5~kcal/mol,
-respectively. Water exchange energy barrier based on the thermodynamic
-sampling is also estimated by Bühl et.\ al \cite{Buehl06b} using
-the \textit{ab initio} molecular dynamics (AIMD). Their result of
-6.7~kcal/mol is however lower than the experiment and the QM+RISM
+by Rotzinger \citep{Rotzinger07} using the complete active space
+self-consistent field (CAS SCF) technique. The corresponding energy
+barriers of the associative pathway are equal to 9.4~kcal/mol and
+6.5~kcal/mol, respectively. Water exchange energy barrier based on
+the thermodynamic sampling is also estimated by Bühl et.\ al \citep{Buehl06b}
+using the \textit{ab initio} molecular dynamics (AIMD). Their result
+of 6.7~kcal/mol is however lower than the experiment and the QM+RISM
 value in this study. The good agreement of QM+RISM result with the
@@
 
+\bibliographystyle{elsarticle-num}
 \bibliography{uo2}

### 2015-04-07 doc/uo2_paper/uo2-II.tex
AM: re-apply figures from cwd
AM: regenerate tex fro JPCA

@@
 %% Do not edit unless you really know what you are doing.
-\documentclass[english,5p,sort&compress,times]{elsarticle}
+\documentclass[english,journal=jpcafh,layout=twocolumn,maxauthors=0]{achemso}
 \usepackage[T1]{fontenc}
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{esint}
+\usepackage[numbers]{natbib}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
+
+\title{Uranyl solvation by a three-dimensional reference interaction site
+model}
+\author{Alexei Matveev}
+\email{alexei.matveev@gmail.com}
+\author{Bo Li}
+\author{Notker Rösch}
+\email{roesch@mytum.de}
+\affiliation{Department Chemie, Technische Universität München, 85747 Garching,
+Germany}
+\alsoaffiliation{Catalysis Research Center, Technische Universität München, 85747
+Garching, Germany}
+\alsoaffiliation{Institute of High Performance Computing, Agency for Science, Technology
+and Research, 1 Fusionopolis Way, Connexis \#16-16, Singapore 138632,
+Singapore}
 %% Because html converters don't know tabularnewline
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\usepackage{mathptmx}
-%\usepackage{times}
+\usepackage{mathptmx}
+\usepackage{times}
+%\usepackage{microtype}
 %\def\thetable{\arabic{table}}
-\usepackage{microtype}
-\journal{Computational and Theoretical Chemistry}
+%\usepackage{ragged2e}
+%\usepackage[textfont=it]{caption}
+\captionsetup{font={bf}}
+\usepackage{natmove}
 
@@
 \begin{document}
-
-\title{Uranyl solvation by the three dimensional reference interaction site
-model}
-
-\author[tum]{Alexei Matveev\corref{corr}}
-
-\ead{alexei.matveev@gmail.com}
-
-\author[tum]{Bo Li}
-
-\ead{fixme@abc.com}
-
-\author[tum,crc]{Notker Rösch}
-
-\ead{fixme@xyz.com}
-
-\cortext[corr]{Corresponding author}
-
-\address[tum]{Department Chemie, Technische Universität München, 85747 Garching,
-Germany}
-
-\address[crc]{Catalysis Research Center, Technische Universität München, 85747
-Garching, Germany}
 \begin{abstract}
 We report an implementation of the three-dimensional reference interaction
-site model (3D RISM) specifically describing the treatment of the
-long-range Coulomb field of charged species represented by point charges
-and/or distributed charge density. A comparison of 1D and 3D results
-for atomic ions demonstrates the reasonable accuracy even for moderate
-box size and grid resolution. In an application to uranyl aqua complexes
-with 4\textendash{}6 explicit aqua ligands and an implicit bulk solvent
-modeled by RISM, we demonstrate that the 3D technique is not susceptible
-to the deficiencies of 1D technique exposed in our previous work {[}Li,
-Matveev, Krüger, Rösch, \emph{Comp. Theor. Chem.} \textbf{1051}, p.~151,
-(2015){]}. The 3D method eliminates the artificial superposition of
-explicit aqua ligands and the RISM medium. It predicts essentially
-the same values for uranyl and uranyl-water bond lengths as a state
-of the art polarizable continuum model. We find that with the first
-solvation shell treated explicitly, the observables are nearly independent
-of the closure order. The water exchange activation barrier obtained
-in a hybrid approach combining the quantum mechanical description
-of the uranyl interaction within the first solvation shell and the
-RISM model for the bulk solvent is in close agreement with the experimental
-and the best theoretical estimates. \end{abstract}
-\begin{keyword}
-Reference interaction site model, uranyl, aqueous solution
-\end{keyword}
+site model (3D RISM) that in particular addresses the treatment of
+the long-range Coulomb field of charged species, represented by point
+charges and/or a distributed charge density. A comparison of 1D and
+3D results for atomic ions demonstrates a reasonable accuracy, even
+for a moderate size of the unit cell and a moderate grid resolution.
+In an application to uranyl complexes with 4\textendash{}6 explicit
+aqua ligands and an implicit bulk solvent modeled by RISM, we show
+that the 3D technique is not susceptible to the deficiencies of the
+1D technique exposed in our previous work {[}Li, Matveev, Krüger,
+Rösch, \emph{Comp. Theor. Chem.} \textbf{1051}, p.~151, (2015){]}.
+The 3D method eliminates the artificial superposition of explicit
+aqua ligands and the RISM medium and predicts essentially the same
+values for uranyl and uranyl-water bond lengths as a state-of-the-art
+polarizable continuum model. With the first solvation shell treated
+explicitly, the observables are nearly independent of the order of
+the closure relationship used when solving the set of integral equations
+for the various distribution functions. Furthermore, we calculated
+the activation barrier of water exchange with a hybrid approach that
+combines the 3D RISM model for the bulk aqueous solvent and a quantum
+mechanical description (at the level of electronic Density Functional
+Theory) of uranyl interacting with explicitly represented water molecules.
+The calculated result agrees very well with experiment and the best
+theoretical estimates. 
+\end{abstract}
 \maketitle
@@
 Thermodynamic and structural properties of the uranyl(VI) cation,
-UO$_{2}^{2+}$, the most common uranium species in aqueous solution
+UO$_{2}^{2+}$, the most common uranium species in aqueous solution,
 are crucial for understanding its environmental chemistry as well
@@
 of used nuclear fuel \citep{book:Morss06,book:Grenthe92,Silva95}.
-Quantum mechanics (QM) and classical molecular mechanics (MM) simulations
-provide the two complimentary ways to address microscopic structure
-and energetics of solvated uranyl species. Till now, the most popular
-strategy to combine the accuracy of QM methods and the complexity
-of the statistical solvent description is to apply a polarizable continuum
-model (PCM) to represent the effective electrostatic field of the
-solvent, usually employed in combination with phenomenological corrections
-for dispersion-repulsion interactions and cavity formation energy
-using empiric parameters \citep{Tomasi05,Fuchs02}. The state of the
-art PCM is owing its success in part to the conductor-like screening
-model (COSMO) of electrostatic interactions \citep{Klamt95} and has
-successfully been applied in various studies of uranyl complexes \citep{Fuchs02,Moskaleva04,Cao05,Shamov05,Gutowski06}.
-On the other hand, there is a significant effort to improve the empirical
-model potentials for MM simulations, such as molecular dynamics (MD)
-or Monte-Carlo (MC) methods, that yield the most explicit statistical
-solvent description \citep{Guilbaud93,Guilbaud96,Rai12,Pomogaev13,Kerisit13,Tiwari14}.
-Though several studies \citep{Gordon00,hagberg05,Frick09,Tirler13}
-attempt to combine QM and stochastic MM models such combinations will
-inevitably lack the precise reproducibility  and efficiency of analytic
-continuum models due to statistical sampling techniques. Yet more
-demanding QM dynamic simulations \citep{Buehl05,Buehl06,Buehl11,Nichols08}
+Simulations on the basis of quantum mechanics (QM) or classical molecular
+mechanics (MM) provide two complimentary ways of addressing the microscopic
+structure and the energetics of solvated uranyl species. Till now,
+the most popular strategy for combining the accuracy of QM methods
+and the complexity of the statistical solvent description is to apply
+a polarizable continuum model (PCM) to represent the effective electrostatic
+field of the solvent, usually employed in combination with phenomenological
+corrections (largely relying on empirical parameters) for dispersion-repulsion
+interactions and the cavity formation energy \citep{Tomasi05,Fuchs02}.
+The conductor-like screening model (COSMO) of electrostatic interactions
+is a state-of-the-art PCM \citep{Klamt95}; it was successfully applied
+in various studies of uranyl complexes \citep{Fuchs02,Moskaleva04,Cao05,Shamov05,Gutowski06}.
+On the other hand, significant efforts are being made on improving
+the empirical model potentials for MM simulations, be it molecular
+dynamics (MD) or Monte-Carlo (MC) methods, that yield the most explicit
+statistical description of the solvent \citep{Guilbaud93,Guilbaud96,Rai12,Pomogaev13,Kerisit13,Tiwari14,book:Griebel07}.
+Although several studies \citep{Gordon00,hagberg05,Frick09,Tirler13}
+attempted to combine QM and stochastic MM models, such combinations
+will inevitably lack the precise reproducibility and the efficiency
+of analytic continuum models due to statistical sampling techniques.
+More demanding QM dynamic simulations \citep{Buehl05,Buehl06,Buehl11,Nichols08}
 are restricted to comparably small numbers of solvent molecules.
 
-Meanwhile, the integral equation theories of molecular liquids \citep{book:Hirata03,book:Hansen06},
-may provide sufficient accuracy at an affordable cost for simulating
-chemical processes in solution by employing standard numerical methods
-on contemporary hardware. Two approaches of this type are the reference
-interaction site model (RISM) \citep{Chandler72,Hirata81,Hirata82,Kinoshita97}
-and its three-dimensional generalization (3D RISM) \citep{Beglov95,Kovalenko98}
-which provide the solvent structure in form of site-site radial distribution
+Alternatively, the integral equation theories of molecular liquids
+\citep{book:Hirata03,book:Hansen06} may provide sufficient accuracy
+at an affordable cost for simulating chemical processes in solution
+by employing standard numerical methods on contemporary hardware.
+Two variants of this type are the reference interaction site model
+(RISM) \citep{Chandler72,Hirata81,Hirata82,Kinoshita97} and its three-dimensional
+generalization (3D RISM) \citep{Beglov95,Kovalenko98}. These methods
+provide the solvent structure in form of site-site radial distribution
 functions (RDFs) or three-dimensional distribution functions of solvent
 sites in an external (solute) field, respectively. Both 1D and 3D
-theories provide variational expressions for the solvation energy
-in the form of the excess chemical potential for several closure types
-\citep{Singer85,Kast08} and, hence, intuitive expressions for the
-first derivatives. The dielectrically consistent RISM (DRISM) \citep{Perkyns92}
-was designed to compensate the error of the gas-like dielectric response
-of RISM for electrolytes and ionic solutes \citep{Perkyns94,Chiodo07,Chuev08,Joung13}.
-
-A correct handling of the electrostatic field of a charge distribution
+methods offer variational expressions for the solvation energy in
+the form of the excess chemical potential for several types of closure
+of the hierarchy of equations \citep{Singer85,Kast08}, hence admit
+intuitive expressions for the first-order derivatives with respect
+to nuclear displacements. For another class of integral equations
+based on the Born--Green--Yvon hierarchy \citep{Fischer80,Griebel08}
+the lack of a practical free energy expression complicates exploring
+free energy landscapes. Note that the dielectrically consistent RISM
+(DRISM) \citep{Perkyns92} is able to compensate the error of the
+gas-like dielectric response of standard RISM for electrolytes and
+ionic solutes \citep{Perkyns94,Chiodo07,Chuev08,Joung13}.
+
+Handling correctly the electrostatic field of a charge distribution
 requires additional care when periodic boundary conditions are employed
 to simulate finite systems \citep{Makov95}. It was understood early
-that the long-range asymptotes of the radial correlation functions
+on that the long-range asymptotes of the radial correlation functions
 induced by Coulomb interactions in RISM need a separate treatment
-as well \citep{Ng74}. For 3D RISM the few practical approaches include
-the background charge correction technique \citep{Kovalenko00b},
-monopole correction \citep{Kloss08b}, and an exact re-summation for
-a superposition of gaussian charges \citep{Perkyns10}. In this work
+as well \citep{Ng74}. The few practical approaches to 3D RISM pursue
+various strategies, e.g.\ inclusion of a background charge correction
+\citep{Kovalenko00b}, monopole correction \citep{Kloss08b}, and
+an exact re-summation for a superposition of Gaussian-type charge
+distributions \citep{Perkyns10}. Very recently, a new efficient technique
+has been presented to avoid integrating in real space the long-range
+contributions to the solvation energy \citep{Heil15}. In this work
 we present a generalization of the exact re-summation technique for
@@
 electronic structure theory for self-consistent field calculations
-(RISM SCF) in the pioneering work of Ten-no, Hirata, and Kato and
-has been first applied to formaldehyde and carbonyl compounds in water
-\citep{Tenno93,Tenno94}. An extension to 3D RISM SCF was developed
+(RISM SCF) in the pioneering work of Ten-no, Hirata, and Kato who
+first applied the method to formaldehyde and carbonyl compounds in
+water \citep{Tenno93,Tenno94}. An extension to 3D RISM SCF was developed
 for density functional theory (DFT) \citep{Kovalenko99} and \emph{ab
-initio} calculations \citep{Kloss08}. The analytical gradients for
-the self-consistent combination of 3D RISM and DFT \citep{Gusarov06}
-simplify explorations of free energy landscapes. Some of the recent
-developments of hybrid approaches combining integral equation theories
-and quantum chemistry are discussed in the recent review by Sato \citep{Sato13}.
-
-In this work, we first describe the implementation of the 3D RISM
-approach and a numerical treatment of the long-range Coulomb field
-of the charged solute, Section~\ref{sec:methodology}, and then validate
-the implementation by comparing the results of 1D and 3D techniques
-for alkali and halide ions, Section~\ref{sec:simple-ions}. We further
-evaluate the performance of the method for the uranyl(VI) dication
-in aqueous solution represented by aqua complexes $[\text{UO}_{2}(\text{H}_{2}\text{O})_{n}]^{2+}$,
-($n=4-6$), using a force field, Section~\ref{sub:MM+RISM}, and
-a quantum mechanical representation of interactions within the first
-solvation shell, Section~\ref{sub:QM+RISM}. We also compare the
-results to those of PCM calculations. The activation barrier of water
-exchange in the first solvation shell is analyzed in Section~\ref{sub:water-exchange}
-using QM and MM models of interactions in the first solvation shell.
+initio} calculations \citep{Kloss08}. Analytic energy gradients with
+respect to nuclear displacements for the self-consistent combination
+of DFT and 3D RISM \citep{Gusarov06} simplify exploring free energy
+landscapes of complex molecular systems. Some of the recent developments
+of hybrid approaches combining integral equation theories and quantum
+chemistry have been discussed in a recent review \citep{Sato13}.
+
+In this work, we first describe our implementation of the 3D RISM
+approach as well as a numerical treatment of the long-range Coulomb
+field of a charged solute, Section~\ref{sec:methodology}. Then we
+validate the implementation by comparing the results of 1D and 3D
+techniques for alkali and halide ions, Section~\ref{sec:simple-ions}.
+We further evaluate the performance of the method for the uranyl(VI)
+dication in aqueous solution, represented by aqua complexes $[\text{UO}_{2}(\text{H}_{2}\text{O})_{n}]^{2+}$,
+($n=4-6$), using a force field description, Section~\ref{sub:MM+RISM},
+and a quantum mechanical representation of the interactions within
+the first solvation shell, Section~\ref{sub:QM+RISM}. We also compare
+the results to those of PCM calculations. Finally, we analyze the
+activation barrier for exchanging a water molecule in the first solvation
+shell, Section~\ref{sub:water-exchange}, using QM and MM models
+of interactions. With the exception of the water exchange process
+the test cases used to evaluate the 3D RISM approach are the same
+as in our preceding work on 1D RISM \citep{Bo15}.
 
@@
 field of a solute of general shape. The solute-solvent interactions
-are thus assumed to be pairwise, which in this formulation excludes
-\emph{polarizable} solutes. Further, the interactions are usually
+are assumed to be pairwise, which in the present formulation excludes
+\emph{polarizable} solutes.  Furthermore, the interactions are usually
 approximated by superposition of pairwise site-site contributions.
 The 3D RISM equations operating with site-specific distributions can
-be written in the following form \citep{book:Hirata03}:
+be written as follows \citep{book:Hirata03}:
 
@@
 \end{equation}
-Here $h$, $c$, and $t$ are the vectors of total-, direct-, and
-indirect correlation functions $h_{\alpha}(\vec{r})$, $c_{\alpha}(\vec{r})$,
+Here $h$, $c$, and $t$ are the vectors of total, direct, and indirect
+correlation functions $h_{\alpha}(\vec{r})$, $c_{\alpha}(\vec{r})$,
 and $t_{\alpha}(\vec{r})$. All are dimensionless unknowns of the
-equations that will be represented on a uniform grid of points. The
-greek subscripts, such as $\alpha$, enumerate solvent interaction
-sites.
+equations that will be represented on a uniform grid of points. Greek
+subscripts, e.g.\,$\alpha$, enumerate solvent interaction sites.
 
 The solvent susceptibility $\chi$ in the first of the three Eqs.~\ref{eq:rsim-h-c-t}
-is represented by its (square symmetric) matrix $\chi_{\alpha\gamma}(r)$
+is represented by a (symmetric square) matrix $\chi_{\alpha\gamma}(r)$
 derived from the radial site-site correlation functions of the pure
 solvent in 1D RISM equations for a specific temperature and solvent
-number density $\rho$. \citep{book:Hirata03}. The star denotes the
-convolution integral and the sum over corresponding matrix indices.
-For that reason it is convenient to represent the solvent susceptibility
-by its (dimensionless) Fourier image $\tilde{\chi}_{\alpha\gamma}(k)$
-\citep{Chandler72,book:Hirata03}. The solvent susceptibility as a
-function of either $r$ or $k$ is spherically symmetric by construction.
+number density $\rho$ \citep{book:Hirata03}. The star denotes the
+convolution integral and a summation over corresponding matrix indices.
+It is convenient to represent the solvent susceptibility by its (dimensionless)
+Fourier image $\tilde{\chi}_{\alpha\gamma}(k)$ \citep{Chandler72,book:Hirata03}.
+The solvent susceptibility as a function of either $r$ or $k$ is
+spherically symmetric by construction.
 
@@
 --- the third and last of Eqs.~\ref{eq:rsim-h-c-t}. Closure is a
-local relation between the values of site potential $v_{\alpha}(\vec{r})$
-and, in this case, the indirect correlation $t_{\alpha}(\vec{r})$
-at each point $\vec{r}$ and the value of the direct correlation $h_{\alpha}(\vec{r})$
-at this point. It is the only non-linear relation between the unknowns
-$h$, $c$ and $t$ in Eqs.~\ref{eq:rsim-h-c-t}. The site-specific
-potentials $v$ are measured in units of temperature and thus are
-also dimensionless. For the purposes of this work we have chosen a
-very specific, though popular, form of the closure which happens to
-be a relation between $h$ and the sum $x=-v+t$ sometimes referred
-to as a renormalized indirect correlation \citep{Kast08}:
+local relation between the values of the site potential $v_{\alpha}(\vec{r})$
+and, in this case, the values of the indirect correlation $t_{\alpha}(\vec{r})$
+and the direct correlation $h_{\alpha}(\vec{r})$, at each point $\vec{r}$.
+It is the only non-linear relation between the unknowns $h$, $c$
+and $t$ in Eqs.~\ref{eq:rsim-h-c-t}. The site-specific potentials
+$v$ are measured in units of temperature and thus are also dimensionless.
+For the purposes of this work we have chosen a very specific, though
+popular, form of the closure, which happens to be a relation between
+$h$ and the sum $x=-v+t$ sometimes referred to as renormalized indirect
+correlation \citep{Kast08}:
 \begin{equation}
@@
 \mathrm{e}^{x}-1 & x\le0\\
-\sum_{k=1}^{n}x^{k}/k! & x>0
+\sum_{k=1}^{p}x^{k}/k! & x>0
 \end{cases}\label{eq:closure-pse}
 \end{equation}
-In the limit of $n\to\infty$ one recovers the hypernetted chain (HNC)
-closure. Another popular closure is obtained for $n=1$. It was proposed
-by Kovalenko and Hirata (KH) for improving the numerical stability
-of the equations \citep{Kovalenko99}. Higher order closures with
-$n>1$ are referred to as partial series expansion closures of order
-$n$ (PSE$n$). For consistency, we will also refer to the KH closure
-as PSE1 in this work.
+In the limit of $p\to\infty$ one recovers the hypernetted chain (HNC)
+closure \citep{Leeuwen59}. Kovalenko and Hirata (KH) proposed another
+popular closure $(p=1)$ for improving the numerical stability of
+the equations \citep{Kovalenko99}. Higher-order closure relations,
+with $p>1$, are referred to as partial series expansion closures
+of order $p$ (PSE$p$). For consistency, in this work we will refer
+to the KH closure as PSE1.
 
@@
 is defined as the sum of the Coulomb interaction $u_{i\alpha}^{\text{C}}(r)=q_{i}q_{\alpha}a[f_{\text{CS}}(ar)+f_{\text{CL}}(ar)]$,
-where we separated the singular short-range $f_{\text{CS}}(r)=\mathrm{erfc}(r)/r$
-from the regular long-range $f_{\text{CL}}(r)=\mathrm{erf}(r)/r$
-part, and a Lennard--Jones (LJ) type potential, $u_{i\alpha}^{\text{LJ}}(r)=\epsilon_{i\alpha}f_{\text{LJ}}(r/\sigma_{i\alpha})$
-with a LJ form $f_{\text{LJ}}(r)=4(r^{-12}-r^{-6})$. Parameters of
-the pair interactions were derived from site parameters using the
-Lorentz--Berthelot rules \citep{book:Hirata03}. The Ewald range parameter
-$a=1.2$~Å$^{-1}$ is common for all site pairs. Further, to circumvent
-numerical problems with singularities the primitive forms $f_{\text{CS}}(r)$
-and $f_{\text{LJ}}(r)$ were regularized by substituting them with
-quadratic functions $\bar{f}(r)=f(b)+f^{\prime}(b)(r^{2}-b^{2})/2b$
+where we separated the singular short-range part $f_{\text{CS}}(r)=\mathrm{erfc}(r)/r$
+from the regular long-range part $f_{\text{CL}}(r)=\mathrm{erf}(r)/r$,
+and a Lennard--Jones (LJ) type potential, $u_{i\alpha}^{\text{LJ}}(r)=\epsilon_{i\alpha}f_{\text{LJ}}(r/\sigma_{i\alpha})$
+with $f_{\text{LJ}}(r)=4(r^{-12}-r^{-6})$. Parameters of the pair
+interactions were derived from site parameters using the Lorentz--Berthelot
+rules \citep{book:Hirata03}. The Ewald range parameter $a=1.2$~Å$^{-1}$
+is common to all site pairs. To avoid numerical problems due to singularities,
+the primitive forms $f_{\text{CS}}(r)$ and $f_{\text{LJ}}(r)$ were
+regularized by replacing them by quadratic functions $\bar{f}(r)=f(b)+f^{\prime}(b)(r^{2}-b^{2})/2b$
 for $r<b$. For our systems at room temperature the results were almost
-indistinguishable for finite $b\le0.5$ so that we settled with $b=0.2$
-in this work.
+indistinguishable for $0<b\le0.5$ so that we chose $b=0.2$.
 
@@
 
-The analytical expression for the excess chemical potential $\mu$
-of an unstructured solute and the HNC closure has been derived by
-Singer and Chandler (here in temperature units) \citep{Singer85}:
+Singer and Chandler derived an analytical expression for the excess
+chemical potential $\mu$ of an unstructured solute in the case of
+HNC closure (in temperature units) \citep{Singer85}:
 \begin{equation}
@@
 A modification for the KH closure \citep{Kovalenko99} and a generalization
-for the arbitrary closure of the form $h=f(x)$ with $x=-v+t$ are
-available: 
+to an arbitrary closure of the form $h=f(x)$ with $x=-v+t$ are available:
 \begin{equation}
@@
 \end{equation}
-with $\phi(x)=f(x)-x-\int_{0}^{x}f(y)dy$ having particularly a simple
-form for the PSE$n$ closures: $\phi_{n}(x)=-x^{n+1}/(n+1)!$ for
-$x>0$ and zero otherwise \citep{Kast08}. In the case of the KH (PSE1)
-closure the additional term $-x^{2}/2$ cancels the $h^{2}/2$ term
-of the HNC integrand, Eq.~\ref{eq:chem-pot-hnc}, in the regions
-where $x=h>0$.
+For PSE$p$ closures, $\phi(x)=f(x)-x-\int_{0}^{x}f(y)dy$ has a particularly
+simple form: $\phi_{p}(x)=-x^{p+1}/(p+1)!$ for $x>0$ and zero otherwise
+\citep{Kast08}. In the case of the KH (PSE1) closure, the additional
+term $-x^{2}/2$ cancels the $h^{2}/2$ term of the HNC integrand,
+Eq.~\ref{eq:chem-pot-hnc}, in the regions where $x=h>0$.
 
@@
 \end{equation}
-where $E[n_{\text{e}},s]$ is the solute internal energy depending
-on the solutes' geometric degrees of freedom $s$ and, in the case
+where $E[n_{\text{e}},s]$ is the solute internal energy that depends
+on the solute's geometric degrees of freedom $s$ and, in the case
 of a QM solute, also on the electronic density $n_{\text{e}}$, but
@@
 
-The uranyl solvation energy is then estimated as\textcolor{green}{{}
-}
-\begin{eqnarray}
-\Delta G_{\text{sol}} & = & G[\text{UO}_{2}(\text{H}_{2}\text{O})_{n}^{2+}]-nG[\text{H}_{2}\text{O}]-E[\text{UO}_{2}^{2+}]\label{eq:dG-sol-definition}
-\end{eqnarray}
+The solvation energy of uranyl is estimated as\textcolor{green}{{} }
+\begin{equation}
+\Delta G_{\text{sol}}=G[\text{UO}_{2}(\text{H}_{2}\text{O})_{n}^{2+}]-nG[\text{H}_{2}\text{O}]-E[\text{UO}_{2}^{2+}]\label{eq:dG-sol-definition}
+\end{equation}
 for $n\ge0$ explicit aqua ligands. Note that without proper statistical
@@
 of the solvation energy, justified to some extent for sufficiently
-rigid solutes, breaks for increasingly many explicit water molecules
-outside of the tightly bound first solvation shell as $n\to\infty$.
-Indeed, in a gedankenexperiment the difference $G[(\text{H}_{2}\text{O})_{n}]-nG[\text{H}_{2}\text{O}]$
+rigid solutes, does not hold for systems with a large number $n\to\infty$
+of explicit water molecules outside of the tightly bound first solvation
+shell. Indeed, in a gedanken experiment the difference $G[(\text{H}_{2}\text{O})_{n}]-nG[\text{H}_{2}\text{O}]$
 for the above definition of $G$ applied to water droplets asymptotically
@@
 where $\bar{E}$ is the average configurational energy of a water
-molecule which is about $-$9.9~kcal/mol \citep{Berendsen87,Mahoney00}
-to be compared to the solvation energy of $-$6.3~kcal/mol derived
-from the ratio of the equilibrium vapor/liquid concentrations \citep{Bridgeman64}.
-
-\subsection{Numerical solution of 3D RISM equations}
-
-Because of the highly non-linear closure relation, a single variable
-fix point iteration scheme, $t^{(n+1)}=T[t^{(n)}]$, with $T[t]=(\chi-1)\ast c(t)$
+molecule; its value is about $-$9.9~kcal/mol \citep{Berendsen87,Mahoney00}.
+This result is to be compared to the solvation energy of $-$6.3~kcal/mol
+derived from the ratio of the equilibrium vapor/liquid concentrations
+\citep{Bridgeman64}.
+
+\subsection{Numerical solution of the 3D RISM equations}
+
+Because of the highly non-linear closure relation, a single-variable
+fix-point iteration scheme, $t^{(n+1)}=T[t^{(n)}]$, with $T[t]=(\chi-1)\ast c(t)$
 and $c(t)=f(-v+t)-t$ as derived from the non-linear equation system,
@@
 cases. To overcome the divergence problem we solve the reduced non-linear
-problem $F[t]=T[t]-t=0$ using the Newton--Krylov non-linear solver
-\citep{petsc-user-ref} with the analytic matrix-free Jacobian, $J[t]=\delta F/\delta t$,
-occasionally preceded by a few Picard iterations.
+problem $F[t]=T[t]-t=0$ using a Newton--Krylov non-linear solver
+\citep{petsc-user-ref} with the analytic matrix-free Jacobian, $J[t]=\delta F/\delta t$;
+the procedure is occasionally preceded by a few Picard iterations.
 
 The long-range field of a charged solute requires a special treatment
-in the 3D case. All of the practical approaches, including the Ng
-scheme \citep{Ng74}, separate the smooth asymptote $v_{\text{L}}$
-of the site specific potential off the indirect correlation $t$ as
-well as $-v_{\text{L}}$ off the direct correlation $c$ to operate
+in the 3D case. All practical approaches, including the scheme of
+Ng \citep{Ng74}, separate the smooth asymptote $v_{\text{L}}$ of
+the site-specific potential from the indirect correlation $t$ as
+well as $-v_{\text{L}}$ from the direct correlation $c$ to operate
 primarily with the short range counterparts $v_{\text{S}}$ , $c_{\text{S}}$,
@@
 $t_{\text{S}}$ (and the resulting non-linear problem) will look slightly
-different, though:
-
-\begin{eqnarray}
-T_{\text{S}}[t_{\text{S}}] & = & (\chi-1)\ast c_{\text{S}}(t_{\text{S}})-\tau
-\end{eqnarray}
-with $c_{\text{S}}(t_{\text{S}})=f(-v_{\text{S}}+t_{\text{S}})-t_{S}$
-and the only term that involves a long-range operand being the constant
-``renormalization'' term $\tau=\chi\ast v_{\text{L}}$. This term
-is the convolution of solvent susceptibility with the asymptotic potential.
-
-The low $k$ behavior of the asymptotic potential $v_{\text{L}}$
-originates from the Coulomb potential and is, in general, singular.
+different:
+
+\begin{equation}
+T_{\text{S}}[t_{\text{S}}]=(\chi-1)\ast c_{\text{S}}(t_{\text{S}})-\tau
+\end{equation}
+with $c_{\text{S}}(t_{\text{S}})=f(-v_{\text{S}}+t_{\text{S}})-t_{S}$.
+The constant ``renormalization'' term $\tau=\chi\ast v_{\text{L}}$,
+the convolution of the solvent susceptibility with the asymptotic
+potential, is the only one that involves a long-range operand.
+
+The behavior of the asymptotic potential $v_{\text{L}}$ at low $k$
+values originates from the Coulomb potential and is, in general, singular.
 With our choice of separation into short- and long-range parts it
@@
 \end{equation}
-with $\tilde{n}(\vec{k})$ being the solute electric form factor,
-$Q=\tilde{n}(0)$ being the net solute charge, $q$ being the vector
-of the solvent site charges $q_{\alpha}$ and $\tilde{\varphi}(k)=4\pi k^{-2}\exp(-k^{2}/4a^{2})$
-the Fourier transform of the long range potential function $\varphi(r)=\mathrm{erf}(ar)/r$.
+where $\tilde{n}(\vec{k})$ is the solute electric form factor, $Q=\tilde{n}(0)$
+the net solute charge, $q$ the vector of the solvent site charges
+$q_{\alpha}$ and $\tilde{\varphi}(k)=4\pi k^{-2}\exp(-k^{2}/4a^{2})$
+the Fourier transform of the long-range potential function $\varphi(r)=\mathrm{erf}(ar)/r$.
 A straightforward representation of this convolution operand on a
-3D grid is thus problematic and should be avoided. The result of the
-convolution, $\tau$, is of short range, though.
-
-Indeed, the low $k$ behavior of another convolution operand, the
+3D grid is thus problematic and should be avoided. However, the result
+of the convolution, $\tau$, is of short range.
+
+Indeed, the behavior of another convolution operand at low $k$, the
 solvent susceptibility is regular 
@@
 importantly, this limit is the same for all site pairs $\alpha$ and
-$\gamma$. Though not enforced, this property is obeyed in practical
-calculations with reasonable accuracy. The second, $o(k^{2})$, term
+$\gamma$. Though not enforced, this property is obeyed with reasonable
+accuracy in practical calculations. The second term, of $o(k^{2})$,
 is site pair specific and is in particular responsible for the dielectric
-response of the solvent --- it is this term that is addressed by the
-dielectric correction of the DRISM approach \citep{Perkyns92,Cummings81,Kovalenko00b}.
-The result of the convolution, 
+response of the solvent. This term is addressed by the dielectric
+correction in the DRISM approach \citep{Perkyns92,Cummings81,Kovalenko00b}.
+The result of the\textcolor{blue}{{} }convolution, 
 \begin{equation}
-\tilde{\tau}_{\alpha}(\vec{k})=\tilde{n}(\vec{k})\left[\tilde{\varphi}(k)\sum_{\gamma}\tilde{\chi}_{\alpha\gamma}(k)q_{\gamma}\right]\sim2\pi Q\sum_{\gamma}d_{\alpha\gamma}q_{\gamma}\label{eq:tau}
+\tilde{\tau}_{\alpha}(\vec{k})=\tilde{n}(\vec{k})\tilde{\theta}_{\alpha}(k)\label{eq:tau}
 \end{equation}
-is regular assuming net neutrality of the solvent, $\sum_{\gamma}q_{\gamma}=0$,
-which makes the compressibility term vanish. Note that the terms in
-square brackets, $\tilde{\varphi}(k)\sum_{\gamma}\tilde{\chi}_{\alpha\gamma}(k)q_{\gamma}=:\tilde{\theta}_{\alpha}(k)$
-(one per solvent site $\alpha$), are convolutions of spherically
-symmetric functions neither of which is solute specific. The functions
-$\tilde{\theta}_{\alpha}(k)$ are site-specific solvent properties
-representing response to a unit (here gaussian-broadened) charge that
-should be computed using 1D techniques, just as the solvent susceptibility
+expressed here via the regular auxiliary functions
+\begin{equation}
+\tilde{\theta}_{\alpha}(k)=\tilde{\varphi}(k)\sum_{\gamma}\tilde{\chi}_{\alpha\gamma}(k)q_{\gamma}\sim2\pi\sum_{\gamma}d_{\alpha\gamma}q_{\gamma}\label{eq:theta}
+\end{equation}
+is also regular for an overall neutral solvent, $\sum_{\gamma}q_{\gamma}=0$.
+In that case, the compressibility term in Eq.~\ref{eq:theta} vanishes
+after summation. Note that the functions $\tilde{\theta}_{\alpha}(k)$
+(one per solvent site $\alpha$) are convolutions of spherically symmetric
+functions neither of which is solute specific. The functions $\tilde{\theta}_{\alpha}(k)$
+are site-specific solvent properties which represent the response
+to a unit charge (here broadened as a Gaussian). These quantities
+can be computed using 1D techniques, just as the solvent susceptibility
 $\tilde{\chi}_{\alpha\gamma}(k)$ itself. In the 1D case our deliberate
 choice of the self-inverse discrete sine transform of type IV (DST-IV)
-as a basic operation eliminates the need to treat the singularities
-at $r=0$ and $k=0$ as the discrete representation do not include
-these points by construction \citep{FFTW05}. The solute part is the
-electric form factor $\tilde{n}(\vec{k})$ of an arbitrary solute
-charge distribution $n(\vec{r})$ with a low $k$ expansion including
-all multipoles of the charge distribution; it is treated by regular
-3D techniques.
-
-A similar approach was applied before to a monopole field of the solute
-\citep{Kloss08b} and to a superposition of gaussian charge distributions
-\citep{Perkyns10}.  The argument line used above assumes the net
-neutrality of the solvent species and is not directly applicable to
-electrolytes, though even then the low $k$ behavior of a convolution
-like in Eq.~\ref{eq:tau} can be assumed to be regular, cp.\ the
-discussion of the background charge correction in Ref.~\citep{Kovalenko00b}.
-An interpretation of Eq.~\ref{eq:tau} as a linear response of the
-pure solvent distribution to a (weak) electrostatic field of the charge
-distribution $n(\vec{r})$ offers an intuitive way to judge the range
-assumptions for most solvent-solute combinations, cf.~Eqs.~\ref{eq:rsim-h-c-t}
-for sound closures that fulfill $f(x)\sim x$ upon linearization.
+as a basic operation eliminates the need for treating the singularities
+at $r=0$ and $k=0$ as the discrete representation does not include
+these points by construction \citep{FFTW05}. A similar approach was
+applied before to a monopole field of the solute \citep{Kloss08b}
+and to a superposition of Gaussian-type charge distributions \citep{Perkyns10}.
+
+An arbitrary solute charge distribution $n(\vec{r})$ gives rise to
+its electric form factor $\tilde{n}(\vec{k})$, the expansion of which
+at low $k$ includes all multipoles of the charge distribution. For
+a superposition of point charges $z_{i}$, representing solute sites
+at $\vec{r}_{i}$, the form factor is evaluated on the 3D $k$-grid
+using the analytic expression, $\tilde{n}(\vec{k})=\sum_{i}z_{i}\exp(-\mathrm{i}\vec{k}\vec{r}_{i})$,
+that is easy to generalize to other spherical distributions. For continuous
+charge distributions of zero net charge, the form factor is approximated
+by a fast Fourier transform of a periodically extended cell representation
+of $n(\vec{r})$ \citep{FFTW05}. A particle mesh Ewald method has
+recently been shown to accelerate significantly the construction of
+a form factor $\tilde{n}(\vec{k})$ and its convolution with a (renormalized)
+Coulomb kernel $\tilde{\varphi}(k)$ for 3D RISM calculations involving
+many charged solute sites \citep{Heil15}. The same technique should,
+in principle, be applicable to combinations of the form factor $\tilde{n}(\vec{k})$
+and site-specific response functions $\tilde{\theta}_{\alpha}(k)$
+without the need of a renormalization.
+
+The line of arguments used above assumes neutral solvent species and
+thus is not directly applicable to electrolytes. However, even then
+the behavior at low $k$ of a convolution like in Eq.~\ref{eq:tau}
+can be assumed to be regular; see the discussion of the background
+charge correction in Ref.~\citep{Kovalenko00b}. Indeed, a physically
+viable electrolyte model with a suitable susceptibility $\chi$ is
+intuitively expected to reproduce charge screening at finite Debye
+length. Then the convolution $h\sim-\chi\ast v_{\text{L}}$, interpreted
+here as a first-order response of the solvent distribution to an electrostatic
+field $v_{\text{L}}$, cf.~Eqs.~\ref{eq:rsim-h-c-t}, also decays
+at that same length scale \citep{Kovalenko00b}.
 
@@
 
-To model uranyl solvation we employ the SPC/E water model \citep{Berendsen87},
-modified as suggested by Pettitt and Rossky \citep{Pettitt82} by
-addition of a Lennard--Jones (LJ) term for hydrogen atoms with $\sigma_{\text{H}}=0.4$~Å
-and $\epsilon_{\text{H}}=0.046$~kcal/mol (PR-SPC/E). For atomic
-ions we adopted the coincident SPC/E water model (cSPC/E) with $\sigma_{\text{H}}=1.1658$~Å
-and $\epsilon_{\text{H}}=0.01553$~kcal/mol as proposed by Luchko
-and Kovalenko \citep{Luchko10} in order to compare with published
-results. The uranyl-water interaction parameters were derived from
-the second model proposed by Kerisit and Liu to strengthen the uranyl\textendash{}water
-interaction with the uranium charge of 3.5~\emph{e} (KL2) \citep{Kerisit13}.
-For convenience the reader may refer to our previous work \citep{Bo15}
-where we summarized the literal values for the charge and LJ parameters
-of several force-fields.
-
-All free energy calculations were carried out for the temperature
-of 298~K, with the dielectric constant set to 78.4 and the solvent
-particle density $\rho=0.0333295$~Å$^{-3}$. These parameters correspond
-to a pressure of 1~atm. The 3D domain was represented by a cubic
-cell of $20^{3}$~Å$^{3}$ and a Cartesian grid with 96$^{3}$ points
-and a resolution of about 0.208~Å. The radial grid used to precompute
-and interpolate the solvent susceptibility and its contractions with
-the Coulomb kernel for use in Eq.~\ref{eq:tau} was chosen with a
-radial range of 40~Å and 1536 points corresponding to a resolution
-of about 0.026~Å.
+As in our preceding study of 1D RISM, applied to uranyl \citep{Bo15},
+we modeled uranyl solvation using the SPC/E water model \citep{Berendsen87},
+modified by adding a Lennard--Jones (LJ) term for hydrogen atoms,
+with $\sigma_{\text{H}}=0.4$~Å and $\epsilon_{\text{H}}=0.046$~kcal/mol
+(PR-SPC/E) \citep{Pettitt82}. For benchmark calculations on mono-atomic
+alkali and halide ions \citep{Joung13,Bo15} we rather adopted the
+coincident SPC/E water model (cSPC/E) \citep{Luchko10}, with $\sigma_{\text{H}}=1.1658$~Å
+and $\epsilon_{\text{H}}=0.01553$~kcal/mol. The parameters characterizing
+the uranyl-water interaction were derived from the second model proposed
+by Kerisit and Liu (KL2) \citep{Kerisit13}. All charge values and
+the LJ parameters of several force-fields can be found in our previous
+work \citep{Bo15}.
+
+The free energy calculations were carried out for the temperature
+298~K, with the dielectric constant of bulk water set to 78.4, and
+the solvent particle density $\rho=0.0333295$~Å$^{-3}$. These parameters
+correspond to a pressure of 1~atm. The 3D domain was represented
+by a cubic cell of $20^{3}$~Å$^{3}$ and a Cartesian grid with 96$^{3}$
+points and a resolution of about 0.208~Å. The radial grid used to
+precompute and interpolate the solvent susceptibility and its contractions
+with the Coulomb kernel for use in Eq.~\ref{eq:tau} was chosen with
+a radial range of 40~Å and 1536 points corresponding to a resolution
+of about 0.026~Å. With these parameters, starting from scratch, a
+calculation of the solvent susceptibility, the solvation energy, and
+the corresponding forces for uranyl with six explicit water ligands
+takes about 40 seconds on 8 cores of two Intel$^{\circledR}$ Xeon$^{\circledR}$
+E5630 @ 2.53~GHz CPUs, where one achieves a parallel efficiency above
+40\%. The non-linear solver converges in 8 iterations with about 380
+Jacobian applications.
 
@@
 of Gaussian-type orbitals fitting-functions density functional method
-\citep{Dunlap90} as implemented in the parallel code ParaGauss \citep{ParaGaussV3.1}.
-The atomic approximation \citep{Matveev08} to the second-order Douglas\textendash{}Kroll\textendash{}Hess
-scalar-relativistic approach \citep{rev:roesch96} was used to account
-for relativistic effects; this approximation was shown to be very
-accurate for uranyl aqua complexes. We employed the gradient-corrected
+\citep{Dunlap90} as implemented in the parallel DFT code ParaGauss
+\citep{ParaGaussV3.2}.  The atomic approximation \citep{Matveev08}
+to the second-order Douglas\textendash{}Kroll\textendash{}Hess scalar-relativistic
+approach \citep{rev:roesch96} was used to account for relativistic
+effects; this approximation was shown to be very accurate for uranyl
+aqua complexes \citep{Matveev08}. We employed the gradient-corrected
 exchange-correlation functional (generalized gradient approximation,
@@
 
-We used a COSMO PCM solvation model for comparison \citep{Fuchs02},
+For comparison, we used a COSMO PCM (CPCM) solvation model \citep{Fuchs02},
 with the cavity defined via an effective solvent radius of 1.4~Å
@@
 
-Geometry optimizations have been carried out using the quasi-Newton
-method and the constrained minimization algorithm as implemented in
-the utility package ParaTools \citep{Chaffey-Millar11,Nikodem13}.
-In QM+RISM calculations of symmetric uranyl aqua complexes the U\textendash{}O
-bonds were relaxed but constrained to have the same value by symmetry.
-In the corresponding MM+RISM calculations uranyl was rigid with the
-U\textendash{}O distances fixed to 179~pm motivated by earlier results
-\citep{Moskaleva04,Cao05,Buehl05}. However, for a study of the water
-exchange with MM+RISM we adopted a flexible uranyl model without symmetry
-constraints and compared two sets of intra-molecular parameters \citep{Kerisit13,Tiwari14}
-with unconstrained QM+RISM results. The two uranyl force field models
-we employed differ only by uranyl equilibrium bond length and force
-constants for stretching and bending modes. The SPC/E geometry of
-explicit water ligands in the aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$,
+Molecular structures were optimized using the quasi-Newton method
+and the constrained minimization algorithm as implemented in the utility
+package ParaTools \citep{Chaffey-Millar11,Nikodem13}. In QM+RISM
+calculations of symmetric uranyl aqua complexes the U\textendash{}O
+bonds were relaxed, but constrained to have the same value by symmetry.
+In the corresponding MM+RISM calculations uranyl was kept rigid with
+the U\textendash{}O distances fixed to 179~pm, motivated by earlier
+results \citep{Moskaleva04,Cao05,Buehl05}. However, for a study of
+the water exchange with the MM+RISM approach, we adopted a flexible
+uranyl model without symmetry constraints and compared two sets of
+intra-molecular parameters \citep{Kerisit13,Tiwari14} with unconstrained
+QM+RISM results. The two uranyl force field models employed differ
+only by the uranyl equilibrium bond length and the force constants
+for stretching and bending modes. The SPC/E geometry of explicit water
+ligands in the aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$,
 $n=4-6$, features an O\textendash{}H distance of 100~pm and an H\textendash{}O\textendash{}H
@@
 
-To verify our 3D RISM implementation for charged solutes, we performed
-a comparison to 1D RISM results for mono-atomic alkali and halide
-ions \citep{Joung13,Bo15} applying the cSPC/E water model \citep{Berendsen87,Luchko10}.
-With moderate settings for the grid \textcolor{blue}{size} of $64^{3}$
-points in total and a cubic box volume of $20^{3}$~Å$^{3}$ we are
-able to match the 1D results up to \textcolor{blue}{0.2}~kcal/mol
-or better. \textcolor{blue}{With a 50\% larger box size at the same
-grid density the agreement improves to 0.07~kcal/mol; a further increase
-of the grid density by 50\% improves the results insignificantly.}
-The accuracy is slightly better for small cations than for the larger
-anions. The values are comparable to MD with SPC/E water and experimental
-results. The agreement may be slightly improved by a correction proportional
-to the molar volume of the solute \citep{Sergiievskyi14} as was demonstrated
-in our previous work \citep{Bo15}.
-
-\begin{table}[t]
+To verify our 3D RISM implementation for charged solutes, we compared
+to 1D RISM results for mono-atomic alkali and halide ions \citep{Joung13,Bo15}
+applying the cSPC/E water model \citep{Berendsen87,Luchko10}. With
+moderate settings for the grid resolution ($64^{3}$ points in total,
+a cubic volume of $20^{3}$~Å$^{3}$), we were able to match the
+1D results up to 0.2~kcal/mol or better (Table~\ref{tab:simple-ions}).
+With a 50\% larger box size at the same grid density the agreement
+improved to 0.07~kcal/mol; a subsequent increase of the grid density
+by 50\% improved the results only insignificantly. The values are
+comparable to MD results with SPC/E water and to experimental results
+(Table~\ref{tab:simple-ions}). The accuracy is slightly better for
+small cations than for the larger anions. The 10--15\% deviation from
+the experiment for anions may be reduced to 1--5\% by applying a correction
+proportional to the molar volume of the solute \citep{Sergiievskyi14}
+as was demonstrated in our previous work: for the largest anion, I$^{-}$,
+these corrections amount to $-10$~kcal/mol and $-8$~kcal/mol at
+KH and HNC levels, respectively \citep{Bo15}.
+
+\begin{table*}[tbph]
 \begin{centering}
-\caption{The solvation energy (excess chemical potential) for alkali and halide
-ions by 3D and 1D RISM approaches with the cSPC/E water model. MD
-and experimental values are also shown for comparison.$^{a}$ \label{tab:simple-ions}}
+\caption{Solvation energies (excess chemical potential) for alkali and halide
+ions in water by 3D and 1D RISM approaches with the cSPC/E model.
+MD and experimental values are also shown.$^{a}$ \label{tab:simple-ions}}
 
@@
 
-\small\vspace{1ex}
+\smallskip{}
+
+\small
 
@@
 \hline 
-KH & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.4 & $-$72.\textcolor{blue}{6} & $-$62.9\tabularnewline
-KH$^{b}$ & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.5 & $-$72.6 & $-$63.1\tabularnewline
-HNC & $-$111.\textcolor{blue}{7} & $-$86.3 & $-$69.3 & $-$79.8 & $-$74.1 & $-$64.7\tabularnewline
-HNC$^{b}$ & $-$111.7 & $-$86.3 & $-$69.3 & $-$79.9 & $-$74.2 & $-$64.9\tabularnewline
+3D KH & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.4 & $-$72.5 & $-$62.9\tabularnewline
+1D KH$^{b}$ & $-$108.5 & $-$85.0 & $-$68.5 & $-$78.5 & $-$72.6 & $-$63.1\tabularnewline
+3D HNC & $-$111.7 & $-$86.3 & $-$69.3 & $-$79.8 & $-$74.1 & $-$64.7\tabularnewline
+1D HNC$^{b}$ & $-$111.7 & $-$86.3 & $-$69.3 & $-$79.9 & $-$74.2 & $-$64.9\tabularnewline
 MD$^{c}$ & $-$113.3 & $-$88.4 & $-$71.0 & $-$89.3 & $-$82.7 & $-$74.4\tabularnewline
 Exp$^{d}$ & $-$113.8 & $-$88.7 & $-$71.2 & $-$89.1 & $-$82.7 & $-$74.3\tabularnewline
+\hline 
 \end{tabular}
@@
 
-\vspace{1ex}\footnotesize
-
-$^{a}$~Energies in kcal/mol. KH and HNC closure relations have been
-used. 3D domain is $20^{3}$ Å$^{3}$ / $64^{3}$ pts. $^{b}$~1D
-RISM from Ref.~\citep{Bo15}. $^{c}$~MD with SPC/E water from Ref.~\citep{Joung08}.
+\begin{raggedright}
+\smallskip{}
+
+\par\end{raggedright}
+
+\begin{raggedright}
+\footnotesize
+\par\end{raggedright}
+
+\raggedright{}$^{a}$~Energies in kcal/mol. KH and HNC closure relations
+have been used. 3D domain of $20^{3}$~Å$^{3}$, grid of $64^{3}$~points.
+$^{b}$~Ref.~\citep{Bo15}. $^{c}$~MD with SPC/E water; Ref.~\citep{Joung08}.
 $^{d}$~Ref.~\citep{Schmid00}. 
-\end{table}
+\end{table*}
 
@@
 
-In this section we examine a combination of 3D RISM for bulk water
-and a force field potential energy surface for the interactions of
-uranyl and $n$ explicit water ligands for $n=4-6$. The complex with
-$n=4$ features a missing equatorial ligand and demonstrates a qualitative
-difference of the optimal structures obtained by minimizing the MM
-and and MM+RISM expressions of the potential (free) energy. Similar
-to the findings in our previous work, using the 1D RISM formalism
-\citep{Bo15}, the MM+RISM combination features a vacancy in the equatorial
-plane of the uranyl corresponding to the location of the missing water
-ligand, Figure~\ref{fig:spatial-distribution-4w}. The MM potential
-energy surface features a minimum at a $D_{4\text{h}}$-like structure
-with four water ligands evenly distributed in the equatorial plane;
-structures featuring a vacancy in the equatorial plane are by 13--15~kcal/mol
-higher on that potential energy surface. The average U--OW bond in
-the MM-optimized structure of 239~pm is slightly shorter than the
-corresponding bonds of 240--242~pm of the structures optimized at
-the MM+RISM level, Table~\ref{tab:mm-rism}. The shorter distances
+Next we examine a combination of 3D RISM for bulk water and a force
+field potential for the interactions of uranyl and $n$ explicit water
+ligands, for $n=4-6$. The complex with $n=4$ features a missing
+equatorial ligand and demonstrates the qualitative difference of the
+optimal structures obtained by minimizing the MM and MM+RISM expressions
+of the (free) energy. Similar to the findings in our previous work,
+using the 1D RISM formalism \citep{Bo15}, the MM+RISM combination
+features a vacancy in the equatorial plane of the uranyl that corresponds
+to the location of the missing water ligand, Fig.~\ref{fig:spatial-distribution-4w}.
+A straightforward MM calculation yields a minimum at a $D_{4\text{h}}$-like
+structure with four water ligands evenly distributed in the equatorial
+plane; structures with a vacancy in the equatorial plane are by 13--15~kcal/mol
+higher on that potential energy surface. The average U--O$_{\text{w}}$
+bond, 239~pm, in the MM-optimized structure is slightly shorter than
+the corresponding bonds, 240--242~pm, of the structures optimized
+at the MM+RISM level, Table~\ref{tab:mm-rism}. The shorter distances
 indicate weaker steric ligand-ligand repulsions in the under-coordinated
-gas-phase complex. Note, however, there is a non-obvious opposite
-effect of shortening U--OW distances from 243--244~pm to 240~pm
-in five-coordinated uranyl complexes after embedding a gas phase complex
-into the RISM medium, Table~\ref{tab:mm-rism}. This could be an
-indication of a partial screening of steric repulsions of the equatorial
-ligands by the solvent of the second solvation shell in saturated
-complexes, also observed in PCM calculations, see below.
+complex in the gas phase. Note, however, the not so obvious opposite
+effect of shortened U--O$_{\text{w}}$ distances, from 243--244~pm
+to 240~pm, in five-coordinated uranyl complexes after embedding a
+complex with a structure optimized for the gas phase into the RISM
+medium, Table~\ref{tab:mm-rism}. This result may indicate a partial
+screening of the steric repulsion of the equatorial ligands of a saturated
+complex by the second solvation shell. This effect seems to be present
+also in PCM calculations, see below.
 
@@
 ligand is a local increase of the water density at the location of
-the missing explicit ligand, Figure~\ref{fig:spatial-distribution-4w}.
-Due to a notable width of the local distribution, both in equatorial
-and apical directions and less so in radial direction the RISM representation
-of the missing ligand is not an exact substitute for an explicit water
-ligand. Indeed, there is a measurable difference of the average U--OW
-distances between the complexes with 4 and 5 explicit waters in the
-first solvation shell: 241--242~pm for $n=4$ and 240~pm for $n=5$,
+the missing explicit ligand, Fig.~\ref{fig:spatial-distribution-4w}.
+Yet, the RISM representation of the missing ligand is not an exact
+substitute for an explicit water ligand, due to the notable width
+of the local distribution, both in equatorial and apical directions,
+but less so in radial direction. Indeed, the average distances U--O$_{\text{w}}$
+between the complexes with 4 and 5 explicit aqua ligands in the first
+solvation shell differ: 241--242~pm for $n=4$ and 240~pm for $n=5$,
 Table~\ref{tab:mm-rism}.
@@
 The strong localization of the solvent medium at the equatorial vacancy
-leads to divergence of the RISM equations with the PSE3 closure. This
-is manifested by the missing values in Table~\ref{tab:mm-rism}.
+induces a divergence of the RISM equations in combination with the
+PSE3 closure; see the missing values in Table~\ref{tab:mm-rism}.
 Results for the five-coordinated uranyl models, on the other hand,
-are essentially independent of the closure choice. The bare uranyl
+are essentially independent of the closure chosen. The bare uranyl
 model, where the axial symmetry makes the localization issue less
-severe, does converge for all three closures, though the results are
-significantly affected by the closure choice --- the solvation energy
-at the PSE3 level of theory is by 20~kcal/mol lower than at the PSE1
-level. A different source of errors is the approximation of the solute
-by a single structure minimizing the conformation free energy instead
-of an ensemble at thermal equilibrium. Whereas this could be justified
+severe, converges for each of the three closure relations, although
+the results vary significantly. The solvation energy at the PSE3 level
+is by 20~kcal/mol more negative than at the PSE1 level.
+
+A different source of errors is the approximation of the solute by
+a single structure, minimizing the conformation free energy instead
+of an ensemble at thermal equilibrium. This approximation can be justified
 to some extent for solutes with rigid intra-molecular and strong ligand
-bonds, the weakly associated water molecules of the second solvation
-shell need a better statistical treatment. The model with six explicit
-waters with one of them in the second shell ($n=5+1$, Table~\ref{tab:mm-rism})
-illustrates the point. Whereas the structure of the first solvation
-shell is very similar for $n=5$ and $n=5+1$ explicit waters, the
-solvation energy, computed by Eq.~\ref{eq:dG-sol-definition} is
-by about 5~kcal lower in the latter case. This is likely an artifact
-of a missing statistical sum over positions and orientations of the
-second shell water molecule. Hence a successful model shall treat
-the strongly bound aqua ligands explicitly as a part of the solute
-and weakly associated or free-floating water as a RISM medium in order
-to be numerically stable and closure independent.
-
-\begin{table}[t]
-\caption{Average uranyl--water distance of the first solvation shell, $d$(U--OW),
-and an estimate for the uranyl solvation energy, $\Delta G_{\text{sol}}$,
-obtained using gas phase MM (GP) and MM+RISM models with PSE closures
-of order $n=1-3$ (PSE$n$), see Eq.~\ref{eq:closure-pse}.$^{a}$
+bonds, but the weakly associated water molecules of the second solvation
+shell do require a statistical treatment. The model with six explicit
+water molecules illustrates the point; in that structure, one water
+molecule is in the second shell ($n=5+1$, Table~\ref{tab:mm-rism}).
+Whereas the structure of the first solvation shell is very similar
+for $n=5$ and $n=5+1$ explicit water molecules, the solvation energy,
+calculated by Eq.~\ref{eq:dG-sol-definition}, is by about 5~kcal
+higher (in absolute terms) in the latter case. This is likely an artifact
+of the missing statistical averaging over positions and orientations
+of the water molecule in the second shell. Hence, in order to yield
+numerically stable results, independent of the closure chosen, a successful
+model has to treat the strongly bound aqua ligands explicitly as a
+part of the solute and weakly associated or free-floating water molecules
+as part of the RISM medium.
+
+\begin{figure}[tbph]
+\caption{Water oxygen distribution around the uranyl aqua complex with four
+explicit water ligands. The isosurface of the O$_{\text{w}}$ distribution
+function, $g(\vec{r})=3$, is shown in red.  The area around the
+missing equatorial ligand is characterized by values of $g>20$. Other
+lobes formed by the isosurface, with $g<7$, represent water molecules
+in the second solvation shell. \label{fig:spatial-distribution-4w}
+}
+
+\smallskip{}
+
+\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/4w-composite}
+\end{figure}
+\begin{table}[tbph]
+\caption{Average uranyl--water distance, $d$(U--O$_{\text{w}}$), of the first
+solvation shell and estimate, $\Delta G_{\text{sol}}$, of the uranyl
+solvation energy obtained using gas phase MM (GP) and MM+RISM models
+with PSE closures of order $p=1-3$ (PSE$p$), see Eq.~\ref{eq:closure-pse}.$^{a}$
 \label{tab:mm-rism}}
-\small\vspace{1ex}
+
+\smallskip{}
+
+\small
 
@@
 \hline 
-$d$(U--OW) & GP & --- & 239 & 244 & 243\tabularnewline
+$d$(U--O$_{\text{w}}$) & GP & --- & 239 & 244 & 243\tabularnewline
  & PSE1 & --- & 241 & 240 & 240\tabularnewline
@@
  & PSE3 & $-$383 & na$^{d}$ & $-$392 & $-$397\tabularnewline
+\hline 
 \end{tabular}
@@
 
-\vspace{1ex}\footnotesize
-
-$^{a}$~Bond lengths in pm, energies in kcal/mol. $^{b}$~GP structure
-is $D_{4\text{h}}$ but the RISM structures feature a vacancy in the
-equatorial plane corresponding to the missing ligand, see Figure~\ref{fig:spatial-distribution-4w}.
-$^{c}$~Five-coordinated structure with the sixth water in the second
-shell. $^{d}$~Value not available due to divergence. $^{e}$~The
-GP row shows binding energies of $n$ water molecules to a rigid uranyl
-with $d$(U--O) of 179~pm. 
+\begin{raggedright}
+\smallskip{}
+\footnotesize
+\par\end{raggedright}
+
+\raggedright{}$^{a}$~Bond lengths in pm, energies in kcal/mol. $^{b}$~$D_{4\text{h}}$
+structure in the gas phase (GP), but the structures from RISM calculations
+feature a vacancy in the equatorial plane that corresponds to the
+missing ligand, see Fig.~\ref{fig:spatial-distribution-4w}. $^{c}$~Five-coordinated
+structure with the sixth water molecule in the second solvation shell.
+$^{d}$~Value not available due to divergence. $^{e}$~The GP results
+show binding energies of $n$ water molecules to a rigid uranyl moiety
+with $d$(U--O) = 179~pm. 
 \end{table}
-\begin{figure}[t]
-\caption{Water oxygen distribution around the uranyl aqua complex with four
-explicit water ligands. The isosurface corresponds to $h=2$. The
-area around the missing equatorial ligand is characterized by values
-of $h$ more than ten times as high. Other lopes of the isosurface
-correspond to second solvation shell and feature $h<6$. \label{fig:spatial-distribution-4w}}
-
-\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/4w}
-\end{figure}
 
@@
 
-We use symmetric models of uranyl aqua complexes to compare PCM and
-RISM representations of the solvent in QM calculations. The under-coordinated
-species does not feature a vacancy in the first solvation shell. Instead
-all water ligands are equivalent due to symmetry, also for the five-
-and six-coordinated uranyl. These models feature an additional intra-molecular
-degree of freedom, the uranyl bond length, in contrast to the MM models
-discussed in the previous section where the uranyl bond was fixed
-to a reasonable value.
-
-In our previous work exploring these models with 1D RISM \citep{Bo15}
-we observed consistently longer U--O and U--OW bond lengths in QM+RISM
-calculations in comparison to QM+PCM and experiment. We tentatively
-attributed that to the excess coordination of uranyl, manifested by
-the presence of a peak in the U--OW radial distribution at the distance
-corresponding to the first solvation shell for models with explicit
-ligands. This artifact is specific to 1D RISM and is not reproduced
-in 3D RISM calculations, cp.\ Figure~\ref{fig:rdfs-vs-closure-5w}
-showing no features at that separation and Figure~\ref{fig:rdfs-vs-closure-0w}.
-The number integrals of 5.3, 6.2, and 6.4 for the first peak of U--OW
-RDFs with PSE closures of order 1 to 3, respectively, for uranyl without
-explicit aqua ligands (Figure~\ref{fig:rdfs-vs-closure-0w}) are
-overestimated though.
-
-Apart from the bare uranyl model the QM+RISM and QM+PCM results for
-uranyl bond lengths of 179--180~pm agree in Table~\ref{tab:qm+rism-vs-qm+pcm}.
-For the bare uranyl model the QM+RISM result is qualitatively better.
-The QM+PCM predicts 174~pm for the uranyl bond length which is shorter
-than for uranyl with just four water ligands. The QM+RISM results,
-on the other hand, are around 179~pm being equally long (PSE1) or
-even marginally longer by 0.3~pm and 0.7~pm for (PSE2 and PSE3,
-respectively) than for four-coordinated uranyl (not visible in Table~\ref{tab:qm+rism-vs-qm+pcm}
-due to rounding).  The effect of the RISM medium on the bond length
-of the bare uranyl is thus comparable to that of the explicit water
-ligands; the elongation of the uranyl bond by 7--8~pm corresponds
-to 5--6~kcal/mol change in internal energy.
+In QM calculations, we use symmetric models of uranyl aqua complexes
+to compare PCM and RISM representations of the solvent in QM calculations.
+The under-coordinated species does not feature a vacancy in the first
+solvation shell. Instead, all water ligands are equivalent due to
+symmetry, also for the five- and six-coordinated uranyl. These models
+feature an additional intra-molecular degree of freedom, the uranyl
+bond length, in contrast to the MM models discussed in the previous
+section where the uranyl bond was fixed at a reasonable value.
+
+When previously exploring these models with 1D RISM \citep{Bo15},
+we observed consistently longer U--O and U--O$_{\text{w}}$ bond lengths
+in QM+RISM calculations as compared to QM+PCM and experiment. We tentatively
+attributed this trend to the excess coordination of uranyl, manifested
+by the presence of a peak in the U--O$_{\text{w}}$ RDF at the distance
+which corresponds to the first solvation shell for models with explicit
+ligands. From the present 3D RISM calculations, it is clear that this
+artifact is specific to 1D RISM. Comparing the U--O$_{\text{w}}$
+RDFs from the models without any explicit water ligands of uranyl
+(Fig.~\ref{fig:rdfs-vs-closure-0w}) and with five such water ligands
+(Fig.~\ref{fig:rdfs-vs-closure-5w}), one does not note any such
+feature at that crucial separation U--O$_{\text{w}}$. However, the
+number integrals, 5.3, 6.2, and 6.4, for the first peak of the U--O$_{\text{w}}$
+RDFs with PSE closures of orders 1 to 3, respectively, are overestimated
+in the model without explicit aqua ligands (Fig.~\ref{fig:rdfs-vs-closure-0w}).
+
+The QM+RISM and QM+PCM results agree for uranyl bond lengths, 179--180~pm,
+with the QM+PCM result, 174~pm, for the bare uranyl being the only
+less satisfactory exception (Table~\ref{tab:qm+rism-vs-qm+pcm}).
+On the other hand, the QM+RISM results are around 179~pm, hence equally
+long (PSE1) or even marginally longer by 0.3~pm and 0.7~pm for PSE2
+and PSE3, respectively, than for four-coordinated uranyl (not visible
+in Table~\ref{tab:qm+rism-vs-qm+pcm} due to rounding).  The effect
+of the RISM medium on the bond length of bare uranyl is thus comparable
+to that of the explicit water ligands; the elongation of the uranyl
+bond by 7--8~pm corresponds to 5--6~kcal/mol change in internal
+energy.
 
 Another notable discrepancy between the PCM and 1D RISM results mentioned
-in our previous work is also resolved by the 3D RISM approach: the
-uranyl--water distances U--OW are consistently shorter in solvated
-complexes than in the gas phase by measurable 4--6~pm. The 1D RISM
-approach did not show this effect convincingly \citep{Bo15}. Again,
-we interpret this effect of the medium as a screening of repulsive
-ligand-ligand interactions in the first coordination shell by the
-solvent.
-
-The magnitude of the uranyl solvation energy estimated with QM+RISM
-models is comparable with that obtained in our previous work using
-1D RISM. The estimates of $-$388 and $-$389~kcal/mol using the
-standard model of hydrated uranyl with five explicit water ligands
-vary only little with the closure order and agree well with $-$390~kcal/mol
+in our previous work \citep{Bo15} is also resolved by the 3D RISM
+approach, namely the uranyl--water distances U--O$_{\text{w}}$ are
+consistently shorter in solvated complexes than in the gas phase,
+by measurable 4--6~pm. The 1D RISM approach did not show this effect
+in a convincing manner \citep{Bo15}. Again, we interpret this effect
+of the medium as a screening of repulsive ligand-ligand interactions
+in the first coordination shell due to the solvent.
+
+The results for the solvation energy $\Delta G_{\text{sol}}$ of uranyl,
+estimated with QM+RISM models, are comparable to those obtained in
+our previous work using 1D RISM \citep{Bo15}. The estimates of $-$388~kcal/mol
+and $-$389~kcal/mol using the standard model of hydrated uranyl
+with five explicit water ligands vary only little with the order of
+the closure and agree very well with the result $-$390~kcal/mol
 from our previous work \citep{Bo15}. However, the relative order
 of the estimates based on the four- and six-coordinated complexes
-is not reproduced. The estimate of $\Delta G_{\text{sol}}$ based
-on the four-coordinated complex is by $\sim$10~kcal/mol less negative
-than the result of the standard model, Table~\ref{tab:qm+rism-vs-qm+pcm}.
-This difference (but not the absolute numbers) is the same as the
-PCM result, Table~\ref{tab:qm+rism-vs-qm+pcm}. However, this difference
-is significantly larger than the value of 1.6~kcal/mol obtained using
-1D RISM \citep{Bo15}, which is remarkably close to the free energy
-difference corresponding to the relative weights of four- and five-coordinated
-species determined from an average coordination number in a high-energy
-x-ray scattering (HEXS) experiment \citep{Soderholm05}. Given the
-limitations of 1D RISM exposed by the current 3D RISM approach we
-interpret that past agreement as fortuitous.  $\Delta G_{\text{sol}}$
-calculated from the six-coordinated $D_{3\text{d}}$ complex is by
-4--5~kcal/mol more negative than the result of the standard model
-with five explicit water ligands, Table~\ref{tab:qm+rism-vs-qm+pcm}.
-The corresponding PCM difference is only 1~kcal/mol and of opposite
-sign. It is likely that the limitations of the Eq.~\ref{eq:dG-sol-definition}
-with the free energy of the complex approximated from a single symmetry-constrained
-conformation is the source of this counter-intuitive ordering of estimates
-for $\Delta G_{\text{sol}}$. For similar reasons the MM+RISM estimates
-of the uranyl solvation energy using the models with different number
+is not reproduced. The estimate of $\Delta G_{\text{sol}}$ from the
+four-coordinated complex is by $\sim$10~kcal/mol less negative than
+the result of the standard model with five aqua ligands, Table~\ref{tab:qm+rism-vs-qm+pcm}.
+That difference of $\Delta G_{\text{sol}}$ values is the same as
+obtained from the PCM calculations, although the PCM values are by
+$\sim40$~kcal/mol more negative, Table~\ref{tab:qm+rism-vs-qm+pcm}.
+With 1D RISM the $\Delta G_{\text{sol}}$ values differ by 1.6~kcal/mol
+only \citep{Bo15}.
+
+The solvation energy $\Delta G_{\text{sol}}$ calculated from the
+six-coor\-dinated $D_{3\text{d}}$ complex is by 4--5~kcal/mol more
+negative than the result of the standard model with five explicit
+water ligands, Table~\ref{tab:qm+rism-vs-qm+pcm}. The corresponding
+PCM difference is only 1~kcal/mol and of opposite sign. The limitations
+of Eq.~\ref{eq:dG-sol-definition} with the free energy of the complex
+approximated from a single symmetry-constrained conformation likely
+is the source of this counter-intuitive ordering of estimates for
+$\Delta G_{\text{sol}}$. For similar reasons the MM+RISM estimates
+of the uranyl solvation energy using the models with different numbers
 of explicit water molecules agree between each other no better than
-to 3--5~kcal/mol even at the highest closure level, Table~\ref{tab:mm-rism}.
- Again, the solvation energy as determined by the QM+PCM model without
-explicit water ligands ($n=0$) is by $\sim$100~kcal/mol off the
-results obtained with $n=4-6$ explicit ligands, Table~\ref{tab:qm+rism-vs-qm+pcm}.
- The QM+PCM solvation energy of $-426$~kcal/mol given by the standard
-five-coordinated uranyl model appears to be too low \citep{Moskaleva04}
-in the view of most recent experimental estimate of $-369\pm15$~kcal/mol
-\citep{Gibson05}. The corresponding QM+RISM solvation energy of $-389$~kcal/mol
-is significantly closer to that interval. The properties in Table~\ref{tab:qm+rism-vs-qm+pcm}
-obtained using the models with explicit water ligands separating the
-uranyl di-cation and the RISM medium ($n=4-6$) hardly depend on the
-closure order. This observation may justify the widespread use of
-the lowest order PSE1 (KH) closure.
-
-\begin{table}[t]
-\caption{Uranyl bond length, $d$(U--O), uranyl--water distance, $d$(U--OW),
-and the uranyl solvation energy estimate, $\Delta G_{\text{sol}}$,
-obtained using gas phase QM, QM+PCM and QM+RISM models.$^{a}$ \label{tab:qm+rism-vs-qm+pcm}}
-
-\small\vspace{1ex}
+to 3--5~kcal/mol, even at the highest closure level, Table~\ref{tab:mm-rism}.
+The solvation energy as determined by the QM+PCM model without explicit
+water ligands ($n=0$) is again by $\sim$100~kcal/mol off the results
+obtained with $n=4-6$ explicit ligands, Table~\ref{tab:qm+rism-vs-qm+pcm}.
+ The QM+PCM solvation energy of $-426$~kcal/mol, obtained with
+the standard five-coordinated uranyl model \citep{Moskaleva04}, appears
+to be too high in absolute value in view of the most recent experimental
+estimate, $-369\pm15$~kcal/mol \citep{Gibson05}. The corresponding
+QM+RISM solvation energy, $-389$~kcal/mol, is significantly closer
+to the interval estimated in experiment. The properties in Table~\ref{tab:qm+rism-vs-qm+pcm}
+obtained using the models with explicit aqua ligands, that separate
+the uranyl di-cation and the RISM medium ($n=4-6$), hardly depend
+on the closure order. This observation may justify the widespread
+use of the lowest-order closure PSE1 (KH).
+
+\begin{table}[tbph]
+\caption{Uranyl bond length, $d$(U--O), uranyl--water distance, $d$(U--O$_{\text{w}}$),
+ and estimate, $\Delta G_{\text{sol}}$, of the uranyl solvation
+energy, as obtained using a QM model for a system in the gas phase
+(GP) as well as QM+PCM and QM+RISM models.$^{a}$ \label{tab:qm+rism-vs-qm+pcm}}
+
+\smallskip{}
+
+\small
 
@@
  & PSE3 & 179 & 179 & 179 & 180\tabularnewline
-$d$(U--OW) & GP & --- & 242 & 249 & 253\tabularnewline
+$d$(U--O$_{\text{w}}$) & GP & --- & 242 & 249 & 253\tabularnewline
  & PCM & --- & 237 & 243 & 249\tabularnewline
@@
  & PSE3 & $-$378 & $-$379 & $-$389 & $-$393\tabularnewline
+\hline 
 \end{tabular}
@@
 
-\vspace{1ex}\footnotesize
-
-$^{a}$~Bond lengths in pm, energies in kcal/mol. Symmetric uranyl
-aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$, $4\leq n\leq6$
-and a bare uranyl ($n=0$) are treated by QM while interactions with
-the PR-SPC/E water solvent are treated by KL2 force field. $^{b}$~$D_{4\text{h}}$
-structure. $^{c}$~$D_{5\text{h}}$ structure. $^{d}$~$D_{3\text{d}}$
-structure. $^{e}$~The row shows binding energies of $n$ water ligands. 
+\begin{raggedright}
+\smallskip{}
+\footnotesize
+\par\end{raggedright}
+
+\raggedright{}$^{a}$~Bond lengths in pm, energies in kcal/mol. Symmetric
+uranyl aqua complexes {[}UO$_{2}$(H$_{2}$O)$_{n}${]}$^{2+}$, $4\leq n\leq6$
+and the bare uranyl ($n=0$) are treated by QM while interactions
+with the PR-SPC/E water solvent are treated by the KL2 force field.
+$^{b}$~$D_{4\text{h}}$ structure. $^{c}$~$D_{5\text{h}}$ structure.
+$^{d}$~$D_{3\text{d}}$ structure. $^{e}$~Binding energies of
+$n$ aqua ligands. 
 \end{table}
 
-\begin{figure}[t]
-\caption{Radial distribution functions of the solvated uranyl without explicit
-aqua ligands: U--OW (top panel, black), U--HW (top panel, gray), O--OW
-(bottom panel, black) and O--HW (bottom panel, gray). Results employing
-PSE1 closure are shown with the thick lines in the foreground. Results
-employing PSE2 and PSE3 closures are displayed using thinner lines
-in the background.  \label{fig:rdfs-vs-closure-0w}}
-
-\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig2}
+\begin{figure}[tbph]
+\caption{Radial distribution functions of solvated uranyl without explicit
+aqua ligands: U--O$_{\text{w}}$ (top panel, black), U--H$_{\text{w}}$
+(top panel, gray), O--O$_{\text{w}}$ (bottom panel, black) and O--H$\text{\ensuremath{_{w}}}$
+(bottom panel, gray). Results obtained with the PSE1 closure are shown
+as thick lines in the foreground, results obtained with the closure
+relations PSE2 and PSE3 are displayed in the background using thinner
+lines.  \label{fig:rdfs-vs-closure-0w}}
+
+\smallskip{}
+
+\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig2}
 \end{figure}
 
-\begin{figure}[t]
+\begin{figure}[tbph]
 \caption{Radial distribution functions of the solvated uranyl complex with
-five explicit water ligands in the equatorial plane. See Fig.~\ref{fig:rdfs-vs-closure-0w}
-for explanation. \label{fig:rdfs-vs-closure-5w}}
-
-\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig3}
+five explicit aqua ligands in the equatorial plane. Same lay-out as
+Fig.~\ref{fig:rdfs-vs-closure-0w}. \label{fig:rdfs-vs-closure-5w}}
+
+\smallskip{}
+
+\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig3}
 \end{figure}
@@
 
-By way of the Eyring equation the activation energy of aqua ligand
-exchange process in the first solvation shell of uranyl computed from
-the experimentally available water exchange rate estimate of 1.3$\times$10$^{6}$~s$^{-1}$
-is about 9~kcal/mol \citep{Farkas00}. Many theoretical studies predict
-much faster water exchange than the available experimental data suggests
-\citep{Vallet01,hagberg05,Frick09,Nichols08,Kerisit10,Tiwari14}.
-A simple way to obtain an estimate of the activation energy for the
-water exchange may be the inspection of the U--OW potential of mean
-force (PMF), $W(r)=-T\ln g(r)$, derived from the uranyl--water oxygen
-RDF $g(r)$, Fig.\ \ref{fig:potential-of-mean-force}. In Table~\ref{tab:pmf-extrema}
-we give the locations of the extrema of such PMF profiles, obtained
-using the bare uranyl model and three different closure relations.
-The activation energy estimated by 3D RISM as the PMF difference of
-the first-shell minimum and the barrier top ranges from 3.4 to 5.2
-kcal/mol, increasing with the closure order. The corresponding number
-from the MD RDF is 4.4 kcal/mol. These values are comparable to the
-5~kcal/mol derived from the residence time by Tiwari et al.~\citep{Tiwari14},
-but significantly lower than the experimental estimate. The transition
-between the first- and second solvation shells happens at U--OW separations
-$r_{\text{b}}=309-324$~pm according to the radial PMF profiles.
-Note that the one-particle radial PMF $W(r)$ provides no information
-on the collective behavior of the water ligands during the water exchange,
-thus it offers only limited insight into possible mechanisms. Also,
-the RISM model for bare uranyl assumes a rigid linear geometry whereas
-it was shown that uranyl bending essentially facilitates the water
-exchange \citep{Tiwari14}.
-
-Additional insight is provided by a three-dimensional PMF $W(\vec{r})=-T\ln g(\vec{r})$
-given as $W(x,y)$ in Fig.\ \ref{fig:potential-of-mean-force-2d}
-on a quarter of the axial plane with the uranium atom at the origin
-and the uranyl oxygen on the $y$-axis. The location of the saddle
-point  in Fig.\ \ref{fig:potential-of-mean-force-2d} corresponds
-to a U--OW distance of 318~pm comparable with the respective PSE1
-value in Table~\ref{tab:pmf-extrema}. The angle of attack for the
-incoming/leaving water ligand estimated from that location is about
-60$^{\circ}$ to the uranyl axis. This value is in good agreement
-with the result of 61.8$^{\circ}$ from the transition state search
-at Hartree--Fock level of theory with water solvent described by CPCM
-model \citep{Vallet01}. A hypothetical water exchange in the equatorial
-plane appears to be significantly hindered at room temperature. The
-PMF values at the location of the saddle point and at the equatorial
-minimum are 1.4~kcal/mol and $-1.7$~kcal/mol, respectively (PSE1).
-Thus the activation barrier is about 3.1~kcal/mol which is only slightly
-less than the estimate from the radial PMF, Table~\ref{tab:pmf-extrema}.
-Again it is not possible to draw a conclusion on a collective behavior
-of the incoming and leaving ligands from this one-particle property.
-
-\begin{figure}[t]
-\caption{Potential of mean force, $W(r)$, for U--OW distribution functions
-obtained with PSE1, PSE2 and PSE3 closures (solid lines of decreasing
-weight). The dashed line represents the MD result. \label{fig:potential-of-mean-force}}
-
-\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf}
+The activation energy for exchanging an aqua ligand of uranyl and
+a water molecule from the surrounding bulk is estimated at about 9~kcal/mol
+when one applies the Eyring equation to the experimentally available
+water exchange rate, 1.3$\times$10$^{6}$~s$^{-1}$ \citep{Farkas00}.
+ Many theoretical studies \citep{Vallet01,hagberg05,Frick09,Nichols08,Kerisit10,Tiwari14}
+predict a water exchange much faster than the available experimental
+data suggests.  As the experimentally observed exchange rate is many
+orders of magnitude lower than a diffusion-controlled reaction rate
+$4\pi\rho DR$ for a solvent molecule characterized by self-diffusion
+coefficient $D$ and a solute approximated as spherical particle of
+radius $R$ (uranyl, $R=3$\textasciitilde{}Å), the rate limiting
+step of the reaction must be an exchange of a uranyl ligand and a
+solvent molecule from the second solvation shell.  Such an exchange
+involves breaking a uranyl--ligand bond for the molecule leaving the
+first solvation shell and breaking a network of hydrogen bonds for
+the incoming water molecule; bonds are restored as a result of exchange.
+Below, we will focus on water exchange between the first and second
+solvation shells, without assuming a specific mechanism of this reaction.
+However, given that a six-coordinated uranyl species was determined
+to be significantly more stable in the gas phase than a four-coordinated
+species, Table~\ref{tab:qm+rism-vs-qm+pcm}, we anticipate an associative
+mechanism via the six-coordinated intermediate to be preferred to
+a dissociative mechanism via the four-coordinated species. Furthermore,
+the equivalence of the solvent molecules participating in the exchange
+allows us to limit the attention to a half-reaction with the intermediate
+as final state.
+
+Inspection of the U--O$_{\text{w}}$ potential of mean force (PMF),
+$W(r)=-T\ln g(r)$, derived from the uranyl--water oxygen RDF $g(r)$,
+offers a simple way for estimating the activation energy of water
+exchange (Fig.\ \ref{fig:potential-of-mean-force}). Table~\ref{tab:pmf-extrema}
+collects the extrema of such PMF profiles, obtained using the bare
+uranyl model and three closure relations. The activation energy estimated
+by 3D RISM as the PMF difference of the first-shell minimum and the
+barrier top ranges from 3.4~kcal/mol to 5.2~kcal/mol, increasing
+with the order of the closure. The corresponding value from the MD
+RDF is 4.4~kcal/mol \citep{Bo15}.  These values are comparable
+to the result of 5~kcal/mol, derived from the residence time in MD
+simulation \citep{Tiwari14}, but they are significantly lower than
+the experimental estimate, 9~kcal/mol \citep{Farkas00}. The transition
+between the first and the second solvation shells occurs at U--O$_{\text{w}}$
+separations $r_{\text{b}}=309-324$~pm, according to the radial PMF
+profiles. Note that the one-particle radial PMF $W(r)$ does not provide
+any information on the collective behavior of the water ligands during
+the water exchange. Thus it offers only limited insight into a possible
+mechanism. Also, the RISM model for bare uranyl assumes a rigid linear
+geometry whereas it was shown that uranyl bending significantly facilitates
+the exchange of water ligands \citep{Tiwari14}.
+
+Additional insight can be gained from a three-dimensional PMF $W(\vec{r})=-T\ln g(\vec{r})$
+(Fig.~\ref{fig:potential-of-mean-force-2d}), given as $W(\varrho,z)$
+in a quarter of the axial plane with the uranium atom at the origin
+and one of the uranyl terminal oxygen centers on the $z$-axis. The
+location of the saddle point  in Fig.~\ref{fig:potential-of-mean-force-2d}
+corresponds to a U--O$_{\text{w}}$ distance of 318~pm, comparable
+with the PSE1 value in Table~\ref{tab:pmf-extrema}. The angle of
+attack for the incoming or leaving aqua ligand, estimated from that
+location, is about 60$^{\circ}$ to the uranyl axis. This value is
+in good agreement with the result of 61.8$^{\circ}$ from a transition
+state search at the Hartree--Fock level with the water solvent described
+by the CPCM model \citep{Vallet01}. A hypothetical water exchange
+in the equatorial plane appears to be significantly hindered at room
+temperature. The PMF values at the location of the saddle point and
+at the equatorial minimum are determined at 1.4~kcal/mol and $-1.7$~kcal/mol,
+respectively (PSE1). Thus the activation barrier is about 3.1~kcal/mol
+which is only slightly less than the estimate from the radial PMF,
+Table~\ref{tab:pmf-extrema}. Both types of analysis, based on a
+radial or a three-dimensional PMF, carried out for a rigid uranyl
+model using force field interactions, thus predict too low a value
+of the activation energy in comparison to experiment, but they offer
+valuable qualitative insight into possible reaction pathways. Of course,
+a radical reduction of the dimensionality of the free energy surface
+to a few degrees of freedom will likely obscure mechanistic details.
+For example, the (one-dimensional) PMF profile does not indicate any
+metastable intermediate. It is not possible to draw any conclusion
+regarding the collective behavior of the incoming and leaving ligands
+from one-particle PMF profiles.
+
+\begin{figure}[tbph]
+\begin{centering}
+\caption{Potential of mean force, $W(r)$, from the U--O$_{\text{w}}$ radial
+distribution function, determined with the closure relations PSE1,
+PSE2 and PSE3 (solid lines of decreasing weight). The dashed line
+represents the MD result of Ref.~\citep{Bo15}\textcolor{blue}{.}
+\label{fig:potential-of-mean-force}}
+
+\par\end{centering}
+
+\smallskip{}
+
+\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf}
 \end{figure}
 
-\begin{table}[t]
+\begin{table*}[tbph]
 \begin{centering}
-\caption{PMF extrema $\{r_{i},W(r_{i})\}$,$^{a}$ and an estimate for the
-water exchange activation energy, $\Delta W_{\text{ab}}=W(r_{\text{b}})-W(r_{\text{a}})$
-extracted from the curves in Fig.\ \ref{fig:potential-of-mean-force}.
-\label{tab:pmf-extrema}}
+\caption{Extrema $\{r_{i},W(r_{i})\}$$^{a}$ in the potential of mean force
+and estimates of the activation energy for exchange of aqua ligands
+of uranyl, $\Delta W_{\text{ab}}=W(r_{\text{b}})-W(r_{\text{a}})$,
+between the first and the second solvation shell, extracted from the
+curves in Fig.~\ref{fig:potential-of-mean-force}. \label{tab:pmf-extrema}}
 
@@
 MD & 246 & $-$1.5 & 297 & 2.9 & 460 & $-$0.3 & 4.4\tabularnewline
+\hline 
 \end{tabular}
@@
 
-\vspace{1ex}\footnotesize
-
-$^{a}$~Distances in pm, energies in kcal/mol. Locations $r_{\text{a}}$
-and $r_{\text{c}}$ correspond to the minima of first- and second
-solvation shells, respectively; $r_{\text{b}}$ is the highest point
-of the barrier connecting the two minima. 
-\end{table}
-
-\begin{figure}[t]
-\caption{Potential of mean force, $W(x,y)$, for the U--OW distribution functions
-obtained with PSE1 closure. Here $x$ and $y$ are the radial and
-axial displacements off the uranium atom. \label{fig:potential-of-mean-force-2d}}
-
-\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf-2d}
+\begin{raggedright}
+\smallskip{}
+\footnotesize
+\par\end{raggedright}
+
+\raggedright{}$^{a}$~Distances in pm, energies in kcal/mol. Locations
+$r_{\text{a}}$ and $r_{\text{c}}$ correspond to the minima of the
+first and the second solvation shells, respectively; $r_{\text{b}}$
+marks the highest point of the PMF curve between the two minima. 
+\end{table*}
+
+\begin{figure}[tbph]
+\caption{Potential of mean\textcolor{blue}{{} }force, $W(\varrho,z)$, for the
+U--O$_{\text{w}}$ distribution function, determined with the closure
+PSE1. $\varrho$ and $z$ are the radial and axial displacements from
+the uranium center. \label{fig:potential-of-mean-force-2d}}
+
+\smallskip{}
+
+\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf-2d}
 \end{figure}
 
-\begin{figure}[t]
+\begin{figure}[tbph]
 \caption{Free energy profiles, $G(q)$, along the water exchange pathway using
-two intra-molecular force fields with soft- and hard uranyl bending
-modes (squares and diamonds, respectively). The QM+RISM profile is
-represented by three stationary states (circles) connected by strait
-lines to guide the eye. The reaction coordinate $q$ is the difference
-of U--OW distances of two designated water molecules. \label{fig:free-energy-profiles}}
-
+uranyl force fields with soft (squares) and hard (diamonds) uranyl
+bending modes. The QM+RISM profile is represented by three stationary
+states (circles) connected by strait lines to guide the eye. The reaction
+coordinate $q$ is the difference of U--O$_{\text{w}}$ distances
+of two selected water molecules (see text). \label{fig:free-energy-profiles}
+}
+
+\centering{}\smallskip{}
 \includegraphics[width=1\linewidth]{uo2-II\lyxdot d/profiles}
@@
 
+\begin{table}[tbph]
+
+\caption{Free energy values $G(q)$ at local minima $q=q^{\ast}$, corresponding
+to a five-coordinated uranyl complex with an explicit water molecule
+in the second shell, and $q=0$, representing a six-coordinated metastable
+intermediate, as well as at the corresponding transition state, $q=q^{\#}$,
+obtained by MM+RISM and QM+RISM approaches, see Fig.~\ref{fig:free-energy-profiles}.$^{a}$
+\label{tab:stationary-points}}
+\smallskip{}
+
+\small
+
+\begin{centering}
+\begin{tabular}{lrrrr}
+\hline 
+ & \multicolumn{2}{c}{MM+RISM} &  & QM+RISM\tabularnewline
+\cline{2-3} \cline{5-5} 
+ & soft & hard &  & \tabularnewline
+\hline 
+$G(q^{\ast})$ & -426.0 & -430.8 &  & -434.4\tabularnewline
+$G(q^{\#})$  & -423.7 & -426.5 &  & -425.2\tabularnewline
+$G(0)$ & -426.1 & -427.2 &  & -428.7\tabularnewline
+$\Delta G^{\#}$$^{b}$ & 2.3 & 4.3 &  & 9.1\tabularnewline
+\end{tabular}
+\par\end{centering}
+
+\begin{raggedright}
+\smallskip{}
+\footnotesize
+\par\end{raggedright}
+
+\raggedright{}$^{a}$~Free energies in kcal/mol. The exchange reaction
+corresponds to a transition from $q^{\ast}$ to $-q^{\ast}$, but
+due to symmetry $G(q^{\ast})=G(-q^{\ast})$ and $G(q^{\#})=G(-q^{\#})$.
+$^{b}$~The activation free energy is estimated as $\Delta G^{\#}=G(q^{\#})-G(q^{\ast})$.
+\end{table}
+
 To investigate the collective behavior of the ligands in the water
-exchange process we need to treat the first solvation shell and the
-incoming water molecule explicitly and also employ a flexible uranyl
-model as it was shown that the uranyl bending may affect the exchange
-rate \citep{Tiwari14}. The two intra-molecular force fields for a
-flexible uranyl species we compare below differ primarily by the force
-constant corresponding to the bending modes of uranyl and the equilibrium
-bond length $d_{0}$(U--O). The force field here referred to as ``soft''
-corresponds to the frequency of the bending mode of 159~cm$^{-1}$
-and $d_{0}$(U--O) of 176~pm \citep{Tiwari14}. The alternative ``hard''
-force field features a significantly higher frequency of the uranyl
-bending mode of 393~cm$^{-1}$ and a somewhat too long a uranyl bond
- $d_{0}$(U--O) of 180~pm \citep{Kerisit13}. The bending vibration
-in crystals and solution has been experimentally characterized by
-frequencies in the range of 180--270~cm$^{-1}$ \citep{Denning76,Denning79,Toth81,Gal92}.
-The DFT models with 5--6 explicit water ligands that we employ in
-this and earlier works predict the bending frequency in this experimental
-range \citep{Matveev08}. For the gas phase uranyl the standard DFT
-predicts too soft a bending mode likely because of the self-interaction
-error for $f$-electrons \citep{Ramakrishnan11} while hybrid DFT
-and CCSD(T) estimate the bending frequency around 160--180~cm$^{-1}$
-\citep{Jackson08}. In this light the force field with the ``soft''
-bending mode appears to be a more reliable choice.
+exchange process, one needs an explicit treatment of the first solvation
+shell and the incoming water molecule; also, one needs to employ a
+flexible uranyl moiety (see above) \citep{Tiwari14}. To describe
+a flexible uranyl species, we employed two intra-molecular force fields
+that primarily differ by the force constant of the uranyl bending
+modes and the equilibrium bond length $d_{0}$(U--O). The force field,
+referred to as ``soft'' \citep{Tiwari14}, yields $d_{0}$(U--O)
+= 176~pm and the harmonic frequency of the bending mode is 159~cm$^{-1}$.
+The alternative ``hard'' force field \citep{Kerisit13} features
+a uranyl bond $d_{0}$(U--O) = 180~pm, which is somewhat too long,
+and a significantly higher frequency, 393~cm$^{-1}$, of the uranyl
+bending mode. The bending vibrations in solution and in crystals 
+have experimentally been characterized by frequencies in the range
+of 180--270~cm$^{-1}$ \citep{Denning76,Denning79,Toth81,Gal92}.
+The DFT models at the BP level, with 5--6 explicit water ligands,
+employed in the present and in earlier works, predict the bending
+frequency in this experimental range \citep{Matveev08}. For uranyl
+species in the gas phase, such standard DFT calculations with semi-local
+exchange-correlation functionals, e.g.\ BP, predict too soft a bending
+mode slightly below 100~cm$^{-1}$,  likely because of the self-interaction
+error associated with the $f$-electrons \citep{Ramakrishnan11}.
+In contrast, hybrid DFT and CCSD(T) methods estimate the bending frequency
+at around 160--180~cm$^{-1}$ \citep{Jackson08}. In this light the
+force field with the ``soft'' bending mode appears to be a more
+adequate choice.
 
 However, a comparison of the free energy profiles obtained with QM
-and the two force field models of uranyl aqua complexes in RISM medium
-reveals significant quantitative deficiencies for both force field
-models. The Fig.~\ref{fig:free-energy-profiles} shows the three
-free energy profiles along the reaction coordinate, $q=d(\text{U--OW}_{1})-d(\text{U--OW}_{2})$,
-which is a difference of two uranium--water oxygen distances for incoming
-and leaving water molecules. The QM+RISM profile is approximated by
-just three stationary points. All of the three profiles feature an
-intermediate metastable state at around $q=0$ roughly corresponding
-to the six-coordinated $D_{3\text{d}}$-like uranyl aqua complex.
-With the ``soft'' force field the uranyl species in the metastable
-state is bent featuring an O--U--O angle of 155$^{\circ}$ and distances
-to the two water ligands participating in exchange which are by about
-15~pm longer than the distances to the other four water ligands;
-the deviations from the $D_{3\text{d}}$ geometry for the other two
-models, QM and ``hard'' force field, are less pronounced.  The
-transition state approximated by the structure with the highest free
-energy is characterized by $q>0.6$~Å for both the QM and the ``soft''
-uranyl models with the longest U--OW distance of 303~pm and 310~pm,
-respectively. These distances are comparable to or slightly below
-the estimates derived from the 1D and 3D PMFs above. The ``hard''
-uranyl model, on the other hand, suggests a transition state with
-a much shorter distance to the outermost water ligand of about 280~pm
-and a shallow metastable state. This is likely due to the significantly
-longer U--O bond of the solvated uranyl species of about 186~pm which
-is more than 6~pm longer than those of the other models employed
-here and in other works \citep{Tiwari14,Kerisit13} and should be
-considered a deficiency of the parametrization.  In turn, the ``soft''
-model does not convincingly show the preference for five-coordinated
-uranyl with the sixth water in the second solvation shell over the
-six-coordinated broken symmetry metastable structure. The QM and ``hard''
-force field models predict the six-coordinated $D_{3\text{d}}$-like
-state by 5.6~kcal/mol and 3.6~kcal/mol less stable than the five-coordinated
-structure with an extra water in the second shell, respectively. The
-presence of such a six-coordinated intermediate suggests an associative
-two-step water exchange mechanism as one of the possibilities. This
-is consistent with the findings of the other theoretical studies \citep{Tsushima07,Vallet01,Buehl06b,Rotzinger07,Tiwari14}
-suggesting that the associative mechanism is more favorable than the
-dissociative mechanism.
+calculations and the two force field models of uranyl aqua complexes
+in a RISM medium reveals significant quantitative deficiencies of
+both force field models. Fig.~\ref{fig:free-energy-profiles} shows
+the three free energy profiles along the reaction coordinate, $q=d(\text{U--O}_{\text{w1}})-d(\text{U--O}_{\text{w2}})$
+where the labels ``1'' and ``2'' refer to the incoming and leaving
+water molecules, respectively. Small values of $q$ correspond to
+a metastable six-coordinated complex and large absolute values of
+$q$ correspond to a five-coordinated uranyl complex with one ($q>0$)
+or the other ($q<0$) explicit water molecule in the second shell.
+The full profile is symmetric around $q=0$. Due to the apparent symmetry
+of the metastable state and the exchange process as the whole, only
+the non-redundant half of the actual reaction profile is displayed
+in Fig.~\ref{fig:free-energy-profiles}. The QM+RISM profile is approximated
+by just three stationary points: two local minima and a saddle point
+obtained as root of the Lagrange multiplier in a series of constrained
+optimizations. The reaction path was sampled in a preparation step
+using lower quality settings for the QM and RISM energy terms. All
+of the three profiles feature an intermediate metastable state at
+around $q=0$, roughly corresponding to the six-coordinated $D_{3\text{d}}$-like
+uranyl aqua complex. With the ``soft'' force field, the uranyl species
+in this metastable state is bent, featuring an O--U--O angle of 155$^{\circ}$
+and distances to the two water ligands participating in the exchange
+which are by about 15~pm longer than the distances to the other four
+water ligands. The deviations from the $D_{3\text{d}}$ geometry are
+less pronounced for the other two models, QM and ``hard'' force
+field.
+
+The transition state approximated by the structure with the highest
+free energy is characterized by $q>0.6$~Å for both the QM and the
+``soft'' uranyl models, with the longest U--O$_{\text{w}}$ distances
+at 303~pm and 310~pm, respectively. These distances are comparable
+to or slightly below the estimates derived above from the 1D and 3D
+PMFs. The ``hard'' uranyl model, on the other hand, suggests a transition
+state with a much shorter distance to the outermost water ligand of
+\textasciitilde{}280~pm and a shallow metastable state. This is likely
+a consequence of the significantly longer U--O bond of the solvated
+uranyl species, \textasciitilde{}186~pm, which is more than 6~pm
+longer than the corresponding distances of the other models employed
+here and in other works \citep{Tiwari14,Kerisit13}. This particular
+elongation should be considered as deficiency of the parametrization.
+ In turn, the ``soft'' model does not convincingly show the preference
+for five-coordinated uranyl with the sixth water in the second solvation
+shell over the six-coordinated broken symmetry metastable structure.
+The QM model and ``hard'' force field model predict the six-coordinated
+$D_{3\text{d}}$-like state by 5.6~kcal/mol and 3.6~kcal/mol, respectively
+(Table~\ref{tab:stationary-points}), less stable than the five-coordinated
+structure with an extra water in the second shell. The presence of
+such a six-coordinated intermediate may be taken to suggest an associative
+two-step water exchange mechanism. This latter finding is consistent
+with the results of other theoretical studies \citep{Tsushima07,Vallet01,Buehl06b,Rotzinger07,Tiwari14}
+that suggest an associative mechanism as more favorable than the dissociative
+mechanism via the four-coordinated intermediate structure.
 
@@
 and can be described as a redistribution in the torus-shaped first
-solvation shell roughly corresponding to the low PMF region, cp.~Figs.~\ref{fig:potential-of-mean-force}
-and \ref{fig:potential-of-mean-force-2d}, while the sixth water molecule
-enters or leaves the first solvation shell. The water exchange barrier
-of 2.3 and 4.4~kcal/mol calculated with the ``soft'' and ``hard''
-force field models, respectively, though not very far from the estimates
-we obtained above from the PMF analysis, are significantly lower than
-the 9.1~kcal/mol barrier of the QM+RISM model, Fig.~\ref{fig:free-energy-profiles},
-which is in good agreement with the experimental estimate of 9~kcal/mol
-\citep{Farkas00}.  A comparison of locations and orientations of
-the second-shell explicit water which is connected by hydrogen bonds
-to the two explicit aqua ligands of uranyl between the three models
-suggests that the force field models may slightly underestimate the
-strength and directional nature of the hydrogen bonds. This may explain
-the low water exchange activation energy and the stability of the
-six-coordinated metastable state relative to the structure with an
-explicit water in the second shell, as predicted by the force field
-models, Fig.~\ref{fig:free-energy-profiles}.
-
-It should be noted that water activation energy barrier computed by
-state of the art QM is sensitive to the choice of reaction path, level
-of theory, and solvation model \citep{Tiwari14}. Tsushima \citep{Tsushima07}
-reported the activation energy barriers of 10.1~kcal/mol and 4.8~kcal/mol
-at the B3LYP level with the conductor-like polarizable continuum model
-(CPCM) by considering the terminal states with one and two hydrogen
-bonds, respectively. The latter value is close to the prediction of
-4.5~kcal/mol in an earlier study \citep{Vallet01}, yet both are
-lower than the experimental result. The comparison of PCM and a spherical
-cavity self-consistent reaction field (SCRF) solvation models is reported
-by Rotzinger \citep{Rotzinger07} using the complete active space
-self-consistent field (CAS SCF) technique. The corresponding energy
-barriers of the associative pathway are equal to 9.4~kcal/mol and
-6.5~kcal/mol, respectively. Water exchange energy barrier based on
-the thermodynamic sampling is also estimated by Bühl et.\ al \citep{Buehl06b}
-using the \textit{ab initio} molecular dynamics (AIMD). Their result
-of 6.7~kcal/mol is however lower than the experiment and the QM+RISM
-value in this study. The good agreement of QM+RISM result with the
-experiment and the best of other theoretical models indicates that
-the RISM model provides a viable alternative solvation model for uranyl
-aqua complexes in hybrid calculations and offers a reasonable accuracy
-at an affordable cost.
+solvation shell which roughly corresponds to the region of low PMF
+(see Figs.~\ref{fig:potential-of-mean-force} and \ref{fig:potential-of-mean-force-2d})
+when the sixth water molecule enters or leaves the first solvation
+shell. The water exchange barrier of 2.3~kcal/mol and 4.3~kcal/mol,
+calculated with the ``soft'' and the ``hard'' force field models,
+respectively (Table~\ref{tab:stationary-points}), are rather close
+to the estimates that we obtained above from the PMF analysis. Yet,
+these barrier values are significantly lower than the result of 9.1~kcal/mol
+barrier, determined with the QM+RISM model (Table~\ref{tab:stationary-points})
+which is in excellent agreement with the experimental estimate of
+9~kcal/mol \citep{Farkas00}. A comparison of the three models regarding
+the location and the orientation of the second-shell explicit water
+molecule suggests that the force field models may slightly underestimate
+the strength and the directional nature of the hydrogen bonds between
+the incoming explicit water molecule and two of the explicit aqua
+ligands of uranyl. This may explain the low activation energy for
+water exchange and the stability of the six-coordinated metastable
+state relative to the structure with an explicit water molecule in
+the second shell, as predicted by the force field models, Fig.~\ref{fig:free-energy-profiles}.
+
+Finally, we compare the present results for the activation barrier
+of exchanging two water molecules of the first and the second solvation
+shell of uranyl to results of some state-of-the-art QM calculations.
+The computed values are sensitive to the choice of the reaction path,
+the level of theory, and the solvation model \citep{Tiwari14}. Tsushima
+\citep{Tsushima07} reported activation barriers of 10.1~kcal/mol
+and 4.8~kcal/mol at the B3LYP level with the CPCM approach by considering
+final states with one and two hydrogen bonds, respectively. The latter
+value is close to the prediction of 4.5~kcal/mol of an earlier study
+\citep{Vallet01}, yet both values are lower than the experimental
+result. Rotzinger applied a complete active space self-consistent
+field method in combination with two solvation models, a PCM approach
+and a spherical cavity self-consistent reaction field (SCRF) solvation
+models \citep{Rotzinger07}; for the associative pathway he obtained
+barriers of 9.4~kcal/mol and 6.5~kcal/mol, respectively. Bühl and
+Kabrede obtained the free energy profile for ligand exchange of uranyl,
+employing thermodynamic sampling via constrained Car--Parrinello molecular
+dynamics, using a computational approach where they combined BLYP
+density functional with norm conserving pseudo-potentials \citep{Buehl06b}.
+They observed a ``shallow ragged plateau'' around the six-coordinated
+intermediate and estimated the activation barrier by 6.7~kcal/mol,
+only slightly lower than the experiment and the present QM+RISM value.
+
+The good agreement between the QM+RISM result and experiment \citep{Farkas00}
+as well as the values of the best other theoretical models indicates
+that the 3D RISM model provides an attractive solvation approach,
+in particular as a hybrid solvation strategy applied to uranyl aqua
+complexes. Using a semi-local DFT method in combination with the 3D
+RISM strategy offers an adequate accuracy at an affordable cost.
 
@@
 
-We presented a parallelized implementation of 3D RISM for charged
-species and verified its performance on examples of atomic ions and
-uranyl aqua complexes. We demonstrated that the method is not susceptible
-to a deficiency exposed in our previous work for the 1D RISM technique
-where an artificial superposition of explicit aqua ligands and the
-RISM medium leads to an overestimation of effective uranyl--water
-coordination and an excessive elongation of the U--O and U--OW bonds:
-for the uranyl aqua complexes with 4--6 explicit ligands RISM predict
-essentially the same structure of the first solvation shell as PCM.
-A net effect of a missing explicit aqua ligand is qualitatively correctly
-represented by an excess density of the RISM medium, leading to comparable
-results for uranyl with 5 and 4 explicit water ligands. However, because
-of the strong closure order dependence of the results and occasional
-divergence for higher order closures a successful model will avoid
-treating the first solvation shell of highly charged ions implicitly.
-With a first solvation shell treated by explicit aqua ligands, the
-observables are nearly independent of the closure order.
-
-A hybrid approach employing two very different models of ion--water
-interactions, \emph{ab initio} and force field, relies heavily on
-the consistence of the two interaction models. Use of the KL2 force
-field, featuring high charges of interacting sites of uranyl, as selected
-in our previous study as the one predicting the highest uranyl solvation
-energy, leads to a substitution of explicit QM aqua ligands by the
-RISM medium far too easily for routine calculations to be practical.
-Such artifacts affected the 1D RISM approach as well suggesting that
-this particular QM+RISM combination may need to be revisited.
+We presented a parallelized implementation of the 3D RISM approach
+for charged species in (aqueous) solution and verified its performance
+on examples of atomic ions and uranyl aqua complexes. We demonstrated
+that the method is not susceptible to the deficiency exposed in our
+previous work for the 1D RISM technique \citep{Bo15} where an artificial
+superposition of explicit aqua ligands and the RISM medium leads to
+an overestimation of the effective uranyl--water coordination, resulting
+in an excessive elongation of the U--O and U--O$_{\text{w}}$ bonds.
+With the present 3D RISM method, semi-local DFT calculations on uranyl
+aqua complexes with 4--6 explicit ligands predict essentially the
+same structure of the first solvation shell as the conventional PCM
+strategy. The net effect of a missing explicit aqua ligand in the
+first solvation shell is correctly represented in a qualitative fashion
+by an excess solvent density of the RISM medium, leading to comparable
+results for uranyl with 5 and 4 explicit water ligands. However, for
+``bare'' solvated ions, without explicit solvent molecules as ligands,
+the results depend notably on the order of the closure; the iterative
+procedure used to solve the coupled integral equations, even diverged
+occasionally for closure relations of higher order. Yet, a successful
+model will avoid describing the first solvation shell of highly charged
+ions in implicit fashion. When the first solvation shell is treated
+by explicit aqua ligands, we did not note any numerical problems and
+the observables turned out to be nearly independent of the closure
+order.
+
+Any hybrid approach combining two models of ion--water interactions,
+\emph{``first principles'' }and force field, relies heavily on the
+consistency of the two interaction models. For uranyl hydration, the
+force field KL2, which features large charges of interacting sites,
+as selected in our previous study \citep{Bo15} for predicting the
+largest uranyl solvation energy (in absolute terms), is far from being
+practical for routine calculations because it too strongly favors
+substitution of explicit QM aqua ligands by the RISM medium. Such
+artifacts affected the 1D RISM approach, suggesting that this particular
+QM+RISM combination may need to be revisited.
 
@@
 shell and the bulk water, the hybrid QM+RISM model produces meaningful
-structure of the first solvation shell and solvation energies competitive
-to the state of the art PCM. By examining the water exchange activation
-energy we demonstrated the usefulness of the hybrid QM+RISM models
-to address properties that are not easily reproduced by pure force
-field models. The activation energy is predicted too low with the
-force field models, likely due to the misrepresentation of hydrogen
-bonding. It is in remarkably good agreement with the experiment when
-the hybrid QM+RISM model is employed.
+structures of the first solvation shell and results in solvation energies
+that are competitive to the state-of-the-art PCM procedures.
+
+By examining the activation energy for exchange of aqua ligands, we
+convincingly demonstrated the benefits of the hybrid QM+RISM model
+for addressing properties that are not easily reproduced by pure force
+field models. This type of activation energy is predicted too low
+with the current force field models, likely due to a misrepresentation
+of hydrogen bonding. In contrast, the present hybrid QM+RISM model
+yields an activation barrier for ligand exchange that agrees remarkably
+well with experiment \citep{Farkas00}.
 
@@
 
-We thank Prof.~H.-J.~Bungartz and co-workers for valuable discussions
-and Dr.~L.~Jager and Prof.~M.~Griebel for providing the source
-of the BGY3D program package. BL is grateful for a fellowship by the
-International Graduate School of Science and Engineering at Technische
-Universität München. This work was partially supported by Bundesministerium
-für Bildung und Forschung (grant No.\ 01RC1106A).
-
-\bibliographystyle{elsarticle-num}
-\bibliography{uo2}
-
+We thank \textcolor{blue}{Prof.~H.-J.~Bungartz, his co-workers,
+and Dr.\ S.\ Krüger} for valuable discussions. BL is grateful for
+a fellowship by the International Graduate School of Science and Engineering
+at Technische Universität München. The early stage of this work benefitted
+from Dr.\ L.\ Jager making his BGY3D program package available.
+This work was partially supported by Bundesministerium für Bildung
+und Forschung (grant No.\ 01RC1106A).
+
+\bibliographystyle{achemso}
+\bibliography{../bibliography}
+
+\begin{tocentry}
+\includegraphics[height=3.5cm]{uo2-II.d/4w-composite}
+\end{tocentry}
 \end{document}

### 2015-04-17 doc/uo2_paper/uo2-II.tex
AM: graphics from CWD again
AM: bib from CWD
AM: achemso is the default here
AM: rm refs to sections

@@
 \providecommand{\tabularnewline}{\\}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
 
@@
 approach as well as a numerical treatment of the long-range Coulomb
-field of a charged solute, Section~\ref{sec:methodology}. Then we
+field of a charged solute. Then we
 validate the implementation by comparing the results of 1D and 3D
-techniques for alkali and halide ions, Section~\ref{sec:simple-ions}.
+techniques for alkali and halide ions.
 We further evaluate the performance of the method for the uranyl(VI)
 dication in aqueous solution, represented by aqua complexes $[\text{UO}_{2}(\text{H}_{2}\text{O})_{n}]^{2+}$,
-($n=4-6$), using a force field description, Section~\ref{sub:MM+RISM},
+($n=4-6$), using a force field description,
 and a quantum mechanical representation of the interactions within
-the first solvation shell, Section~\ref{sub:QM+RISM}. We also compare
+the first solvation shell. We also compare
 the results to those of PCM calculations. Finally, we analyze the
 activation barrier for exchanging a water molecule in the first solvation
-shell, Section~\ref{sub:water-exchange}, using QM and MM models
+shell, using QM and MM models
 of interactions. With the exception of the water exchange process
@@
 
-\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/4w-composite}
+\centering{}\includegraphics[width=1\linewidth]{4w-composite}
 \end{figure}
@@
 
-\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig2}
+\centering{}\includegraphics[width=1\linewidth]{fig2}
 \end{figure}
@@
 
-\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/fig3}
+\centering{}\includegraphics[width=1\linewidth]{fig3}
 \end{figure}
@@
 
-\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf}
+\centering{}\includegraphics[width=1\linewidth]{pmf}
 \end{figure}
@@
 
-\centering{}\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/pmf-2d}
+\centering{}\includegraphics[width=1\linewidth]{pmf-2d}
 \end{figure}
@@
 \centering{}\smallskip{}
-\includegraphics[width=1\linewidth]{uo2-II\lyxdot d/profiles}
+\includegraphics[width=1\linewidth]{profiles}
 \end{figure}
@@
 
-\bibliographystyle{achemso}
-\bibliography{../bibliography}
+\bibliography{bibliography}
 
 \begin{tocentry}
-\includegraphics[height=3.5cm]{uo2-II.d/4w-composite}
+\includegraphics[height=3.5cm]{4w-composite}
 \end{tocentry}

### 2015-04-17 doc/uo2_paper/uo2-II.tex
AM: rm blue text

@@
 PSE2 and PSE3 (solid lines of decreasing weight). The dashed line
-represents the MD result of Ref.~\citep{Bo15}\textcolor{blue}{.}
+represents the MD result of Ref.~\citep{Bo15}.
 \label{fig:potential-of-mean-force}}
@@
 
-We thank \textcolor{blue}{Prof.~H.-J.~Bungartz, his co-workers,
-and Dr.\ S.\ Krüger} for valuable discussions. BL is grateful for
+We thank Prof.~H.-J.~Bungartz, his co-workers,
+and Dr.\ S.\ Krüger for valuable discussions. BL is grateful for
 a fellowship by the International Graduate School of Science and Engineering

### 2015-04-17 doc/uo2_paper/uo2-II.tex
AM: inline bibliography

@@
 
-\bibliography{bibliography}
+% %\bibliography{bibliography}
+\begin{thebibliography}{96}
+
+\bibitem{book:Morss06}
+Morss,~L.~R.; Edelstein,~N.~M.; Fuger,~J. \emph{The Chemistry of the Actinide
+  and Transactinide Elements}, 3rd ed.; Springer: Netherlands, 2006; Vol.~1; pp
+  253--697.
+\bibitem{book:Grenthe92}
+Grenthe,~I.; Fuger,~J.; Konings,~R.; Lemire,~R.; Muller,~A.; Nguyen-Trung,~C.;
+  Wanner,~H. \emph{Chemical Thermodynamics of Uranium}; North-Holland:
+  Amsterdam, 1992.
+\bibitem{Silva95}
+Silva,~R.; Nitsche,~H. Actinide Environmental Chemistry. \emph{Radiochim. Acta}
+  \textbf{1995}, \emph{70}, 377--396.
+\bibitem{Tomasi05}
+Tomasi,~J.; Mennucci,~B.; Cammi,~R. Quantum Mechanical Continuum Solvation
+  Models. \emph{Chem. Rev.} \textbf{2005}, \emph{105}, 2999--3094.
+\bibitem{Fuchs02}
+Fuchs,~M.; Shor,~A.~M.; R\"{o}sch,~N. The Hydration of the Uranyl Dication.
+  {I}ncorporation of Solvent Effects in Parallel Density Functional
+  Calculations with the Program \textsc{{P}ara{G}auss}. \emph{Int. J. Quantum
+  Chem.} \textbf{2002}, \emph{86}, 487--501.
+\bibitem{Klamt95}
+Klamt,~A. Conductor-like Screening Model for Real Solvents: A New Approach to
+  the Quantitative Calculation of Solvation Phenomena. \emph{J. Phys. Chem.}
+  \textbf{1995}, \emph{99}, 2224--2235.
+\bibitem{Moskaleva04}
+Moskaleva,~L.~V.; Kr\"{u}ger,~S.; Sp\"{o}rl,~A.; R\"{o}sch,~N. Role of
+  Solvation in the Reduction of the Uranyl Dication by Water: A Density
+  Functional Study. \emph{Inorg. Chem.} \textbf{2004}, \emph{43},
+  4080--4090.
+\bibitem{Cao05}
+Cao,~Z.; Balasubramanian,~K. Theoretical Studies of
+  {UO$_{2}($H$_{2}$O$)_{n}^{2+}$}, {NpO$_{2}($H$_{2}$O$)_{n}^{+}$}, and
+  {PuO$_{2}($H$_{2}$O$)_{n}^{2+}$} Complexes ($n=4-6$) in Aqueous Solution and
+  Gas Phase. \emph{J. Chem. Phys.} \textbf{2005}, \emph{123}, 114309.
+\bibitem{Shamov05}
+Shamov,~G.~A.; Schreckenbach,~G. Density Functional Studies of Actinyl Aquo
+  Complexes Studied Using Small-Core Effective Core Potentials and a Scalar
+  Four-Component Relativistic Method. \emph{J. Phys. Chem. A} \textbf{2005},
+  \emph{109}, 10961--10974.
+\bibitem{Gutowski06}
+Gutowski,~K.~E.; Dixon,~D.~A. Predicting the Energy of the Water Exchange
+  Reaction and Free Energy of Solvation for the Uranyl Ion in Aqueous Solution.
+  \emph{J. Phys. Chem. A} \textbf{2006}, \emph{110}, 8840--8856.
+\bibitem{Guilbaud93}
+Guilbaud,~P.; Wipff,~G. Hydration of Uranyl ({UO$_2^{2+}$}) Cation and its
+  Nitrate Ion and 18-crown-6 Adducts Studied by Molecular Dynamics Simulations.
+  \emph{J. Phys. Chem.} \textbf{1993}, \emph{97}, 5685--5692.
+\bibitem{Guilbaud96}
+Guilbaud,~P.; Wipff,~G. Force Field Representation of the {UO}$_2^{2+}$ Cation
+  from Free Energy {MD} Simulations in Water. {T}ests on its 18-crown-6 and
+  {NO}$_3$- Adducts, and on its calix[6]arene6- and {CMPO} Complexes. \emph{J.
+  Mol. Struct.: THEOCHEM} \textbf{1996}, \emph{366}, 55--63.
+\bibitem{Rai12}
+Rai,~N.; Tiwari,~S.~P.; Maginn,~E.~J. Force Field Development for Actinyl Ions
+  via Quantum Mechanical Calculations: An Approach to Account for Many Body
+  Solvation Effects. \emph{J. Phys. Chem. B} \textbf{2012}, \emph{116},
+  10885--10897.
+\bibitem{Pomogaev13}
+Pomogaev,~V.; Tiwari,~S.~P.; Rai,~N.; Goff,~G.~S.; Runde,~W.; Schneider,~W.~F.;
+  Maginn,~E.~J. Development and Application of Effective Pairwise Potentials
+  for {UO}$_2^{n+}$, {NpO}$_2^{n+}$, {PuO}$_2^{n+}$, and {AmO}$_2^{n+}$ ($n =
+  1, 2$) Ions with Water. \emph{Phys. Chem. Chem. Phys.} \textbf{2013},
+  \emph{15}, 15954--15963.
+\bibitem{Kerisit13}
+Kerisit,~S.; Liu,~C. Structure, Kinetics, and Thermodynamics of the Aqueous
+  Uranyl({VI}) Cation. \emph{J. Phys. Chem. A} \textbf{2013}, \emph{117},
+  6421--6432.
+\bibitem{Tiwari14}
+Tiwari,~S.~P.; Rai,~N.; Maginn,~E.~J. Dynamics of Actinyl Ions in Water: A
+  Molecular Dynamics Simulation Study. \emph{Phys. Chem. Chem. Phys.}
+  \textbf{2014}, \emph{16}, 8060--8069.
+\bibitem[Griebel \latin{et~al.}(2007)Griebel, Knapek, and
+  Zumbusch]{book:Griebel07}
+Griebel,~M.; Knapek,~S.; Zumbusch,~G. \emph{Numerical Simulation in Molecular
+  Dynamics: Numerics, Algorithms, Parallelization, Applications}; Springer:
+  Berlin Heidelberg, 2007.
+\bibitem{Gordon00}
+Gordon,~M.~S.; Freitag,~M.~A.; Bandyopadhyay,~P.; Jensen,~J.~H.; Kairys,~V.;
+  Stevens,~W.~J. The Effective Fragment Potential Method: A {QM}-Based {MM}
+  Approach to Modeling Environmental Effects in Chemistry. \emph{J. Phys. Chem.
+  A} \textbf{2000}, \emph{105}, 293--307.
+\bibitem{hagberg05}
+Hagberg,~D.; Karlstr\"{o}m,~G.; Roos,~B.~O.; Gagliardi,~L. The Coordination of
+  Uranyl in Water: A Combined Quantum Chemical and Molecular Simulation Study.
+  \emph{J. Am. Chem. Soc.} \textbf{2005}, \emph{127}, 14250--14256.
+\bibitem[Frick \latin{et~al.}(2009)Frick, Hofer, Pribil, Randolf, and
+  Rode]{Frick09}
+Frick,~R.~J.; Hofer,~T.~S.; Pribil,~A.~B.; Randolf,~B.~R.; Rode,~B.~M.
+  Structure and Dynamics of the {UO$_2^{2+}$} Ion in Aqueous Solution: An
+  \textit{Ab Initio} {QMCF MD} Study. \emph{J. Phys. Chem. A} \textbf{2009},
+  \emph{113}, 12496--12503.
+\bibitem{Tirler13}
+Tirler,~A.~O.; Weiss,~A. K.~H.; Hofer,~T.~S. A Comparative Quantum Mechanical
+  Charge Field Study of Uranyl Mono- and Dicarbonate Species in Aqueous
+  Solution. \emph{J. Phys. Chem. B} \textbf{2013}, \emph{117},
+  16174--16187.
+\bibitem{Buehl05}
+B\"{u}hl,~M.; Diss,~R.; Wipff,~G. Coordination Environment of Aqueous
+  Uranyl({VI}) Ion. \emph{J. Am. Chem. Soc.} \textbf{2005}, \emph{127},
+  13506--13507.
+\bibitem{Buehl06}
+B\"{u}hl,~M.; Kabrede,~H.; Diss,~R.; Wipff,~G. Effect of Hydration on
+  Coordination Properties of Uranyl({VI}) Complexes. {A} First-Principles
+  Molecular Dynamics Study. \emph{J. Am. Chem. Soc.} \textbf{2006}, \emph{128},
+  6357--6368.
+\bibitem{Buehl11}
+B\"{u}hl,~M.; Wipff,~G. Insights into Uranyl Chemistry from Molecular Dynamics
+  Simulations. \emph{ChemPhysChem} \textbf{2011}, \emph{12}, 3095--3105.
+\bibitem{Nichols08}
+Nichols,~P.; Bylaska,~E.~J.; Schenter,~G.~K.; de~Jong,~W. Equatorial and Apical
+  Solvent Shells of the {UO$_2^{2+}$} Ion. \emph{J. Chem. Phys.} \textbf{2008},
+  \emph{128}, 124507.
+\bibitem{book:Hirata03}
+Hirata,~F. \emph{Molecular Theory of Solvation}; Kluwer Academic Publisher:
+  Dordrecht, 2003.
+\bibitem{book:Hansen06}
+Hansen,~J.-P.; McDonald,~I.~R. \emph{Theory of Simple Liquids}, 3rd ed.;
+  Academic Press: Burlington, 2006.
+\bibitem{Chandler72}
+Chandler,~D.; Andersen,~H.~C. Optimized Cluster Expansions for Classical
+  Fluids. {II}. {T}heory of Molecular Liquids. \emph{J. Chem. Phys.}
+  \textbf{1972}, \emph{57}, 1930--1937.
+\bibitem{Hirata81}
+Hirata,~F.; Rossky,~P.~J. An Extended {RISM} Equation for Molecular Polar
+  Fluids. \emph{Chem. Phys. Lett.} \textbf{1981}, \emph{83}, 329--334.
+\bibitem{Hirata82}
+Hirata,~F.; Pettitt,~B.~M.; Rossky,~P.~J. Application of an Extended {RISM}
+  Equation to Dipolar and Quadrupolar Fluids. \emph{J. Chem. Phys.}
+  \textbf{1982}, \emph{77}, 509--520.
+\bibitem{Kinoshita97}
+Kinoshita,~M.; Hirata,~F. Analysis of Salt Effects on Solubility of Noble Gases
+  in Water Using the Reference Interaction Site Model Theory. \emph{J. Chem.
+  Phys.} \textbf{1997}, \emph{106}, 5202--5215.
+\bibitem{Beglov95}
+Beglov,~D.; Roux,~B. Numerical Solution of the Hypernetted Chain Equation for a
+  Solute of Arbitrary Geometry in Three Dimensions. \emph{J. Chem. Phys.}
+  \textbf{1995}, \emph{103}, 360--364.
+\bibitem{Kovalenko98}
+Kovalenko,~A.; Hirata,~F. Three-Dimensional Density Profiles of Water in
+  Contact with a Solute of Arbitrary Shape: A {RISM} Approach. \emph{Chem.
+  Phys. Lett.} \textbf{1998}, \emph{290}, 237--244.
+\bibitem{Singer85}
+Singer,~S.~J.; Chandler,~D. Free Energy Functions in the Extended {RISM}
+  Approximation. \emph{Mol. Phys.} \textbf{1985}, \emph{55}, 621--625.
+\bibitem{Kast08}
+Kast,~S.~M.; Kloss,~T. Closed-Form Expressions of the Chemical Potential for
+  Integral Equation Closures with Certain Bridge Functions. \emph{J. Chem.
+  Phys.} \textbf{2008}, \emph{129}, 236101.
+\bibitem{Fischer80}
+Fischer,~J.; Methfessel,~M. Born--Green--Yvon Approach to the Local Densities
+  of a Fluid at Interfaces. \emph{Phys. Rev. A} \textbf{1980}, \emph{22},
+  2836--2843.
+\bibitem{Griebel08}
+Griebel,~M.; Jager,~L. The {BGY3dM} Model for the Approximation of Solvent
+  Densities. \emph{J. Chem. Phys.} \textbf{2008}, \emph{129}, 174511.
+\bibitem{Perkyns92}
+Perkyns,~J.~S.; Pettitt,~B.~M. A Dielectrically Consistent Interaction Site
+  Theory for Solvent--Electrolyte Mixtures. \emph{Chem. Phys. Lett.}
+  \textbf{1992}, \emph{190}, 626--630.
+\bibitem{Perkyns94}
+Perkyns,~J.; Pettitt,~B.~M. Integral Equation Approaches to Structure and
+  Thermodynamics of Aqueous Salt Solutions. \emph{Biophys. Chem.}
+  \textbf{1994}, \emph{51}, 129--146.
+\bibitem[Chiodo \latin{et~al.}(2007)Chiodo, Chuev, Erofeeva, Fedorov, Russo,
+  and Sicilia]{Chiodo07}
+Chiodo,~S.; Chuev,~G.~N.; Erofeeva,~S.~E.; Fedorov,~M.~V.; Russo,~N.;
+  Sicilia,~E. Comparative Study of Electrostatic Solvent Response by {RISM} and
+  {PCM} Methods. \emph{Int. J. Quantum Chem.} \textbf{2007}, \emph{107},
+  265--274.
+\bibitem{Chuev08}
+Chuev,~G.~N.; Fedorov,~M.~V.; Chiodo,~S.; Russo,~N.; Sicilia,~E. Hydration of
+  Ionic Species Studied by the Reference Interaction Site Model with a
+  Repulsive Bridge Correction. \emph{J. Comput. Chem.} \textbf{2008},
+  \emph{29}, 2406--2415.
+\bibitem{Joung13}
+Joung,~I.~S.; Luchko,~T.; Case,~D.~A. Simple Electrolyte Solutions: Comparison
+  of {DRISM} and Molecular Dynamics Results for Alkali Halide Solutions.
+  \emph{J. Chem. Phys.} \textbf{2013}, \emph{138}, 044103.
+\bibitem{Makov95}
+Makov,~G.; Payne,~M.~C. Periodic Boundary Conditions in \textit{ab initio}
+  Calculations. \emph{Phys. Rev. B} \textbf{1995}, \emph{51}, 4014--4022.
+\bibitem{Ng74}
+Ng,~K.-C. Hypernetted Chain Solutions for the Classical One-Component Plasma up
+  to {$\Gamma$}=7000. \emph{J. Chem. Phys.} \textbf{1974}, \emph{61},
+  2680--2689.
+\bibitem{Kovalenko00b}
+Kovalenko,~A.; Hirata,~F. Potentials of Mean Force of Simple Ions in Ambient
+  Aqueous Solution. {I}. {T}hree-Dimensional Reference Interaction Site Model
+  Approach. \emph{J. Chem. Phys.} \textbf{2000}, \emph{112}, 10391--10402.
+\bibitem{Kloss08b}
+Kloss,~T.; Kast,~S.~M. Treatment of Charged Solutes in Three-Dimensional
+  Integral Equation Theory. \emph{J. Chem. Phys.} \textbf{2008}, \emph{128},
+  134505.
+\bibitem[Perkyns \latin{et~al.}(2010)Perkyns, Lynch, Howard, and
+  Pettitt]{Perkyns10}
+Perkyns,~J.~S.; Lynch,~G.~C.; Howard,~J.~J.; Pettitt,~B.~M. Protein Solvation
+  from Theory and Simulation: Exact Treatment of Coulomb Interactions in
+  Three-Dimensional Theories. \emph{J. Chem. Phys.} \textbf{2010}, \emph{132},
+  064106.
+\bibitem{Heil15}
+Heil,~J.; Kast,~S.~M. {3D RISM} Theory with Fast Reciprocal-Space
+  Electrostatics. \emph{J. Chem. Phys.} \textbf{2015}, \emph{142}, 114107.
+\bibitem{Tenno93}
+Ten-no,~S.; Hirata,~F.; Kato,~S. A Hybrid Approach for the Solvent Effect on
+  the Electronic Structure of a Solute Based on the {RISM} and
+  {H}artree--{F}ock Equations. \emph{Chem. Phys. Lett.} \textbf{1993},
+  \emph{214}, 391--396.
+\bibitem{Tenno94}
+Ten-no,~S.; Hirata,~F.; Kato,~S. Reference Interaction Site Model
+  Self-Consistent Field Study for Solvation Effect on Carbonyl Compounds in
+  Aqueous Solution. \emph{J. Chem. Phys.} \textbf{1994}, \emph{100},
+  7443--7453.
+\bibitem{Kovalenko99}
+Kovalenko,~A.; Hirata,~F. Self-Consistent Description of a Metal--Water
+  Interface by the {K}ohn--{S}ham Density Functional Theory and the
+  Three-Dimensional Reference Interaction Site Model. \emph{J. Chem. Phys.}
+  \textbf{1999}, \emph{110}, 10095--10112.
+\bibitem{Kloss08}
+Kloss,~T.; Heil,~J.; Kast,~S.~M. Quantum Chemistry in Solution by Combining
+  {3D} Integral Equation Theory with a Cluster Embedding Approach. \emph{J.
+  Phys. Chem. B} \textbf{2008}, \emph{112}, 4337--4343.
+\bibitem{Gusarov06}
+Gusarov,~S.; Ziegler,~T.; Kovalenko,~A. Self-Consistent Combination of the
+  Three-Dimensional {RISM} Theory of Molecular Solvation with Analytical
+  Gradients and the {A}msterdam Density Functional Package. \emph{J. Phys.
+  Chem. A} \textbf{2006}, \emph{110}, 6083--6090.
+\bibitem{Sato13}
+Sato,~H. A Modern Solvation Theory: Quantum Chemistry and Statistical
+  Chemistry. \emph{Phys. Chem. Chem. Phys.} \textbf{2013}, \emph{15},
+  7450--7465.
+\bibitem{Bo15}
+Li,~B.; Matveev,~A.~V.; Kr\"uger,~S.; R\"osch,~N. Uranyl Solvation by a
+  Reference Interaction Site Model. \emph{Comput. Theor. Chem.} \textbf{2015},
+  \emph{1051}, 151--160.
+\bibitem{Leeuwen59}
+van Leeuwen,~J.; Groeneveld,~J.; de~Boer,~J. New Method for the Calculation of
+  the Pair Correlation Function. {I}. \emph{Physica} \textbf{1959}, \emph{25},
+  792--808.
+\bibitem{Berendsen87}
+Berendsen,~H. J.~C.; Grigera,~J.~R.; Straatsma,~T.~P. The Missing Term in
+  Effective Pair Potentials. \emph{J. Phys. Chem.} \textbf{1987}, \emph{91},
+  6269--6271.
+\bibitem{Mahoney00}
+Mahoney,~M.~W.; Jorgensen,~W.~L. A Five-Site Model for Liquid Water and the
+  Reproduction of the Density Anomaly by Rigid, Nonpolarizable Potential
+  Functions. \emph{J. Chem. Phys.} \textbf{2000}, \emph{112}, 8910--8922.
+\bibitem{Bridgeman64}
+Bridgeman,~O.~C.; Aldrich,~E. Vapor Pressure Tables for Water. \emph{J. Heat
+  Transfer} \textbf{1964}, \emph{86}, 279--286, by way of NIST Chemistry
+  Webbook, \url{http://webbook.nist.gov/chemistry/}.
+\bibitem{petsc-user-ref}
+Balay,~S.; Adams,~M.~F.; Brown,~J.; Brune,~P.; Buschelman,~K.; Eijkhout,~V.;
+  Gropp,~W.~D.; Kaushik,~D.; Knepley,~M.~G.; McInnes,~L.~C. \latin{et~al.}
+  {PETS}c Users Manual. 2013.
+\bibitem{Cummings81}
+Cummings,~P.~T.; Stell,~G. Exact Asymptotic Form of the Site-Site Direct
+  Correlation Function for Rigid Polar Molecules. \emph{Mol. Phys.}
+  \textbf{1981}, \emph{44}, 529--531.
+\bibitem{FFTW05}
+Frigo,~M.; Johnson,~S.~G. The Design and Implementation of {FFTW3}.
+  \emph{Proceedings of the IEEE} \textbf{2005}, \emph{93}, 216--231.
+\bibitem{Pettitt82}
+Pettitt,~B.~M.; Rossky,~P.~J. Integral Equation Predictions of Liquid State
+  Structure for Waterlike Intermolecular Potentials. \emph{J. Chem. Phys.}
+  \textbf{1982}, \emph{77}, 1451--1457.
+\bibitem{Luchko10}
+Luchko,~T.; Gusarov,~S.; Roe,~D.~R.; Simmerling,~C.; Case,~D.~A.;
+  Tuszynski,~J.; Kovalenko,~A. Three-Dimensional Molecular Theory of Solvation
+  Coupled with Molecular Dynamics in {Amber}. \emph{J. Chem. Theory Comput.}
+  \textbf{2010}, \emph{6}, 607--624.
+\bibitem{Dunlap90}
+Dunlap,~B.~I.; R\"{o}sch,~N. In \emph{Density Functional Theory of Many-Fermion
+  Systems}; Trickey,~S.~B., Ed.; Advances in Quantum Chemistry; Academic Press,
+  1990; Vol.~21; pp 317--399.
+\bibitem{ParaGaussV3.2}
+R\"{o}sch,~N. \latin{et~al.}  {\textsc{ParaGauss}}, Version 4.0, Technische Universit\"{a}t
+  M\"{u}nchen, 2014.
+\bibitem{Matveev08}
+Matveev,~A.~V.; R\"{o}sch,~N. Atomic Approximation to the Projection on
+  Electronic States in the {D}ouglas--{K}roll--{H}ess Approach to the
+  Relativistic {K}ohn--{S}ham Method. \emph{J. Chem. Phys.} \textbf{2008},
+  \emph{128}, 244102.
+\bibitem{rev:roesch96}
+{R{\"o}sch},~N.; {Kr{\"u}ger},~S.; Mayer,~M.; Nasluzov,~V.~A. In \emph{Recent
+  Developments and Applications of Modern Density Functional Theory};
+  Seminario,~J.~M., Ed.; Theoretical and Computational Chemistry Series;
+  Elsevier: Amsterdam, 1996; Vol.~4; p 497.
+\bibitem{Goerling99}
+G\"orling,~A.; Trickey,~S.; Gisdakis,~P.; R\"osch,~N. In \emph{Organometallic
+  Bonding and Reactivity}; Brown,~J., Hofmann,~P., Eds.; Topics in
+  Organometallic Chemistry; Springer Berlin Heidelberg, 1999; Vol.~4; pp
+  109--163.
+\bibitem{kochDFT}
+Koch,~W.; Holthausen,~M.~C. \emph{A Chemist's Guide to Density Functional
+  Theory}; Wiley-VCH: Weinheim, 2001.
+\bibitem{Becke88}
+Becke,~A.~D. Density-Functional Exchange-Energy Approximation with correct
+  asymptotic behavior. \emph{Phys. Rev. A} \textbf{1988}, \emph{38},
+  3098--3100.
+\bibitem{Perdew86c}
+Perdew,~J.~P. Density-Functional Approximation for the Correlation Energy of
+  the Inhomogeneous Electron Gas. \emph{Phys. Rev. B} \textbf{1986}, \emph{33},
+  8822.
+\bibitem{Moskaleva06b}
+Moskaleva,~L.~V.; Matveev,~A.~V.; Dengler,~J.; R\"{o}sch,~N. The Heat of
+  Formation of Gaseous {PuO}$_2^{2+}$ from Relativistic Density Functional
+  Calculations. \emph{Phys. Chem. Chem. Phys.} \textbf{2006}, \emph{8},
+  3767--3773.
+\bibitem{Moskaleva06a}
+Moskaleva,~L.~V.; Matveev,~A.~V.; Kr\"{u}ger,~S.; R\"{o}sch,~N. The Heat of
+  Formation of the Uranyl Dication: Theoretical Evaluation Based on
+  Relativistic Density Functional Calculations. \emph{Chem. Eur. J.}
+  \textbf{2006}, \emph{12}, 629--634.
+\bibitem{Kremleva12b}
+Kremleva,~A.; Zhang,~Y.; Shor,~A.~M.; Kr\"uger,~S.; Joseph,~C.; Raditzky,~B.;
+  Schmeide,~K.; Sachs,~S.; Bernhard,~G.; R\"osch,~N. Uranyl({VI}) Complexation
+  by Sulfonate Ligands: A Relativistic Density Functional and Time-Resolved
+  Laser-Induced Fluorescence Spectroscopy Study. \emph{Eur. J. Inorg. Chem.}
+  \textbf{2012}, \emph{2012}, 3636--3644.
+\bibitem{Bondi64}
+Bondi,~A. van der {W}aals Volumes and Radii. \emph{J. Phys. Chem.}
+  \textbf{1964}, \emph{68}, 441--451.
+\bibitem{Su09}
+Su,~P.; Li,~H. Continuous and Smooth Potential Energy Surface for Conductorlike
+  Screening Solvation Model Using Fixed Points with Variable Areas. \emph{J.
+  Chem. Phys.} \textbf{2009}, \emph{130}, 074109.
+\bibitem{Chaffey-Millar11}
+Chaffey-Millar,~H.; Nikodem,~A.; Matveev,~A.~V.; Kr\"{u}ger,~S.; R\"{o}sch,~N.
+  Improving Upon String Methods for Transition State Discovery. \emph{J. Chem.
+  Theory Comput.} \textbf{2012}, \emph{8}, 777--786.
+\bibitem{Nikodem13}
+Nikodem,~A.; Matveev,~A.~V.; Zheng,~B.-X.; R\"{o}sch,~N. Efficient Two-Step
+  Procedures for Locating Transition States of Surface Reactions. \emph{J.
+  Chem. Theory Comput.} \textbf{2013}, \emph{9}, 588--599.
+\bibitem{Sergiievskyi14}
+Sergiievskyi,~V.~P.; Jeanmairet,~G.; Levesque,~M.; Borgis,~D. Fast Computation
+  of Solvation Free Energies with Molecular Density Functional Theory:
+  Thermodynamic-Ensemble Partial Molar Volume Corrections. \emph{J. Phys. Chem.
+  Lett.} \textbf{2014}, \emph{5}, 1935--1942.
+\bibitem{Joung08}
+Joung,~I.~S.; Cheatham,~T.~E. Determination of Alkali and Halide Monovalent Ion
+  Parameters for Use in Explicitly Solvated Biomolecular Simulations. \emph{J.
+  Phys. Chem. B} \textbf{2008}, \emph{112}, 9020--9041.
+\bibitem{Schmid00}
+Schmid,~R.; Miah,~A.~M.; Sapunov,~V.~N. A New Table of the Thermodynamic
+  Quantities of Ionic Hydration: Values and Some Applications (Enthalpy-Entropy
+  Compensation and {B}orn Radii). \emph{Phys. Chem. Chem. Phys.} \textbf{2000},
+  \emph{2}, 97--102.
+\bibitem{Gibson05}
+Gibson,~J.~K.; Haire,~R.~G.; Santos,~M.; Mar\c{c}alo,~J.; Pires~de Matos,~A.
+  Oxidation Studies of Dipositive Actinide Ions, {An}$^{2+}$ ({An} = {Th}, {U},
+  {Np}, {Pu}, {Am}) in the Gas Phase: Synthesis and Characterization of the
+  Isolated Uranyl, Neptunyl, and Plutonyl Ions {UO}$_2^{2+}$(g),
+  {NpO}$_2^{2+}$(g), and {PuO}$_2^{2+}$(g). \emph{J. Phys. Chem. A}
+  \textbf{2005}, \emph{109}, 2768--2781.
+\bibitem{Farkas00}
+Farkas,~I.; B\'{a}nyai,~I.; Szab\'{o},~Z.; Wahlgren,~U.; Grenthe,~I. Rates and
+  Mechanisms of Water Exchange of {UO}$_2^{2+}$(aq) and
+  {UO}$_2$(oxalate){F}({H}$_2${O})$^{2-}$: A Variable-Temperature $^{17}${O}
+  and $^{19}${F} {NMR} Study. \emph{Inorg. Chem.} \textbf{2000}, \emph{39},
+  799--805.
+\bibitem{Vallet01}
+Vallet,~V.; Wahlgren,~U.; Schimmelpfennig,~B.; Szab\'{o},~Z.; Grenthe,~I. The
+  Mechanism for Water Exchange in [{UO}$_2$({H}$_2${O})$_5$]$^{2+}$ and
+  [{UO}$_2$(oxalate)$_2$({H}$_2${O})]$^{2-}$, as Studied by Quantum Chemical
+  Methods. \emph{J. Am. Chem. Soc.} \textbf{2001}, \emph{123},
+  11999--12008.
+\bibitem{Kerisit10}
+Kerisit,~S.; Liu,~C. Molecular Simulation of the Diffusion of Uranyl Carbonate
+  Species in Aqueous Solution. \emph{Geochim. Cosmochim. Acta} \textbf{2010},
+  \emph{74}, 4937--4952.
+\bibitem{Denning76}
+Denning,~R.; Snellgrove,~T.; Woodwark,~D. The Electronic Structure of the
+  Uranyl Ion. \emph{Mol. Phys.} \textbf{1976}, \emph{32}, 419--442.
+\bibitem{Denning79}
+Denning,~R.; Foster,~D.; Snellgrove,~T.; Woodwark,~D. The Electronic Structure
+  of the Uranyl Ion. \emph{Mol. Phys.} \textbf{1979}, \emph{37},
+  1089--1107.
+\bibitem{Toth81}
+Toth,~L.~M.; Begun,~G.~M. Raman Spectra of Uranyl Ion and its Hydrolysis
+  Products in Aqueous Nitric Acid. \emph{J. Phys. Chem.} \textbf{1981},
+  \emph{85}, 547--549.
+\bibitem{Gal92}
+G\'al,~M.; Goggin,~P.; Mink,~J. Vibrational Spectroscopic Studies of Uranyl
+  Complexes in Aqueous and Non-Aqueous Solutions. \emph{Spectrochim. Acta, Part
+  A} \textbf{1992}, \emph{48}, 121 -- 132.
+\bibitem{Ramakrishnan11}
+Ramakrishnan,~R.; Matveev,~A.~V.; R\"osch,~N. Effects of the Self-Interaction
+  Error in {K}ohn--{S}ham Calculations: A {DFT+U} Case Study on Penta-Aqua
+  Uranyl({VI}). \emph{Comput. Theor. Chem.} \textbf{2011}, \emph{963},
+  337--343.
+\bibitem{Jackson08}
+Jackson,~V.~E.; Craciun,~R.; Dixon,~D.~A.; Peterson,~K.~A.; de~Jong,~W.~A.
+  Prediction of Vibrational Frequencies of {UO}$_2^{2+}$ at the {CCSD(T)}
+  Level. \emph{J. Phys. Chem. A} \textbf{2008}, \emph{112}, 4095--4099.
+\bibitem{Tsushima07}
+Tsushima,~S. Hydration and Water-Exchange Mechanism of the {UO}$_2^{2+}$ Ion
+  Revisited: The Validity of the "$n + 1$" Model. \emph{J. Phys. Chem. A}
+  \textbf{2007}, \emph{111}, 3613--3617.
+\bibitem{Buehl06b}
+B\"{u}hl,~M.; Kabrede,~H. Mechanism of Water Exchange in Aqueous Uranyl({VI})
+  Ion. A Density Functional Molecular Dynamics Study. \emph{Inorg. Chem.}
+  \textbf{2006}, \emph{45}, 3834--3836.
+\bibitem{Rotzinger07}
+Rotzinger,~F.~P. The Water-Exchange Mechanism of the
+  [{UO}$_2$({OH}$_2$)$_5$]$^{2+}$ Ion Revisited: The Importance of a Proper
+  Treatment of Electron Correlation. \emph{Chem. Eur. J.} \textbf{2007},
+  \emph{13}, 800--811.
+\end{thebibliography}
 

### 2015-04-17 doc/uo2_paper/uo2-II.tex
AM: PG ref, the, preprint layout

@@
 %% Do not edit unless you really know what you are doing.
-\documentclass[english,journal=jpcafh,layout=twocolumn,maxauthors=0]{achemso}
+\documentclass[english,journal=jpcafh,manuscript=article,layout=twocolumn,maxauthors=0]{achemso}
 \usepackage[T1]{fontenc}
@@
 \citep{Dunlap90} as implemented in the parallel DFT code ParaGauss
-\citep{ParaGaussV3.2}.  The atomic approximation \citep{Matveev08}
+\citep{ParaGaussV4.0}.  The atomic approximation \citep{Matveev08}
 to the second-order Douglas\textendash{}Kroll\textendash{}Hess scalar-relativistic
@@
 $\Delta G^{\#}$$^{b}$ & 2.3 & 4.3 &  & 9.1\tabularnewline
+\hline
 \end{tabular}
@@
 shell over the six-coordinated broken symmetry metastable structure.
-The QM model and ``hard'' force field model predict the six-coordinated
+The QM model and the ``hard'' force field model predict the six-coordinated
 $D_{3\text{d}}$-like state by 5.6~kcal/mol and 3.6~kcal/mol, respectively
@@
   1990; Vol.~21; pp 317--399.
-\bibitem{ParaGaussV3.2}
+\bibitem{ParaGaussV4.0}
 R\"{o}sch,~N. \latin{et~al.}  {\textsc{ParaGauss}}, Version 4.0, Technische Universit\"{a}t
-  M\"{u}nchen, 2014.
+  M\"{u}nchen, 2012.
 \bibitem{Matveev08}

### 2015-04-17 doc/uo2_paper/uo2-II.tex
AM: use Ref.~\citenum{} for inline numbers

@@
 can be assumed to be regular; see the discussion of the background
-charge correction in Ref.~\citep{Kovalenko00b}. Indeed, a physically
+charge correction in Ref.~\citenum{Kovalenko00b}. Indeed, a physically
 viable electrolyte model with a suitable susceptibility $\chi$ is
@@
 have been used. 3D domain of $20^{3}$~Å$^{3}$, grid of $64^{3}$~points.
-$^{b}$~Ref.~\citep{Bo15}. $^{c}$~MD with SPC/E water; Ref.~\citep{Joung08}.
-$^{d}$~Ref.~\citep{Schmid00}. 
+$^{b}$~Ref.~\citenum{Bo15}. $^{c}$~MD with SPC/E water; Ref.~\citenum{Joung08}.
+$^{d}$~Ref.~\citenum{Schmid00}. 
 \end{table*}
@@
 PSE2 and PSE3 (solid lines of decreasing weight). The dashed line
-represents the MD result of Ref.~\citep{Bo15}.
+represents the MD result of Ref.~\citenum{Bo15}.
 \label{fig:potential-of-mean-force}}

### 2015-04-17 doc/uo2_paper/uo2-II.tex
AM: textcolor is not used anymore

@@
 \usepackage[T1]{fontenc}
-\usepackage{color}
 \usepackage{amsmath}
@@
 
-The solvation energy of uranyl is estimated as\textcolor{green}{{} }
+The solvation energy of uranyl is estimated as
 \begin{equation}
@@
 correction in the DRISM approach \citep{Perkyns92,Cummings81,Kovalenko00b}.
-The result of the\textcolor{blue}{{} }convolution, 
+The result of the convolution, 
 \begin{equation}
@@
 \begin{figure}[tbph]
-\caption{Potential of mean\textcolor{blue}{{} }force, $W(\varrho,z)$, for the
+\caption{Potential of mean force, $W(\varrho,z)$, for the
 U--O$_{\text{w}}$ distribution function, determined with the closure

### 2015-06-01 doc/uo2_paper/uo2-II-rebuttal.txt
AM: reviewert comments as is

------------------------------------

Reviewer(s)' Comments to Author:

Reviewer: 1

Recommendation: This paper is publishable subject to minor revisions
noted.  Further review is not needed.

Comments:

The manuscript by Matveev and coworkers presents structure and
energetics of solvated uranyl aqua ion using 3d reference interaction
site model (RISM) for the solvent. Authors have coupled RIS model for
water with pair-wise additive (LJ+Coul) force field and QM (BP86) to
model uranyl ion with 4-6 water ligands that are explicitly considered
in this study. The results show that 3D-RIS model is a significant
improvement on their prior work using 1D RISM approach. The structural
features (of the urnayl aqua ion) presented in this work are in good
agreement with the previous computational work and experimental
data. The new solvent model with QM representation of the aqua uranyl
ion reproduces the experimentally observed activation energy barrier
for the water exchange. This is the key highlight of the work as most
other simulations significantly under predict this value. Authors
attribute this to the inability of force field models to capture
hydrogen boding. The manuscript is well written and presents an
alternate approach to continuum solvation models and atomistic
simulations. The manuscript can be published after authors address the
following comments:

1: I was some what surprised at the use of BP86 functional in this
work given how dated is the functional until I read the water exchange
(dynamics) part of the manuscript. Could authors repeat this
calculations with either M06-2X or PBE0 (does quite well for water)
and see the effect of the functional on exchange energetics?

2: When discussing the structural feature with QM/RISM, symmetry was
imposed uranyl aqua ion. Will relaxing the symmetry constraint result
in enhanced water density in the under coordinated uranyl ion?

3: Page 17, line 26: Authors explain shortening of U-Ow bond length is
due to medium screening of ligand-ligand interaction in first
solvation shell. I can't understand how that is feasible. If RISM
results in water density within the first solvation shell then I would
think that it should increase the bond length rather than shortening
it. Perhaps adding some quantitative discussion can be helpful here.

4: The figures (rdf) could benefit from using different line styles
(with possibly color) rather than linewidths.

------------------------------------
Reviewer: 2

Recommendation: This paper is probably publishable, but should be
reviewed again in revised form before it is accepted.

Comments:

The manuscript presents a computational study of partially hydrated
UO22+ obtained via a new 3D RISM implementation developed by the
authors. This work is a continuation of a previous study using an 1D
RISM approach, which was shown to be not sufficiently accurate to
investigate this system.

Both a force field approach and DFT-based quantum chemical 3D-RISM
investigations are reported and compared to quantum mechanical cluster
calculations in combination with the COSMO solvent model. Special
focus has been given on the influence of the order of the closure
relation (partial series closure of order 1 to 3). A further focus of
this investigation was the prediction of the water exchange rate of
ligands bound to the uranium atom of UO22+.

The new aspects of the implementation of the 3D RISM methodology and
the results for aqueous uranyl(VI) are of interest to readers in
physical chemistry and related disciplines. Nevertheless, the
following points should be addressed by the authors before the
manuscript can be considered for publication:

1. The investigated uranyl(VI) complexes contained four to six water
molecules bound to the uranium atom. Why didn't the authors considered
adding further solvent molecules via hydrogen bonds to the oxygen
atoms of the uranyl, since they act as H-bond acceptors? Does the
current implementation not permit coordination of further molecules to
the uranyl oxygens? If not, this would be a noticable limitation of
the implemenation that should be outlined and discussed in the
text. Especially in case of the force field based methodology, the
authors report a mere computing time of 40s when "starting from
scratch". The authors should in any case include data obtained for a
fully hydrated uranly(VI) and compare it to the results of the
partially hydrated ion (assuming the implementaion permits such a
treatment).

2. In the methods section the authors note that “in QM+RISM
calculations of the symmetric uranly aqua complexes the U-O bonds were
relaxed...” and later “that the SPC/E geometry of explicit water
ligands in the aqua complexes … features an O-H distance of 100 pm and
an H-O-H angel of 109.47°.”

Does this mean that the intramolecular degrees of freedom of the
ligand molecules have been kept rigid at bulk configuration? If yes,
this implies that the water molecules have a wrong (geometric) solvent
polarisation, since the geometry of water molecules coordinated to
polarising agents such as ions differ significantly from that of the
bulk. The authors should critically discuss the associated errors
resulting from such a constraint in their revised manuscript.

Furthermore, if the geometric polarisation in the partial hydration
shell is truly negelected, it is highly surprising that the associated
free energy data is still in reasonable agreement with other
sources. Also this point should be critically discussed in the
manuscript.

3. Different versions of the SPC/E water model have been employed,
namely PR-SPC/E and cSPC/E. The reason for choosing different models
is not clear and should be discussed in detail in the
manuscript. Since the respective benchmark computations for the
monoatomic alkali and halide ions presumably require only seconds to
minutes, a comparison of the influence on the water molecule on the
resulting free energy appears feasible and should be included in the
manuscript for comparison.

Minor point: On page 15 the authors write that “the solvation energy
             .. is about 5 kcal higher”.  The unit of energy is most
             likely kcal/mol.

------------------------------------
Reviewer: 3

Recommendation: This paper is publishable subject to minor revisions
noted.  Further review is not needed.

Comments:

This is a QM and MM study of aqueous uranyl hydrate complexes using a
specific variant of a solvation model (3D-RISM), as implemented by the
authors. I am not qualified to judge the validity of the equations in
the Methodology section (or their correct implementation). The methods
employed (XC functional, MM force fields) appear to be appropriate,
however. Extending their previous study that used a simpler version of
the solvation model (1D-RISM), the authors find that the more refined
3D variant improves the accuracy and numerical stability of the
results. The paper is well written, and the results clearly presented
and discussed. I can recommend acceptance; even though the paper could
be published as is, I have a couple of minor suggestions, which the
authors may want to address before it goes to print:

1) There are force fields for uranyl that do not require freezing of
U-O bond lengths (e.g. by Guilbaud and Wipff); some justification why
this was necessary in the present study would seem in order.

2) An indication of the relative timings for QM+(1D-)RISM and
QM+(3D-)RISM calculations would be interesting.

### 2015-06-17 doc/uo2_paper/uo2-II.tex
AM: these packages are auto loaded
AM: preprint layout
AM: use non-specific names for figures
AM: without fontenc umlauts are missing
AM: pagebreak for single page abstract
AM: edits of BL sent out on Jun 12th

@@
 %% Do not edit unless you really know what you are doing.
-\documentclass[english,journal=jpcafh,manuscript=article,layout=twocolumn,maxauthors=0]{achemso}
+\documentclass[english,journal=jpcafh,manuscript=article,maxauthors=0]{achemso}
 \usepackage[T1]{fontenc}
+\usepackage{color,soul}
 \usepackage{amsmath}
@@
 \usepackage{esint}
-\usepackage[numbers]{natbib}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
+\newcommand{\hlc}[2][yellow]{ {\sethlcolor{#1} \hl{#2}} }
 \usepackage{mathptmx}
@@
 \captionsetup{font={bf}}
-\usepackage{natmove}
 
@@
 \maketitle
+\pagebreak
 
@@
 E5630 @ 2.53~GHz CPUs, where one achieves a parallel efficiency above
-40\%. The non-linear solver converges in 8 iterations with about 380
+40\%. \hlc[green]{With the same computing power, a comparable 1D-RISM calculation takes about 10 seconds.} The non-linear solver converges in 8 iterations with about 380
 Jacobian applications.
@@
 for $n=5$ and $n=5+1$ explicit water molecules, the solvation energy,
-calculated by Eq.~\ref{eq:dG-sol-definition}, is by about 5~kcal
+calculated by Eq.~\ref{eq:dG-sol-definition}, is by about \hlc[green]{5~kcal/mol}
 higher (in absolute terms) in the latter case. This is likely an artifact
@@
 
-\centering{}\includegraphics[width=1\linewidth]{4w-composite}
+\centering{}\includegraphics[width=1\linewidth]{figure1}
 \end{figure}
@@
 The under-coordinated species does not feature a vacancy in the first
-solvation shell. Instead, all water ligands are equivalent due to
+solvation shell \hlc[green]{without the symmetry constraint}. Instead, all water ligands are equivalent due to
 symmetry, also for the five- and six-coordinated uranyl. These models
@@
 
-\centering{}\includegraphics[width=1\linewidth]{fig2}
+\centering{}\includegraphics[width=1\linewidth]{figure2}
 \end{figure}
@@
 
-\centering{}\includegraphics[width=1\linewidth]{fig3}
+\centering{}\includegraphics[width=1\linewidth]{figure3}
 \end{figure}
@@
 
-\centering{}\includegraphics[width=1\linewidth]{pmf}
+\centering{}\includegraphics[width=1\linewidth]{figure4}
 \end{figure}
@@
 
-\centering{}\includegraphics[width=1\linewidth]{pmf-2d}
+\centering{}\includegraphics[width=1\linewidth]{figure5}
 \end{figure}
@@
 \centering{}\smallskip{}
-\includegraphics[width=1\linewidth]{profiles}
+\includegraphics[width=1\linewidth]{figure6}
 \end{figure}
@@
 \begin{tocentry}
-\includegraphics[height=3.5cm]{4w-composite}
+\includegraphics[height=3.5cm]{toc}
 \end{tocentry}

### 2015-06-17 doc/uo2_paper/uo2-II.tex
AM: changes by BL sent out on Jun 16th

@@
 
-\title{Uranyl solvation by a three-dimensional reference interaction site
-model}
+\title{\hlc[green]{Uranyl Solvation by a Three-Dimensional Reference Interaction Site
+Model}}
 \author{Alexei Matveev}
@@
 E5630 @ 2.53~GHz CPUs, where one achieves a parallel efficiency above
-40\%. \hlc[green]{With the same computing power, a comparable 1D-RISM calculation takes about 10 seconds.} The non-linear solver converges in 8 iterations with about 380
+40\%. \hlc[green]{With the same computing equipment, a comparable 1D-RISM calculation takes about 10 seconds.} The non-linear solver converges in 8 iterations with about 380
 Jacobian applications.
@@
 \caption{Radial distribution functions of solvated uranyl without explicit
-aqua ligands: U--O$_{\text{w}}$ (top panel, black), U--H$_{\text{w}}$
-(top panel, gray), O--O$_{\text{w}}$ (bottom panel, black) and O--H$\text{\ensuremath{_{w}}}$
-(bottom panel, gray). Results obtained with the PSE1 closure are shown
-as thick lines in the foreground, results obtained with the closure
-relations PSE2 and PSE3 are displayed in the background using thinner
-lines.  \label{fig:rdfs-vs-closure-0w}}
+aqua ligands: U--O$_{\text{w}}$ (top panel, \hlc[green]{red}), U--H$_{\text{w}}$
+(top panel, \hlc[green]{blue}), O--O$_{\text{w}}$ (bottom panel, \hlc[green]{red}) and O--H$\text{\ensuremath{_{w}}}$ (bottom panel, \hlc[green]{blue}).
+Results obtained with the PSE1 closure are shown
+as \hlc[green]{solid lines}, results obtained with the closure relations PSE2 and PSE3 are displayed \hlc[green]{using dashed lines and dotted lines respectively}.  \label{fig:rdfs-vs-closure-0w}}
 
@@
 distribution function, determined with the closure relations PSE1,
-PSE2 and PSE3 (solid lines of decreasing weight). The dashed line
+PSE2 and PSE3, \hlc[green]{are displayed as solid, dashed and dot-dashed lines respectively}. The \hlc[green]{dotted} line
 represents the MD result of Ref.~\citenum{Bo15}.

### 2015-06-17 doc/uo2_paper/uo2-II.tex
AM: rm redundant spaces in hlc-macro
AM: comma, word wrap

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-\newcommand{\hlc}[2][yellow]{ {\sethlcolor{#1} \hl{#2}} }
+\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
 \usepackage{mathptmx}
@@
 E5630 @ 2.53~GHz CPUs, where one achieves a parallel efficiency above
-40\%. \hlc[green]{With the same computing equipment, a comparable 1D-RISM calculation takes about 10 seconds.} The non-linear solver converges in 8 iterations with about 380
-Jacobian applications.
+40\%. \hlc[green]{With the same computing equipment, a comparable
+1D-RISM calculation takes about 10 seconds.} The non-linear solver
+converges in 8 iterations with about 380 Jacobian applications.
 
@@
 \begin{figure}[tbph]
-\caption{Radial distribution functions of solvated uranyl without explicit
-aqua ligands: U--O$_{\text{w}}$ (top panel, \hlc[green]{red}), U--H$_{\text{w}}$
-(top panel, \hlc[green]{blue}), O--O$_{\text{w}}$ (bottom panel, \hlc[green]{red}) and O--H$\text{\ensuremath{_{w}}}$ (bottom panel, \hlc[green]{blue}).
-Results obtained with the PSE1 closure are shown
-as \hlc[green]{solid lines}, results obtained with the closure relations PSE2 and PSE3 are displayed \hlc[green]{using dashed lines and dotted lines respectively}.  \label{fig:rdfs-vs-closure-0w}}
+
+\caption{Radial distribution functions of solvated uranyl without
+  explicit aqua ligands: U--O$_{\text{w}}$ (top panel,
+  \hlc[green]{red}), U--H$_{\text{w}}$ (top panel, \hlc[green]{blue}),
+  O--O$_{\text{w}}$ (bottom panel, \hlc[green]{red}) and
+  O--H$\text{\ensuremath{_{w}}}$ (bottom panel,
+  \hlc[green]{blue}). Results obtained with the PSE1 closure are shown
+  as \hlc[green]{solid lines}, results obtained with the closure
+  relations PSE2 and PSE3 are displayed \hlc[green]{using dashed lines
+    and dotted lines respectively}.  \label{fig:rdfs-vs-closure-0w}}
 
@@
 \begin{centering}
-\caption{Potential of mean force, $W(r)$, from the U--O$_{\text{w}}$ radial
-distribution function, determined with the closure relations PSE1,
-PSE2 and PSE3, \hlc[green]{are displayed as solid, dashed and dot-dashed lines respectively}. The \hlc[green]{dotted} line
+
+\caption{Potential of mean force, $W(r)$, from the U--O$_{\text{w}}$
+  radial distribution function, determined with the closure relations
+  PSE1, PSE2 and PSE3, \hlc[green]{are displayed as solid, dashed and
+  dot-dashed lines, respectively}. The \hlc[green]{dotted} line
   represents the MD result of Ref.~\citenum{Bo15}.

### 2015-06-17 doc/uo2_paper/uo2-II.tex
AM: there IS a vacany without the symmetry constraints

@@
 In QM calculations, we use symmetric models of uranyl aqua complexes
-to compare PCM and RISM representations of the solvent in QM calculations.
-The under-coordinated species does not feature a vacancy in the first
-solvation shell \hlc[green]{without the symmetry constraint}. Instead, all water ligands are equivalent due to
-symmetry, also for the five- and six-coordinated uranyl. These models
-feature an additional intra-molecular degree of freedom, the uranyl
-bond length, in contrast to the MM models discussed in the previous
-section where the uranyl bond was fixed at a reasonable value.
+to compare PCM and RISM representations of the solvent in QM
+calculations.  \hlc[cyan]{The symmetry constraints for the five- and
+  six-coordinated uranyl are qualitatively consistent with the
+  structures obtained by unconstrained optimization (see next
+  section). However, the under-coordinated species does not feature a
+  vacancy in the first solvation shell as in an unconstrained
+  structure. Instead, all water ligands are evenly distributed in the
+  equatorial plane.}  These models feature an additional
+intra-molecular degree of freedom, the uranyl bond length, in contrast
+to the MM models discussed in the previous section where the uranyl
+bond was fixed at a reasonable value.
 

### 2015-06-19 doc/uo2_paper/uo2-II.tex
AM: version sent by BL on Jun 18

--- 2015-06-17 doc/uo2_paper/uo2-II.tex
@@
 \usepackage[T1]{fontenc}
-\usepackage{color,soul}
+\usepackage{color}
 \usepackage{amsmath}
@@
 
-\title{\hlc[green]{Uranyl Solvation by a Three-Dimensional Reference Interaction Site
-Model}}
+\title{Uranyl Solvation by a Three-Dimensional Reference Interaction Site
+Model}
 \author{Alexei Matveev}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
 \usepackage{mathptmx}
@@
 E5630 @ 2.53~GHz CPUs, where one achieves a parallel efficiency above
-40\%. \hlc[green]{With the same computing equipment, a comparable
-1D-RISM calculation takes about 10 seconds.} The non-linear solver
+40\%. With the same computing equipment, a comparable
+1D-RISM calculation takes about 10 seconds. The non-linear solver
 converges in 8 iterations with about 380 Jacobian applications.
@@
 for $n=5$ and $n=5+1$ explicit water molecules, the solvation energy,
-calculated by Eq.~\ref{eq:dG-sol-definition}, is by about \hlc[green]{5~kcal/mol}
+calculated by Eq.~\ref{eq:dG-sol-definition}, is by about 5~kcal/mol
 higher (in absolute terms) in the latter case. This is likely an artifact
@@
 
-In QM calculations, we use symmetric models of uranyl aqua complexes
-to compare PCM and RISM representations of the solvent in QM calculations.
-The under-coordinated species does not feature a vacancy in the first
-solvation shell \hlc[green]{without the symmetry constraint}. Instead, all water ligands are equivalent due to
-symmetry, also for the five- and six-coordinated uranyl. These models
-feature an additional intra-molecular degree of freedom, the uranyl
-bond length, in contrast to the MM models discussed in the previous
-section where the uranyl bond was fixed at a reasonable value.
-
-When previously exploring these models with 1D RISM \citep{Bo15},
+In QM+3D-RISM calculations, we use symmetric models of uranyl 
+aqua complexes to facilitate a comparison with previous results 
+obtained with PCM and 1D-RISM representations of the solvent. 
+In these symmetric models all aqua ligands are evenly distributed 
+in the equatorial plane. The symmetric structures of the five- and 
+six-coordinated uranyl aqua complexes are qualitatively consistent 
+with the structures obtained by unconstrained optimization 
+(see next section). In contrast, the unconstrained four-coordinated 
+species features a vacancy in the first solvation shell, filled by 
+the RISM solvent. These QM models feature an additional intra-molecular 
+degree of freedom, the uranyl bond length, in contrast to the MM models 
+discussed in the previous section where the uranyl bond was fixed at 
+a reasonable value.
+
+When previously exploring these symmetric models with 1D RISM \citep{Bo15},
 we observed consistently longer U--O and U--O$_{\text{w}}$ bond lengths
@@
   explicit aqua ligands: U--O$_{\text{w}}$ (top panel,
-  \hlc[green]{red}), U--H$_{\text{w}}$ (top panel, \hlc[green]{blue}),
-  O--O$_{\text{w}}$ (bottom panel, \hlc[green]{red}) and
+  red), U--H$_{\text{w}}$ (top panel, blue),
+  O--O$_{\text{w}}$ (bottom panel, red) and
   O--H$\text{\ensuremath{_{w}}}$ (bottom panel,
-  \hlc[green]{blue}). Results obtained with the PSE1 closure are shown
-  as \hlc[green]{solid lines}, results obtained with the closure
-  relations PSE2 and PSE3 are displayed \hlc[green]{using dashed lines
-    and dotted lines respectively}.  \label{fig:rdfs-vs-closure-0w}}
+  blue). Results obtained with the PSE1 closure are shown
+  as solid lines, results obtained with the closure
+  relations PSE2 and PSE3 are displayed using dashed lines
+    and dotted lines respectively.  \label{fig:rdfs-vs-closure-0w}}
 
@@
   radial distribution function, determined with the closure relations
-  PSE1, PSE2 and PSE3, \hlc[green]{are displayed as solid, dashed and
-  dot-dashed lines, respectively}. The \hlc[green]{dotted} line
+  PSE1, PSE2 and PSE3, are displayed as solid, dashed and
+  dot-dashed lines, respectively. The dotted line
   represents the MD result of Ref.~\citenum{Bo15}.
